### 2017-02-02 pyked.tex
Add files for our actual paper and remove template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This work is licensed under the Creative Commons Attribution 4.0
% International License. To view a copy of this license, visit
% http://creativecommons.org/licenses/by/4.0/.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{ussci}

%======================================================================
\addbibresource{pyked.bib}
%======================================================================
\newcommand\papertopic{Reaction Kinetics}
%======================================================================

\title{ ChemKED: a human- and machine-readable data standard for chemical kinetics experiments }

\author[1*]{Bryan W. Weber}
\author[2]{Kyle E. Niemeyer}

\affil[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT, USA}
\affil[2]{Mechanical, Industrial, and Manfacturing Engineering, Oregon State University, Corvalis, OR, USA}
\affil[*]{Corresponding author: \email{bryan.weber@uconn.edu}}

\begin{document}
\maketitle

%====================================================================
\begin{abstract} % not to exceed 200 words
Fundamental experimental measurements of quantities including ignition delay
times, laminar flame speeds, and species profiles serve important roles in
understanding fuel chemistry and validating chemical kinetic models. However,
despite both the importance and abundance of such information in the literature,
the community lacks a widely adopted standard format for this data. This impedes
both sharing and wide use by the community. In this talk, we will introduce a
new Chemical Kinetics Experimental Data format, ChemKED, and the related
Python-based package for creating and validating ChemKED-formatted files called
PyKED. We will also review past and related efforts, and motivate the need for a
new solution. ChemKED currently supports the representation of autoignition
delay time and laminar flame speed measurements. ChemKED-formatted files contain
all of the information needed to simulate experimental data points, including
uncertainty. ChemKED is based on the YAML data serialization language, and is
intended as a human- and machine-readable standard for easy creation and
automated use. Development of ChemKED and PyKED occurs openly on GitHub under
the BSD 3-clause license, and contributions from the community are welcome.
Plans for future development include support for experimental data from jet
stirred reactor, extinction, and speciation measurements.
\end{abstract}

% (Provide 2-4 keywords describing your research. Only abbreviations firmly
% established in the field may be used. These keywords will be used for
% sessioning/indexing purposes.)
\begin{keyword}
    chemical kinetics\sep software\sep database
\end{keyword}

%====================================================================
\section{Introduction}
%

\section{Conclusions}
%
\blindtext

\blindtext

\section{Acknowledgements}
This research was funded by \ldots

\printbibliography[heading=bibintoc]

\end{document}

### 2017-02-14 pyked.tex
added (very) basic structure
First commit of my part

@@
+% !TEX option = -shell-escape
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\usepackage{todonotes}
+\newcommand\ck{ChemKED}
+\usepackage{enumitem}
+\setlist{noitemsep}
 %======================================================================
@@
 %
+\todo[inline]{Kyle to do}
 
-\section{Conclusions}
+%====================================================================
+\section{Related work}
 %
-\blindtext
+\todo[inline]{Kyle to do}
 
-\blindtext
+%====================================================================
+\section{Overview of format}\label{sec:overview-of-format}
 
+The \ck{} file is a YAML-format file. This format offers the advantage of being
+human readable and having parsers in most common programming languages,
+including Python, C++, Java, Perl, MATLAB, and many others. The YAML syntax is
+quite simple. The basic file structure is made of mappings, delimited by a colon:
+
+\begin{minted}{yaml}
+    key: value
+    key: 10 # This is because YAML syntax highlighting in Atom is busted
+\end{minted}
+
+In addition, sequences can be specified by using a hypen \yaml|-| or square
+brackets
+
+\begin{minted}{yaml}
+    - [0, 1, 2]
+    key:
+      - 0
+      - 1
+      - 2
+\end{minted}
+
+The \ck{} format is designed
+to include all of the information necessary to simulate a given experiment. Most
+of the required information is specified as a \yaml|key| with a \yaml|value|.
+
+In
+general, this includes
+
+\todo{Make this smaller}
+\begin{itemize}
+    \item The type of experiment
+    \begin{itemize}
+        \item Ignition delay
+        \item Flame speed
+        \item Speciation
+        \item Extinction
+    \end{itemize}
+    \item The experimental apparatus
+    \begin{itemize}
+        \item Shock tube
+        \item Rapid compression machine
+        \item Counterflow flame
+        \item Burner-stabilized flame
+    \end{itemize}
+    \item The thermodynamic conditions of the experiment
+    \begin{itemize}
+        \item Temperature
+        \item Pressure
+        \item Composition
+    \end{itemize}
+    \item Apparatus specific effects
+    \begin{itemize}
+        \item Non-ideal pressure rise in shock tubes
+        \item Post-compression heat loss in rapid compression machines
+    \end{itemize}
+\end{itemize}
+
+%====================================================================
+\section{Use example}
+%
+\todo[inline]{Bryan/Kyle to do}
+
+%====================================================================
+\section{Conclusions and future work}
+%
+\todo[inline]{both to do}
+
+%====================================================================
 \section{Acknowledgements}
 This research was funded by \ldots
+\todo[inline]{update}
 
-\printbibliography[heading=bibintoc]
+\printbibliography
 

### 2017-02-20 pyked.tex
fixed my address
Add info about some of the keys in the ChemKED file

--- 2017-02-02 pyked.tex
@@
 
+\usepackage{todonotes}
+\newcommand\ck{ChemKED}
+\usepackage[inline]{enumitem}
+\setlist{noitemsep}
 %======================================================================
@@
 \affil[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT, USA}
-\affil[2]{Mechanical, Industrial, and Manfacturing Engineering, Oregon State University, Corvalis, OR, USA}
+\affil[2]{School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
 \affil[*]{Corresponding author: \email{bryan.weber@uconn.edu}}
@@
 %
+\todo[inline]{Kyle to do}
 
-\section{Conclusions}
+%====================================================================
+\section{Related work}
 %
-\blindtext
+\todo[inline]{Kyle to do}
 
-\blindtext
+%====================================================================
+\section{Overview of format}\label{sec:overview-of-format}
 
+The \ck{} file is a YAML-format file. This format offers the advantage of being
+human readable, written in plain text, and having parsers in most common
+programming languages, including Python, C++, Java, Perl, MATLAB, and many
+others. The YAML syntax is quite simple. The basic file structure is made of
+mappings, delimited by a colon. The key for the mapping is on the left of the
+colon, and the value on the right can be a single value, a sequence of values, or a nested mapping:
+
+\begin{minted}{yaml}
+    key1: value  # Single-value mapping
+    key2: [0, 1, 2]  # Mapping with sequence
+    key3:  # Alternative sequence format
+      - 0
+      - 1
+    key4:  # Nested mapping
+      key5: 0
+\end{minted}
+
+The \yaml|value| can be a string, integer, or floating point number. The type is
+inferred from the \yaml|key| when the value is read. The \ck{} format is
+designed to include all of the information necessary to simulate a given
+experiment. The \ck{} format can be broken into two main sections. The first
+section details all of the ``meta'' information about the experiment and the
+\ck{} file. In this section, the common \yaml|author| mapping is used for any
+fields where an author must be specified. The \yaml|author| mapping contains the
+following fields:
+%
+\begin{itemize}
+    \item \yaml|name| (required, string): The author's full name
+    \item \yaml|ORCID| (optional, string): the author's ORCID number
+\end{itemize}
+%
+More generally for the meta information, the required keys in this section
+are:
+%
+\begin{itemize}
+    \item \yaml|file-author| (required, \yaml|author|-type mapping): The author
+    of the \ck{} file
+    \item \yaml|file-version| (required, integer): The version of the \ck{} file
+    \item \yaml|chemked-version| (required, string): The version of the \ck{}
+    schema that this file conforms to
+    \item \yaml|reference| (required, mapping): The reference information for
+    the article associated with the data in the file. The fields in this
+    mapping are:
+    \begin{itemize}
+        \item \yaml|authors| (required, sequence): A sequence of \yaml|author| mappings
+        \item \yaml|journal| (required, string): The name of the publishing journal
+        \item \yaml|year| (required, integer): The year of publication
+    \end{itemize}
+    \item \yaml|experiment-type| (required, string): One of \yaml|ignition delay| or \yaml|flame speed|
+    \item \yaml|apparatus| (required, string): The type of apparatus used to perform the experiment
+\end{itemize}
+
+In addition to the meta information, the data encoded in the \ck{} file is in
+a mapping called \yaml|datapoints|. The \yaml|datapoints| mapping contains a
+sequence of nested mappings, with the following information for each element of
+the sequence:
+%
+\begin{itemize}
+    \item \yaml|temperature| (required, sequence): The temperature of the experiment, with units
+    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the experiment, with units, if the \yaml|experiment-type| is \yaml|ignition delay|
+    \item \yaml|flame-speed| (optional, sequence): The flame speed of the experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
+    \item \yaml|pressure| (required, string): The pressure of the experiment, with units
+    \item \yaml|composition| (required, sequence): The composition of the mixture in the experiment, a sequence of mappings with the following fields
+    \begin{itemize}
+        \item \yaml|species-name| (required, string): The name of the species
+        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent| (required, float): The mole fraction or percent, or the mass fraction.
+Only one of the three is permitted in a given \yaml|composition| sequence.
+    \end{itemize}
+    \item \yaml|ignition-type| (required, mapping); The type of ignition measured in the experiments. The required fields are
+    \begin{itemize}
+        \item \yaml|type| (required, string): How the ignition delay was measured, one of \yaml|d/dt max| to indicate the ignition point is found at the maximum of the time derivative of the \yaml|target|, or \yaml|max| or \yaml|min| to indicate the ignition point is at the maximum or minimum of the \yaml|target|
+        \item \yaml|target| (required, string): The target for the type of the ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|, \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
+    \end{itemize}
+\end{itemize}
+
+%====================================================================
+\section{PyKED Architecture}
+%
+\todo[inline]{Bryan to do}
+
+%====================================================================
+\section{Use example}
+%
+\todo[inline]{Bryan/Kyle to do}
+
+%====================================================================
+\section{Conclusions and future work}
+%
+\todo[inline]{both to do}
+
+%====================================================================
 \section{Acknowledgements}
 This research was funded by \ldots
+\todo[inline]{update}
 
-\printbibliography[heading=bibintoc]
+\printbibliography
 

### 2017-02-21 pyked.tex
note in acks
added some software references for PyKED

@@
 \newcommand\ck{ChemKED}
+\newcommand\pk{PyKED}
 \usepackage[inline]{enumitem}
@@
 \todo[inline]{Kyle to do}
+
+Fundamental combustion experiments provide
 
@@
 
+\pk{} relies on well-established scientific Python software tools.
+These include NumPy~\cite{vanderWalt:2011np} for array manipulation, Pint~\cite{Grecco2016} for
+interpreting and converting between units, \texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{}
+Cerberus for validating \ck{} files\todo{needs ref}{},
+and pytest~\cite{pytest:3.0.1} for unit testing.
+Travis-CI~\cite{Travis2016} also provides continuous integration testing.
+
 %====================================================================
@@
 \todo[inline]{update}
+% Kyle: I don't think I have any funding yet to acknowledge for this
 

### 2017-02-21 pyked.tex
mostly spacing updates

--- 2017-02-20 pyked.tex
@@
 
-The \ck{} file is a YAML-format file. This format offers the advantage of being
+The \ck{} file is a YAML-format file. This format offers the advantages of being
 human readable, written in plain text, and having parsers in most common
@@
 mappings, delimited by a colon. The key for the mapping is on the left of the
-colon, and the value on the right can be a single value, a sequence of values, or a nested mapping:
+colon, and the value on the right can be a single value, a sequence of values,
+a nested mapping, or some combination of these:
 
@@
     key3:  # Alternative sequence format
-      - 0
-      - 1
+      - value
+      - value
     key4:  # Nested mapping
@@
 designed to include all of the information necessary to simulate a given
-experiment. The \ck{} format can be broken into two main sections. The first
-section details all of the ``meta'' information about the experiment and the
-\ck{} file. In this section, the common \yaml|author| mapping is used for any
-fields where an author must be specified. The \yaml|author| mapping contains the
-following fields:
+experiment and can be broken into two main sections. The first section details
+all of the ``meta'' information about the experiment and the \ck{} file. In this
+section, the common \yaml|author| mapping is used for any fields where an author
+must be specified. The \yaml|author| mapping contains the following fields:
 %
@@
     \begin{itemize}
-        \item \yaml|authors| (required, sequence): A sequence of \yaml|author| mappings
-        \item \yaml|journal| (required, string): The name of the publishing journal
+        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
+        mappings
+        \item \yaml|journal| (required, string): The name of the publishing
+        journal
         \item \yaml|year| (required, integer): The year of publication
     \end{itemize}
-    \item \yaml|experiment-type| (required, string): One of \yaml|ignition delay| or \yaml|flame speed|
-    \item \yaml|apparatus| (required, string): The type of apparatus used to perform the experiment
+    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
+    delay| or \yaml|flame speed|
+    \item \yaml|apparatus| (required, string): The type of apparatus used to
+    perform the experiment
 \end{itemize}
@@
 \begin{itemize}
-    \item \yaml|temperature| (required, sequence): The temperature of the experiment, with units
-    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the experiment, with units, if the \yaml|experiment-type| is \yaml|ignition delay|
-    \item \yaml|flame-speed| (optional, sequence): The flame speed of the experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
-    \item \yaml|pressure| (required, string): The pressure of the experiment, with units
-    \item \yaml|composition| (required, sequence): The composition of the mixture in the experiment, a sequence of mappings with the following fields
+    \item \yaml|temperature| (required, sequence): The temperature of the
+    experiment, with units
+    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the
+    experiment, with units, if the \yaml|experiment-type| is \yaml|ignition
+    delay|
+    \item \yaml|flame-speed| (optional, sequence): The flame speed of the
+    experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
+    \item \yaml|pressure| (required, string): The pressure of the experiment,
+    with units
+    \item \yaml|composition| (required, sequence): The composition of the
+    mixture in the experiment, a sequence of mappings with the following fields
     \begin{itemize}
         \item \yaml|species-name| (required, string): The name of the species
-        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent| (required, float): The mole fraction or percent, or the mass fraction.
-Only one of the three is permitted in a given \yaml|composition| sequence.
+        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent|
+        (required, float): The mole fraction or percent, or the mass fraction.
+        Only one of the three is permitted in a given \yaml|composition|
+        sequence.
     \end{itemize}
-    \item \yaml|ignition-type| (required, mapping); The type of ignition measured in the experiments. The required fields are
+    \item \yaml|ignition-type| (required, mapping); The type of ignition
+    measured in the experiments. The required fields are
     \begin{itemize}
-        \item \yaml|type| (required, string): How the ignition delay was measured, one of \yaml|d/dt max| to indicate the ignition point is found at the maximum of the time derivative of the \yaml|target|, or \yaml|max| or \yaml|min| to indicate the ignition point is at the maximum or minimum of the \yaml|target|
-        \item \yaml|target| (required, string): The target for the type of the ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|, \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
+        \item \yaml|type| (required, string): How the ignition delay was
+        measured, one of \yaml|d/dt max| to indicate the ignition point is
+        found at the maximum of the time derivative of the \yaml|target|, or
+        \yaml|max| or \yaml|min| to indicate the ignition point is at the
+        maximum or minimum of the \yaml|target|
+        \item \yaml|target| (required, string): The target for the type of the
+        ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|,
+        \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
     \end{itemize}

### 2017-02-21 pyked.tex
Add idea for usage example
Remove acknowledgements
Update chemked spec

@@
         sequence.
+        \item \yaml|InChI|, \yaml|SMILES|, \yaml|atomic-composition|, or
+        \yaml|elemental-composition| (required, string or sequence): The InChI
+        or SMILES representing the molecule, or its atomic or elemental
+        composition (\yaml|atomic-composition| and \yaml|elemental-composition|
+        are synonyms) as a sequence
     \end{itemize}
@@
 
+Each of the quantities in a datapoint must be specified with units. This should
+be done as a single string associated with the first element of the sequence.
+The units of the quantity are validated to ensure they have the appropriate
+dimensionality for the quanity. In addition, each of the quantities in the
+\yaml|datapoint| can optionally be assigned an uncertainty. This uncertainty can
+be either absolute or relative, and is specified as elements of the sequence of
+the associated key. For example, the absolute uncertainty of the temperature and
+the relative uncertainty of the ignition delay might be specified as follows:
+
+\begin{minted}{yaml}
+datapoints:
+  - temperature:
+      - 1100 kelvin
+      - uncertainty-type: absolute
+        uncertainty: 10 kelvin
+    ignition-delay:
+      - 10 us
+      - uncertainty-type: relative
+        uncertainty: 10
+    ...
+\end{minted}
+
 %====================================================================
@@
 \todo[inline]{Bryan/Kyle to do}
+\todo[inline]{Idea: CONV simulations using Cantera in a loop with conditions
+from the Pandas output for butanol?}
 
@@
 
-%====================================================================
-\section{Acknowledgements}
-This research was funded by \ldots
-\todo[inline]{update}
-% Kyle: I don't think I have any funding yet to acknowledge for this
-
 \printbibliography

### 2017-02-21 pyked.tex
Another usage example idea
First parts of PyKED section

@@
       - uncertainty-type: relative
-        uncertainty: 10
+        uncertainty: 0.1
     ...
@@
 %
-\todo[inline]{Bryan to do}
+\pk{} is a Python package that currently provides the reference implementation
+of the interface to \ck{} files. \pk{} reads \ck{} files, validates their
+structure and content, and allows the user to interact with the datapoints
+contained in the \ck{} file.
+
+The basic user interface to the \ck{} file in \pk{} is through the
+\python|ChemKED| class. The \python|ChemKED| class constructor takes the file
+name of a \ck{} file or a Python dictionary containing the contents of a \ck{}
+file as its argument. When the file or dictionary is loaded, it is validated to
+have the correct format by Cerberus \autocite{cerberus}. The fields of the \ck{}
+file are stored as instance attributes of the \python|ChemKED| class.
 
@@
 from the Pandas output for butanol?}
+\todo[inline]{Idea: VPRO simulations using Cantera from a \ck{} file for an RCM
+datapoint}
 

### 2017-02-22 pyked.tex
incomplete additions on related work

--- 2017-02-21 pyked.tex
@@
 \newcommand\ck{ChemKED}
+\newcommand\pk{\texttt{PyKED}}
 \usepackage[inline]{enumitem}
@@
 
+Fundamental combustion experiments provide
+
 %====================================================================
@@
 
+The combustion\slash chemical kinetics community lacks an accepted, commonly used
+standard for recording and sharing data from fundamental combustion experiments.
+Interestingly, this contrasts the situation for chemical kinetic models, where the
+CHEMKIN~\cite{Kee:1996ck} format dominates. Competing standards such as the
+Cantera~\cite{Cantera:2.3.0} CTI format or FlameMaster lag behind considerably, although
+use of the former continually grows due to Cantera's open availability.\todo{add flamemaster citation}{}
+Similarly,
+
+Instead, such data must typically be found in under-documented, nonstandard comma-separated value
+files and Excel spreadsheets, or even contained in PDF tables or figures rather than as
+supplementary material associated with a paper.
+
+While the community has not widely adopted a standard format,
+---most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
+\textbf{M}od\textbf{e}l) data format~\cite{Frenklach:2007bm,You:2011hy} and an associated
+online database, the PrIMe Data Warehouse~\cite{PrIMe}.
+
+More recently, Varga et al.~\cite{Varga2015a,Varga2015b} developed the \texttt{ReSpecTh} standard,
+an XML-based format that builds on, and is compatible with, PrIMe.
+
+We aim to develop a new, human- and machine-readable data standard for fundamental combustion
+experiments, \ck{}, and offer tools for easily working with data encoded in this format.
+Our motivations are similar to some of those of the PrIMe and \texttt{ReSpecTh} teams, although
+less ambitious in scope: enable easy sharing and use of data, for the (primary) purposes of
+developing and validating predictive chemical kinetic models.
+However, those efforts were generally less successful in being adopted---we hope to succeed
+by developing more user-friendly data files and tools for interacting with them.
+
+Niemeyer recently introduced an initial version of the \ck{} format~\cite{Niemeyer:2016wf},
+and in this work we further formalize and develop the standard. We also discuss a related
+Python software tool for validating and working with \ck{} files: \pk{}.
+
 %====================================================================
@@
 
-The \ck{} file is a YAML-format file. This format offers the advantages of being
+The \ck{} file is a YAML-format file. This format offers the advantage of being
 human readable, written in plain text, and having parsers in most common
@@
 mappings, delimited by a colon. The key for the mapping is on the left of the
-colon, and the value on the right can be a single value, a sequence of values,
-a nested mapping, or some combination of these:
+colon, and the value on the right can be a single value, a sequence of values, or a nested mapping:
 
@@
     key3:  # Alternative sequence format
-      - value
-      - value
+      - 0
+      - 1
     key4:  # Nested mapping
@@
 designed to include all of the information necessary to simulate a given
-experiment and can be broken into two main sections. The first section details
-all of the ``meta'' information about the experiment and the \ck{} file. In this
-section, the common \yaml|author| mapping is used for any fields where an author
-must be specified. The \yaml|author| mapping contains the following fields:
+experiment. The \ck{} format can be broken into two main sections. The first
+section details all of the ``meta'' information about the experiment and the
+\ck{} file. In this section, the common \yaml|author| mapping is used for any
+fields where an author must be specified. The \yaml|author| mapping contains the
+following fields:
 %
@@
     \begin{itemize}
-        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
-        mappings
-        \item \yaml|journal| (required, string): The name of the publishing
-        journal
+        \item \yaml|authors| (required, sequence): A sequence of \yaml|author| mappings
+        \item \yaml|journal| (required, string): The name of the publishing journal
         \item \yaml|year| (required, integer): The year of publication
     \end{itemize}
-    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
-    delay| or \yaml|flame speed|
-    \item \yaml|apparatus| (required, string): The type of apparatus used to
-    perform the experiment
+    \item \yaml|experiment-type| (required, string): One of \yaml|ignition delay| or \yaml|flame speed|
+    \item \yaml|apparatus| (required, string): The type of apparatus used to perform the experiment
 \end{itemize}
@@
 \begin{itemize}
-    \item \yaml|temperature| (required, sequence): The temperature of the
-    experiment, with units
-    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the
-    experiment, with units, if the \yaml|experiment-type| is \yaml|ignition
-    delay|
-    \item \yaml|flame-speed| (optional, sequence): The flame speed of the
-    experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
-    \item \yaml|pressure| (required, string): The pressure of the experiment,
-    with units
-    \item \yaml|composition| (required, sequence): The composition of the
-    mixture in the experiment, a sequence of mappings with the following fields
+    \item \yaml|temperature| (required, sequence): The temperature of the experiment, with units
+    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the experiment, with units, if the \yaml|experiment-type| is \yaml|ignition delay|
+    \item \yaml|flame-speed| (optional, sequence): The flame speed of the experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
+    \item \yaml|pressure| (required, string): The pressure of the experiment, with units
+    \item \yaml|composition| (required, sequence): The composition of the mixture in the experiment, a sequence of mappings with the following fields
     \begin{itemize}
         \item \yaml|species-name| (required, string): The name of the species
-        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent|
-        (required, float): The mole fraction or percent, or the mass fraction.
-        Only one of the three is permitted in a given \yaml|composition|
-        sequence.
+        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent| (required, float): The mole fraction or percent, or the mass fraction.
+Only one of the three is permitted in a given \yaml|composition| sequence.
     \end{itemize}
-    \item \yaml|ignition-type| (required, mapping); The type of ignition
-    measured in the experiments. The required fields are
+    \item \yaml|ignition-type| (required, mapping); The type of ignition measured in the experiments. The required fields are
     \begin{itemize}
-        \item \yaml|type| (required, string): How the ignition delay was
-        measured, one of \yaml|d/dt max| to indicate the ignition point is
-        found at the maximum of the time derivative of the \yaml|target|, or
-        \yaml|max| or \yaml|min| to indicate the ignition point is at the
-        maximum or minimum of the \yaml|target|
-        \item \yaml|target| (required, string): The target for the type of the
-        ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|,
-        \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
+        \item \yaml|type| (required, string): How the ignition delay was measured, one of \yaml|d/dt max| to indicate the ignition point is found at the maximum of the time derivative of the \yaml|target|, or \yaml|max| or \yaml|min| to indicate the ignition point is at the maximum or minimum of the \yaml|target|
+        \item \yaml|target| (required, string): The target for the type of the ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|, \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
     \end{itemize}
@@
 \todo[inline]{Bryan to do}
+
+\pk{} relies on well-established scientific Python software tools.
+These include NumPy~\cite{vanderWalt:2011np} for array manipulation, Pint~\cite{Grecco2016} for
+interpreting and converting between units, \texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{}
+Cerberus for validating \ck{} files\todo{needs ref}{},
+and pytest~\cite{pytest:3.0.1} for unit testing.
+Travis-CI~\cite{Travis2016} also provides continuous integration testing.
+
+\todo[inline]{Need to archive version of PyKED and cite}
+
+\pk{} is available under the permissive, open-source BSD 3-clause license via a GitHub
+repository: \url{https://github.com/pr-omethe-us/PyKED}. It can be installed using setuptools
+by downloading the source code files and executing \texttt{python setup.py install}. More mature
+versions of \pk{} will be distributed on \href{https://pypi.python.org/pypi}{PyPI}
+(Python Package Index) and conda-forge.\todo{add URL}{}
 

### 2017-02-22 pyked.tex

--- 2017-02-21 pyked.tex
@@
 \newcommand\ck{ChemKED}
-\newcommand\pk{PyKED}
+\newcommand\pk{\texttt{PyKED}}
 \usepackage[inline]{enumitem}
@@
 \todo[inline]{Kyle to do}
+
+The combustion\slash chemical kinetics community lacks an accepted, commonly used
+standard for recording and sharing data from fundamental combustion experiments.
+Interestingly, this contrasts the situation for chemical kinetic models, where the
+CHEMKIN~\cite{Kee:1996ck} format dominates. Competing standards such as the
+Cantera~\cite{Cantera:2.3.0} CTI format or FlameMaster lag behind considerably, although
+use of the former continually grows due to Cantera's open availability.\todo{add flamemaster citation}{}
+Similarly,
+
+Instead, such data must typically be found in under-documented, nonstandard comma-separated value
+files and Excel spreadsheets, or even contained in PDF tables or figures rather than as
+supplementary material associated with a paper.
+
+While the community has not widely adopted a standard format,
+---most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
+\textbf{M}od\textbf{e}l) data format~\cite{Frenklach:2007bm,You:2011hy} and an associated
+online database, the PrIMe Data Warehouse~\cite{PrIMe}.
+
+More recently, Varga et al.~\cite{Varga2015a,Varga2015b} developed the \texttt{ReSpecTh} standard,
+an XML-based format that builds on, and is compatible with, PrIMe.
+
+We aim to develop a new, human- and machine-readable data standard for fundamental combustion
+experiments, \ck{}, and offer tools for easily working with data encoded in this format.
+Our motivations are similar to some of those of the PrIMe and \texttt{ReSpecTh} teams, although
+less ambitious in scope: enable easy sharing and use of data, for the (primary) purposes of
+developing and validating predictive chemical kinetic models.
+However, those efforts were generally less successful in being adopted---we hope to succeed
+by developing more user-friendly data files and tools for interacting with them.
+
+Niemeyer recently introduced an initial version of the \ck{} format~\cite{Niemeyer:2016wf},
+and in this work we further formalize and develop the standard. We also discuss a related
+Python software tool for validating and working with \ck{} files: \pk{}.
 
@@
 
+\pk{} relies on well-established scientific Python software tools.
+These include NumPy~\cite{vanderWalt:2011np} for array manipulation, Pint~\cite{Grecco2016} for
+interpreting and converting between units, \texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{}
+Cerberus for validating \ck{} files\todo{needs ref}{},
+and pytest~\cite{pytest:3.0.1} for unit testing.
+Travis-CI~\cite{Travis2016} also provides continuous integration testing.
+
+\todo[inline]{Need to archive version of PyKED and cite}
+
+\pk{} is available under the permissive, open-source BSD 3-clause license via a GitHub
+repository: \url{https://github.com/pr-omethe-us/PyKED}. It can be installed using setuptools
+by downloading the source code files and executing \texttt{python setup.py install}. More mature
+versions of \pk{} will be distributed on \href{https://pypi.python.org/pypi}{PyPI}
+(Python Package Index) and conda-forge.\todo{add URL}{}
+
 %====================================================================

### 2017-02-25 pyked.tex
PyKED shouldn't be tt
Switch citations to autocite and update bib

@@
 \newcommand\ck{ChemKED}
-\newcommand\pk{\texttt{PyKED}}
+\newcommand\pk{PyKED}
 \usepackage[inline]{enumitem}
@@
 Interestingly, this contrasts the situation for chemical kinetic models, where the
-CHEMKIN~\cite{Kee:1996ck} format dominates. Competing standards such as the
-Cantera~\cite{Cantera:2.3.0} CTI format or FlameMaster lag behind considerably, although
+CHEMKIN~\autocite{Kee:1996ck} format dominates. Competing standards such as the
+Cantera~\autocite{Cantera:2.3.0} CTI format or FlameMaster lag behind considerably, although
 use of the former continually grows due to Cantera's open availability.\todo{add flamemaster citation}{}
@@
 ---most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
-\textbf{M}od\textbf{e}l) data format~\cite{Frenklach:2007bm,You:2011hy} and an associated
-online database, the PrIMe Data Warehouse~\cite{PrIMe}.
+\textbf{M}od\textbf{e}l) data format~\autocite{Frenklach:2007bm,You:2011hy} and an associated
+online database, the PrIMe Data Warehouse~\autocite{PrIMe}.
 
-More recently, Varga et al.~\cite{Varga2015a,Varga2015b} developed the \texttt{ReSpecTh} standard,
+More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the \texttt{ReSpecTh} standard,
 an XML-based format that builds on, and is compatible with, PrIMe.
@@
 
-Niemeyer recently introduced an initial version of the \ck{} format~\cite{Niemeyer:2016wf},
+Niemeyer recently introduced an initial version of the \ck{} format~\autocite{Niemeyer:2016wf},
 and in this work we further formalize and develop the standard. We also discuss a related
@@
 
-\pk{} relies on well-established scientific Python software tools.
-These include NumPy~\cite{vanderWalt:2011np} for array manipulation, Pint~\cite{Grecco2016} for
-interpreting and converting between units, \texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{}
-Cerberus for validating \ck{} files\todo{needs ref}{},
-and pytest~\cite{pytest:3.0.1} for unit testing.
-Travis-CI~\cite{Travis2016} also provides continuous integration testing.
+\pk{} relies on well-established scientific Python software tools. These include
+NumPy~\autocite{vanderWalt:2011np} for array manipulation,
+Pint~\autocite{Grecco2016} for interpreting and converting between units,
+\texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{} Cerberus for
+validating \ck{} files~\autocite{cerberus}, and pytest~\autocite{pytest:3.0.1}
+for unit testing. Travis-CI~\autocite{Travis2016} also provides continuous
+integration testing.
 

### 2017-02-25 pyked.tex
Add more citations

@@
 Pint~\autocite{Grecco2016} for interpreting and converting between units,
-\texttt{ruamel.yaml} for parsing YAML files,\todo{needs ref}{} Cerberus for
+\texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML files, Cerberus for
 validating \ck{} files~\autocite{cerberus}, and pytest~\autocite{pytest:3.0.1}
-for unit testing. Travis-CI~\autocite{Travis2016} also provides continuous
+for unit testing. Travis-CI~\autocite{Travis2016} provides continuous
 integration testing.

### 2017-02-25 pyked.tex
Use microtype
Add more about PyKED arch

@@
 \usepackage{todonotes}
+
 \newcommand\ck{ChemKED}
 \newcommand\pk{PyKED}
+
 \usepackage[inline]{enumitem}
 \setlist{noitemsep}
+
+\usepackage{microtype}
 %======================================================================
@@
 file as its argument. When the file or dictionary is loaded, it is validated to
-have the correct format by Cerberus \autocite{cerberus}. The fields of the \ck{}
-file are stored as instance attributes of the \python|ChemKED| class.
+have the correct format by Cerberus \autocite{cerberus}.
+
+The fields of the \ck{} file are stored as instance attributes of the
+\python|ChemKED| class. The following attributes are available with the
+appropriate data:
+%
+\begin{itemize}
+    \item \python|chemked_version|, \python|file_version|, \python|file_author|,
+    \python|experiment_type|: Stores the values from the \ck{} file
+    \item \python|reference|: Instance of a \python|namedtuple| containing all
+    the information from the literature reference associated with the data
+    \item \python|apparatus|: Instance of a \python|namedtuple| containing all
+    the information about the apparatus used to perform the experiment
+    \item \python|datapoints|: A Python list of \python|DataPoint| instances
+\end{itemize}
+
+The \python|DataPoint| class stores the information associated with a single
+data point in the \ck{} file. Similar to the \python|ChemKED| class, the
+information from the \ck{} file are stored as attributes of class instances,
+including
+%
+\begin{itemize}
+    \item \python|equivalence_ratio|: The value of the equivalence ratio, if
+    present. For informational purposes only, no validation is done of the value
+    \item \python|ignition_delay|, \python|temperature|, \python|pressure|,
+    \python|pressure_rise|: The value of these data are stored as instances of
+    the \python|Quantity| class from the Pint \autocite{Grecco2016} package,
+    possibly with associated uncertainty
+    \item \python|composition|: A list of dictionaries of the species and their
+    respective amounts. The values are validated so that \yaml|mole-percent|,
+    \yaml|mole-fraction|, or \yaml|mass-fraction| cannot be mixed for a single
+    data point, and so that the sum of the values is approximately 1.0, or
+    100.0 for \yaml|mole-percent|.
+    \item \python|composition_type|: A string indicating the type of composition
+    information for the data point---one of \python|'mole-percent'|,
+    \python|'mole-fraction'|, or \python|'mass-fraction'|
+    \item \python|volume_history|: If the \yaml|volume-history| of an RCM
+    experiment is provided in the \ck{} file, it is stored in this attribute as
+    a \python|namedtuple|, and the actual values are stored in NumPy arrays
+    \autocite{vanderWalt:2011np}.
+\end{itemize}
+
+The \python|DataPoint| class has two instance methods defined, called
+\python|get_cantera_mole_fraction| and \python|get_cantera_mass_fraction|. With
+these methods, the composition of the reactant mixture can be output to a format
+that can be used to set the composition of a Cantera \python|Solution|. Note
+that the \python|composition| specification is not required to contain the
+molecular weights of the components, so calling
+\python|get_cantera_mass_fraction| when \yaml|mole-percent| or
+\yaml|mole-fraction| has been used in the \ck{} file will raise a
+\python|ValueError| (and vice versa for \python|get_cantera_mole_fraction|).
 

### 2017-02-25 pyked.tex
Add note about release
Describe get_dataframe

@@
 
+The \python|ChemKED| class has one instance method defined, called
+\python|get_dataframe|. Using this method, the user can get an instance of a
+Pandas~\autocite{pandas} \python|DataFrame| that contains the information in the
+list of \python|DataPoint|s. The user can specify the columns provided in the
+\python|DataFrame| by passing a list of column names to the
+\python|output_columns| argument of the \python|get_dataframe| function. The
+possible columns are not case-sensitive and are:
+
+\noindent\begin{itemize*}
+    \item \python|'Temperature'|
+    \item \python|'Pressure'|
+    \item \python|'Ignition Delay'|
+    \item \python|'Composition'|
+    \item \python|'Equivalence Ratio'|
+    \item \python|'Reference'|
+    \item \python|'Apparatus'|
+    \item \python|'Experiment Type'|
+    \item \python|'File Author'|
+    \item \python|'File Version'|
+    \item \python|'ChemKED Version'|
+\end{itemize*}
+
+In addition, specific fields from the Reference and Apparatus attributes can
+be included by specifying the name after a colon. These options are:
+
+\noindent\begin{itemize*}
+    \item \python|'Reference:Volume'|
+    \item \python|'Reference:Journal'|
+    \item \python|'Reference:DOI'|
+    \item \python|'Reference:Authors'|
+    \item \python|'Reference:Detail'|
+    \item \python|'Reference:Year'|
+    \item \python|'Reference:Pages'|
+    \item \python|'Apparatus:Kind'|
+\end{itemize*}
+
+% This weirdness is necessary to prevent the list from running over the margin
+\noindent\begin{itemize*}
+    \item \python|'Apparatus:Facility'|
+    \item \python|'Apparatus:Institution'|
+\end{itemize*}
+
+Only the first author is printed when \python|Reference| or
+\python|Reference:Authors| is selected because the whole author list may be
+quite long.
+
 \pk{} relies on well-established scientific Python software tools. These include
@@
 \todo[inline]{Need to archive version of PyKED and cite}
+\todo[inline, author=Bryan]{Should we do this even without officially releasing a version?}
 

### 2017-02-25 pyked.tex
Add Pandas reference

@@
 \pk{} relies on well-established scientific Python software tools. These include
-NumPy~\autocite{vanderWalt:2011np} for array manipulation,
-Pint~\autocite{Grecco2016} for interpreting and converting between units,
-\texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML files, Cerberus for
-validating \ck{} files~\autocite{cerberus}, and pytest~\autocite{pytest:3.0.1}
-for unit testing. Travis-CI~\autocite{Travis2016} provides continuous
-integration testing.
+NumPy~\autocite{vanderWalt:2011np} and Pandas~\autocite{pandas,McKinney2010} for
+array manipulation, Pint~\autocite{Grecco2016} for interpreting and converting
+between units, \texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML
+files, Cerberus for validating \ck{} files~\autocite{cerberus}, and
+pytest~\autocite{pytest:3.0.1} for unit testing. Travis-CI~\autocite{Travis2016}
+provides continuous integration testing.
 

### 2017-02-25 pyked.tex
Rework installation instructions

@@
 
-\pk{} is available under the permissive, open-source BSD 3-clause license via a GitHub
-repository: \url{https://github.com/pr-omethe-us/PyKED}. It can be installed using setuptools
-by downloading the source code files and executing \texttt{python setup.py install}. More mature
-versions of \pk{} will be distributed on \href{https://pypi.python.org/pypi}{PyPI}
-(Python Package Index) and conda-forge.\todo{add URL}{}
+\pk{} is licensed under the permissive, open-source BSD 3-clause license. The
+source code is publicly available on GitHub at
+\url{https://github.com/pr-omethe-us/PyKED}. Installation of stable versions can
+be accomplished from a package repository, using either \python|pip| to install
+from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
+\python|conda| to install from the
+\href{https://anaconda.org/conda-forge}{\texttt{conda-forge}} channel. In
+addition, \pk{} can be installed from the source code by running \python|python
+setup.py install| or \python|pip install .| from the root directory of the
+repository.
 
 %====================================================================
-\section{Use example}
+\section{Usage example}\label{sec:usage-example}
 %

### 2017-02-25 pyked.tex
added FlameMaster reference

@@
 %
-\todo[inline]{Kyle to do}
-
-Fundamental combustion experiments provide
+Fundamental combustion experiments provide vital data for understanding
+fuel chemistry and validating chemical kinetic models. Important measured quantities
+include autoignition delay times, laminar flame speeds, and species profiles.
+However, despite both the importance and abundance of such information in the literature,
+the community lacks a widely adopted standard format for this data. This impedes
+both sharing and wide use by the community.
 
@@
 CHEMKIN~\autocite{Kee:1996ck} format dominates. Competing standards such as the
-Cantera~\autocite{Cantera:2.3.0} CTI format or FlameMaster lag behind considerably, although
-use of the former continually grows due to Cantera's open availability.\todo{add flamemaster citation}{}
-Similarly,
+Cantera~\autocite{Cantera:2.3.0} CTI format or FlameMaster~\autocite{FlameMaster:ref,FlameMaster:code}
+lag behind considerably, although use of the former continually grows due to Cantera's
+open availability. Similarly,
 

### 2017-02-25 pyked.tex
Add first usage example
The Python bit is untested, probably should be tested before publication

--- 2017-02-25 pyked.tex
@@
 
+The following usage examples provide a guide to the use of \pk{}. They are by no
+means an exhaustive treatement, and are meant to demonstrate the basic
+capabilities of the software.
+
+\subsection{RCM modeling with varying reactor volume}\label{sec:rcm-modeling}
+
+The \ck{} file that will be used in this example can be found in the
+\python|tests| directory of the PyKED repository,
+\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
+Examining that file, we find the first section specifies the information about
+the \ck{} file itself
+%
+\begin{minted}{yaml}
+file-author:
+  name: Kyle E Niemeyer
+  ORCID: 0000-0003-4425-7097
+file-version: (1, 0)
+chemked-version: 0.0.1
+fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
+\end{minted}
+%
+Then, we find the information regarding the article in the literature from which
+this data was taken. In this case, the data point is from the work of \textcite{Mittal2006a}:
+%
+\begin{minted}{yaml}
+reference:
+  doi: 10.1002/kin.20180
+  authors:
+    - name: Gaurav Mittal
+    - name: Chih-Jen Sung
+      ORCID: 0000-0003-2046-8076
+    - name: Richard A Yetter
+  journal: International Journal of Chemical Kinetics
+  year: 2006
+  volume: 38
+  pages: 516-529
+  detail: Fig. 6, open circle
+experiment-type: ignition delay
+apparatus:
+  kind: rapid compression machine
+  institution: Case Western Reserve University
+  facility: CWRU RCM
+  fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
+\end{minted}
+
+Finally, there is just a single datapoint in this file, which codes the ignition
+delay for the experiment, the mixture composition, the initial temperature,
+pressure,
+%
+\begin{minted}{yaml}
+datapoints:
+  - temperature: 297.4 kelvin
+    ignition-delay: 1.0 ms
+    pressure: 958.0 torr
+    composition:
+      - species-name: H2
+        InChI: 1S/H2/h1H
+        mole-fraction: 0.12500
+      - species-name: O2
+        InChI: 1S/O2/c1-2
+        mole-fraction: 0.06250
+      - species-name: N2
+        InChI: 1S/N2/c1-2
+        mole-fraction: 0.18125
+      - species-name: Ar
+        InChI: 1S/Ar
+        mole-fraction: 0.63125
+    ignition-type:
+      target: pressure
+      type: d/dt max
+    compression-time: 38.0 ms
+    volume-history:
+      time:
+        units: s
+        column: 0
+      volume:
+        units: cm3
+        column: 1
+      values:
+        - [0.00E+000, 5.47669375000E+002]
+        - [1.00E-003, 5.46608789894E+002]
+        - [2.00E-003, 5.43427034574E+002]
+        - [3.00E-003, 5.38124109043E+002]
+        - 0 # Dammit Atom
+        ...
+\end{minted}
+
+The values for the \yaml|volume-history| are truncated to save space. The user
+can load this file into \pk{} and run a simulation in Cantera with the following
+code:
+
+\begin{minted}{python}
+import cantera as ct
+from pyked import ChemKED
+
+ck = ChemKED('testfile_rcm.yaml')
+dp = ck.datapoints[0]
+
+T_0 = dp.temperature.to('K').magnitude
+P_0 = dp.pressure.to('Pa').magnitude
+X_0 = dp.get_cantera_mole_fraction()
+
+gas = ct.Solution('h2-co-mechanism.cti')
+gas.TPX = T_0, P_0, X_0
+
+reac = ct.IdealGasReactor(gas)
+env = ct.Reservoir(ct.Solution('air.xml'))
+
+time = dp.volume_history.time
+volume = dp.volume_history.volume
+ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))
+
+netw = ct.ReactorNet([reac])
+while netw.time < 0.05:
+    netw.step()
+
+\end{minted}
 %====================================================================

### 2017-02-25 pyked.tex
Add second example
Again, haven't checked the Python works

@@
 \end{minted}
+
+\subsection{Shock tube modeling with constant volume}\label{sec:shock-tube}
+
+The \ck{} file used in this example can be found online.\todo{add url}
+Examining the file, we first find the section that specifies the information about the \ck{} file itself:
+%
+\begin{minted}{yaml}
+file-author:
+  name: Morgan Mayer
+  ORCID: 0000-0001-7137-5721
+file-version: (0,1)
+chemked-version: 0.0.1
+screw-you-atom: 0
+\end{minted}
+%
+Then the reference information:
+%
+\begin{minted}{yaml}
+reference:
+  doi: 10.1016/j.combustflame.2011.08.014
+  authors:
+    - name: Ivo Stranic
+    - name: Deanna P. Chase
+    - name: Joseph T. Harmon
+    - name: Sheng Yang
+    - name: David F. Davidson
+    - name: Ronald K. Hanson
+  journal: Combustion and Flame
+  year: 2011
+  volume: 159
+  pages: 516-527
+  detail:
+experiment-type: ignition delay
+apparatus:
+  kind: shock tube
+  institution: High Temperature Gasdynamics Laboratory, Stanford University
+  facility: stainless steel shock tube
+dammit-atom: 0
+\end{minted}
+%
+In this \ck{} file, multiple data points are specified with some common
+conditions, including a common mixture composition and common defintion of
+ignition delay. Therefore, a \yaml|common-properties| section is specified.
+%
+\begin{minted}{yaml}
+common-properties:
+  composition: &comp
+    - species-name: t-butanol
+      InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
+      mole-fraction: 0.003333333
+    - species-name: O2
+      InChI:  1S/O2/c1-2
+      mole-fraction: 0.04
+    - species-name: Ar
+      InChI:  1S/Ar
+      mole-fraction: 0.956666667
+  ignition-type:  &ign
+    target: OH* concentration
+    type: 1/2 max
+atom-is-bad-at-syntax-highlighting: 0
+\end{minted}
+%
+This block utilizes the ability of YAML files to define an anchor using the
+\yaml|&| symbol and refer to that section later using the \yaml|*| symbol, as
+can be seen in the definition of the \yaml|datapoints| section:
+%
+\begin{minted}{yaml}
+datapoints:
+  - temperature: 1459 kelvin
+    ignition-delay: 347 us
+    pressure: 1.60 atm
+    composition: *comp
+    ignition-type: *ign
+    equivalence-ratio: 0.5
+  - temperature: 1389 kelvin
+    ignition-delay: 756 us
+    pressure: 1.67 atm
+    composition: *comp
+    ignition-type: *ign
+    equivalence-ratio: 0.5
+  - temperature: 1497 kelvin
+    ignition-delay: 212 us
+    pressure: 1.55 atm
+    composition: *comp
+    ignition-type: *ign
+    equivalence-ratio: 0.5
+  - temperature: 1562 kelvin
+    ignition-delay: 105 us
+    pressure: 1.50 atm
+    composition: *comp
+    ignition-type: *ign
+    equivalence-ratio: 0.5
+\end{minted}
+
+In this example, we would like to run constant volume simulations at each of the
+pressure and temperature conditions in the \yaml|datapoints| list.
+
+\begin{minted}{python}
+import cantera as ct
+from pyked import ChemKED
+
+ck = ChemKED('Stranic2012-tbuoh.yaml')
+
+gas = ct.Solution('tbuoh-mechanism.cti')
+
+X_0 = ck.datapoints[0].get_cantera_mole_fraction()
+
+for dp in ck.datapoints:
+    T_0 = dp.temperature.to('K').magnitude
+    P_0 = dp.pressure.to('Pa').magnitude
+    gas.TPX = T_0, P_0, X_0
+    reac = ct.IdealGasReactor(gas)
+    netw = ct.ReactorNet([reac])
+    while reac.T < T_0 + 400:
+        netw.step()
+
+    print('The ignition delay for T_0={}, P_0={} is: {} seconds'.format(
+        T_0, P_0, netw.time
+    ))
+\end{minted}
 %====================================================================

### 2017-02-25 pyked.tex
Added example file reference

@@
 
-The \ck{} file used in this example can be found online.\todo{add url}
+The \ck{} file used in this example can be found online.\todo{add url}{}
+The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
+shock-tube ignition delays for \textit{tert}-butanol.
 Examining the file, we first find the section that specifies the information about the \ck{} file itself:

### 2017-02-26 pyked.tex
draft of intro/related work section

@@
 However, despite both the importance and abundance of such information in the literature,
-the community lacks a widely adopted standard format for this data. This impedes
-both sharing and wide use by the community.
+the combustion\slash chemical kinetics community lacks an accepted, commonly used
+standard for recording and sharing data from fundamental combustion experiments.
 
-%====================================================================
-\section{Related work}
-%
-\todo[inline]{Kyle to do}
-
-The combustion\slash chemical kinetics community lacks an accepted, commonly used
-standard for recording and sharing data from fundamental combustion experiments.
 Interestingly, this contrasts the situation for chemical kinetic models, where the
@@
 lag behind considerably, although use of the former continually grows due to Cantera's
-open availability. Similarly,
+open availability.
+% Similarly, % with more space, discuss JANAF/NASA polynomial standard ?
 
-Instead, such data must typically be found in under-documented, nonstandard comma-separated value
-files and Excel spreadsheets, or even contained in PDF tables or figures rather than as
-supplementary material associated with a paper.
+Instead, such data must typically be found in non- or under-documented, nonstandard
+comma-separated value files and Excel spreadsheets, or contained in PDF tables
+rather than as supplementary material associated with a paper.
+(In the worst cases, data is only available in figures---and in some cases not at all.)
+\todo{call out specific examples of these...?}{}
+These practices limit the wide use of the valuable, extensive experimental data
+available in the literature. Frenklach~\autocite{Frenklach:2007bm} further explained the
+benefits of standardized, widely available data for combustion.
 
-While the community has not widely adopted a standard format,
----most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
+While the community has not widely adopted a standard format or the practice of openly
+sharing data, some researchers and groups attempted to solve the first problem.
+Most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
 \textbf{M}od\textbf{e}l) data format~\autocite{Frenklach:2007bm,You:2011hy} and an associated
-online database, the PrIMe Data Warehouse~\autocite{PrIMe}.
+online database, the PrIMe Data Warehouse~\autocite{PrIMe}. PrIMe files encode fundamental
+combustion experimental data, along with kinetic models and calculated quantities,
+using the Extensible Markup Language (XML) standard.
 
-More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the \texttt{ReSpecTh} standard,
-an XML-based format that builds on, and is compatible with, PrIMe.
+PrIMe represented an admirable first attempt at standardization, but a number of limitations
+prevented wide adoption: the PrIMe standard does not require or support all information
+needed to simulate an experiment, including a machine-readable definition of ignition or a
+standard way to express the rapid compression machine (RCM) compression process; and
+the XML format, intended more as a markup language than data format, presents a barrier
+to creating and working with files.
+
+More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
+an XML-based format that builds on PrIMe. While ReSpecTh adds important features, it still
+suffers from the same usability issues as PrIMe.
 
@@
 experiments, \ck{}, and offer tools for easily working with data encoded in this format.
-Our motivations are similar to some of those of the PrIMe and \texttt{ReSpecTh} teams, although
+Niemeyer recently introduced an initial version of the \ck{} format~\autocite{Niemeyer:2016wf},
+and in this work we further formalize and develop the standard. We also discuss a related
+Python software tool for validating and working with \ck{} files: \pk{}.
+Our motivations are similar to some of those of the PrIMe and ReSpecTh teams, although
 less ambitious in scope: enable easy sharing and use of data, for the (primary) purposes of
@@
 by developing more user-friendly data files and tools for interacting with them.
-
-Niemeyer recently introduced an initial version of the \ck{} format~\autocite{Niemeyer:2016wf},
-and in this work we further formalize and develop the standard. We also discuss a related
-Python software tool for validating and working with \ck{} files: \pk{}.
+Furthermore, we plan to share all data and software openly, and cultivate a community of
+user-developers.
 

### 2017-02-26 pyked.tex
draft future work
a bit more in conclusions

@@
 \begin{keyword}
-    chemical kinetics\sep software\sep database
+    chemical kinetics\sep data standard\sep software\sep database
 \end{keyword}
@@
 
+In this article, we introduced the \ck{} data format for describing measurements
+taken from fundamental combustion experiments, as well as the Python-based tool
+\pk{} for working with the format. Currently, \ck{} supports ignition delay data
+from studies using shock tubes and rapid compression machines.
+
+We are currently incorporating support for specifying uncertainty in all quantities.
+Version 1.0 of \ck{}\slash \pk{} will include ...
+Later versions will support measurements from additional fundamental experiments,
+including laminar flames and flow reactors.
+
 \printbibliography

### 2017-02-26 pyked.tex
Bryan updates/notes in intro

@@
 open availability.
+\todo[inline, author=Bryan]{This reads more as a footnote, maybe consider putting it there instead?}
 % Similarly, % with more space, discuss JANAF/NASA polynomial standard ?
 
-Instead, such data must typically be found in non- or under-documented, nonstandard
+Instead, such data is typically found in under-documented
 comma-separated value files and Excel spreadsheets, or contained in PDF tables
 rather than as supplementary material associated with a paper.
-(In the worst cases, data is only available in figures---and in some cases not at all.)
-\todo{call out specific examples of these...?}{}
+In the worst case, data is only available in figures and must be digitized by the use of software such as WebPlotDigitizer.
+\todo{give a citation here}
+% I think we're already stirring the pot enough, no need to call out
+% specific people
+% \todo{call out specific examples of these...?}{}
 These practices limit the wide use of the valuable, extensive experimental data
@@
 benefits of standardized, widely available data for combustion.
+\todo[author=Bryan,inline]{It would maybe be good to give a bit of Frenklach's perspective here, rather than just throwing in the citation}
 
-While the community has not widely adopted a standard format or the practice of openly
+While the community has not widely adopted a standard format or the practice of openly \todo{this second bit feels like a slightly separate point, and it makes the second clause here a bit awkward}
 sharing data, some researchers and groups attempted to solve the first problem.
@@
 
-PrIMe represented an admirable first attempt at standardization, but a number of limitations
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represented an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
 prevented wide adoption: the PrIMe standard does not require or support all information
@@
 to creating and working with files.
+\todo{Also because the standard was closed?}
 
@@
 suffers from the same usability issues as PrIMe.
+\todo{slightly more detail?}
 
-We aim to develop a new, human- and machine-readable data standard for fundamental combustion
+In this work, we present a new, open-source, human- and machine-readable data standard for fundamental combustion
 experiments, \ck{}, and offer tools for easily working with data encoded in this format.
@@
 Python software tool for validating and working with \ck{} files: \pk{}.
-Our motivations are similar to some of those of the PrIMe and ReSpecTh teams, although
-less ambitious in scope: enable easy sharing and use of data, for the (primary) purposes of
+Our motivations are similar those of the PrIMe and ReSpecTh teams, although
+less ambitious in scope\todo{its not clear to me how this is less ambitious than the other projects}: the \ck{} project will enable easy sharing and use of fundamental combustion data, for the (primary) purposes of
 developing and validating predictive chemical kinetic models.
-However, those efforts were generally less successful in being adopted---we hope to succeed
-by developing more user-friendly data files and tools for interacting with them.
 Furthermore, we plan to share all data and software openly, and cultivate a community of
 user-developers.
+
+\todo[inline, author=Bryan]{We should probably also mention the CloudFlame project from KAUST.}
 

### 2017-02-26 pyked.tex
Update examples

@@
 Examining that file, we find the first section specifies the information about
-the \ck{} file itself
+the \ck{} file itself:
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 file-author:
@@
 chemked-version: 0.0.1
-fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
-\end{minted}
+\end{yamlbox}
 %
 Then, we find the information regarding the article in the literature from which
-this data was taken. In this case, the data point is from the work of \textcite{Mittal2006a}:
+this data was taken. In this case, the data point is from the work of
+\textcite{Mittal2006a}:
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 reference:
@@
   facility: CWRU RCM
-  fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
-\end{minted}
-
+\end{yamlbox}
+%
 Finally, there is just a single datapoint in this file, which codes the ignition
 delay for the experiment, the mixture composition, the initial temperature,
-pressure,
+pressure, compression time, ignition type, and the volume history that specifies
+how the volume of the reactor varies with time, simulating the compression
+stroke and post-compression processes.
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 datapoints:
@@
         - [3.00E-003, 5.38124109043E+002]
-        - 0 # Dammit Atom
         ...
-\end{minted}
-
-The values for the \yaml|volume-history| are truncated to save space. The user
-can load this file into \pk{} and run a simulation in Cantera with the following
-code:
-
-\begin{minted}{python}
+\end{yamlbox}
+%
+The values for the \yaml|volume-history| are truncated to save space. A user
+might like to load the information from this file to perform a simulation to
+calculate the ignition delay using Cantera. All of the information required to
+perform this simulation is present in the \ck{} file, with the exception of a
+chemical kinetic mechanism for H$_2$/CO combustion.
+%
+\begin{pythonbox}
 import cantera as ct
@@
 
+# Load the ChemKED file and retrieve the first element of the
+# datapoints list, which is an instance of the DataPoint class
 ck = ChemKED('testfile_rcm.yaml')
 dp = ck.datapoints[0]
-
+\end{pythonbox}
+%
+The initial temperature, pressure, and mixture composition can be read from the
+instance of the \python|DataPoint| class. \pk{} uses Pint \python|Quantities| to
+store values with units, while Cantera expects a floating point value in SI
+units as input. Therefore, we use the built-in capabilities of Pint to convert
+the units from those specified in the \ck{} file to SI units:
+%
+\begin{pythonbox}
 T_0 = dp.temperature.to('K').magnitude
@@
 
+# Load the mechanism and set the initial state of the mixture
 gas = ct.Solution('h2-co-mechanism.cti')
@@
 
+# Create the reactor and the outside environment
 reac = ct.IdealGasReactor(gas)
 env = ct.Reservoir(ct.Solution('air.xml'))
-
+\end{pythonbox}
+%
+To apply the effect of the volume trace to the \python|IdealGasReactor|, a
+\python|Wall| is installed between the reactor and the environment and assigned
+a velocity. Although we do not show the details here, a reference implementation
+of a class that computes a wall velocity given the volume history of the reactor
+is available in CanSen~\autocite{cansen}, in the
+\python|cansen.profile.VolumeProfile| class.
+%
+\begin{pythonbox}
 time = dp.volume_history.time
@@
 netw = ct.ReactorNet([reac])
+
+# Integrate for 50 ms
 while netw.time < 0.05:
     netw.step()
-
-\end{minted}
+\end{pythonbox}
+%
+Although not shown in this example, the user would typically store information
+about the \python|IdealGasReactor| (e.g., temperature, pressure, mass fractions)
+at each time step of the integration for post-processing.
 
@@
 
-The \ck{} file used in this example can be found online.\todo{add url}{}
-The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
-shock-tube ignition delays for \textit{tert}-butanol.
-Examining the file, we first find the section that specifies the information about the \ck{} file itself:
-%
-\begin{minted}{yaml}
-file-author:
-  name: Morgan Mayer
-  ORCID: 0000-0001-7137-5721
-file-version: (0,1)
-chemked-version: 0.0.1
-screw-you-atom: 0
-\end{minted}
-%
-Then the reference information:
-%
-\begin{minted}{yaml}
-reference:
-  doi: 10.1016/j.combustflame.2011.08.014
-  authors:
-    - name: Ivo Stranic
-    - name: Deanna P. Chase
-    - name: Joseph T. Harmon
-    - name: Sheng Yang
-    - name: David F. Davidson
-    - name: Ronald K. Hanson
-  journal: Combustion and Flame
-  year: 2011
-  volume: 159
-  pages: 516-527
-  detail:
-experiment-type: ignition delay
-apparatus:
-  kind: shock tube
-  institution: High Temperature Gasdynamics Laboratory, Stanford University
-  facility: stainless steel shock tube
-dammit-atom: 0
-\end{minted}
+The \ck{} file used in this example can be found online.\todo{add url}{} The
+data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
+shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
+meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
 %
 In this \ck{} file, multiple data points are specified with some common
-conditions, including a common mixture composition and common defintion of
+conditions, including a common mixture composition and common definition of
 ignition delay. Therefore, a \yaml|common-properties| section is specified.
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 common-properties:
@@
     type: 1/2 max
-atom-is-bad-at-syntax-highlighting: 0
-\end{minted}
+\end{yamlbox}
 %
@@
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 datapoints:
@@
     equivalence-ratio: 0.5
-\end{minted}
+\end{yamlbox}
 
 In this example, we would like to run constant volume simulations at each of the
-pressure and temperature conditions in the \yaml|datapoints| list.
-
-\begin{minted}{python}
+pressure and temperature conditions in the \yaml|datapoints| list. Once again,
+the \ck{} file specifies all the information required for the simulations except
+for the chemical kinetic mechanism.
+%
+\begin{pythonbox}
 import cantera as ct
@@
 
+# Since the composition is specified by the common-properties,
+# just take the composition of the first DataPoint
 X_0 = ck.datapoints[0].get_cantera_mole_fraction()
 
+# Loop through each of the DataPoints
 for dp in ck.datapoints:
@@
     netw = ct.ReactorNet([reac])
+    # Define ignition delay as T_0 + 400 K
     while reac.T < T_0 + 400:
@@
     ))
-\end{minted}
+\end{pythonbox}
+%
+Note that the ignition delay in this example is defined as equal to $T_0 +
+400\,\mathrm{K}$, a simplified definition used for this example. In general, the
+user could post-process the concentration information from the simulation to
+determine one-half the maximum OH* concentration to match the experimental
+definition of ignition delay.
 %====================================================================

### 2017-02-26 pyked.tex
Updates to boxes and CK format section

@@
 %====================================================================
-\section{Overview of format}\label{sec:overview-of-format}
+\section{Overview of ChemKED format}\label{sec:overview-of-format}
 
-The \ck{} file is a YAML-format file. This format offers the advantages of being
-human readable, written in plain text, and having parsers in most common
-programming languages, including Python, C++, Java, Perl, MATLAB, and many
-others. The YAML syntax is quite simple. The basic file structure is made of
-mappings, delimited by a colon. The key for the mapping is on the left of the
-colon, and the value on the right can be a single value, a sequence of values,
-a nested mapping, or some combination of these:
+This section will provide a brief introduction to the \ck{} database format.
+This is not a complete listing of available fields and more complete examples of
+\ck{} files can be found in the \pybox{tests} directory of the \pk{} package. In
+addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
+files. Note that in the following, text with gray background is a \yabox{yaml}
+keyword or value, while text with a yellow background is \pybox{Python} code.
 
-\begin{minted}{yaml}
+\ck{} files use the YAML data serialization format. This format offers the
+advantages of being human readable, written in plain text, and having parsers in
+most common programming languages, including Python, C++, Java, Perl, MATLAB,
+and many others. The YAML syntax is quite simple. The basic file structure is
+made of mappings, delimited by a colon. The key for the mapping is on the left
+of the colon, and the value on the right can be a single value, a sequence of
+values, a nested mapping, or some combination of these:
+%
+\begin{yamlbox}
 key1: value  # Single-value mapping
-    key2: [0, 1, 2]  # Mapping with sequence
-    key3:  # Alternative sequence format
+key2:  # Sequence format
   - value
   - value
-    key4:  # Nested mapping
-      key5: 0
-\end{minted}
+key3:  # Nested mapping
+  key4: 0
+key5:  # Sequence of mappings
+  - key-s1: 0
+    key-s2: value
+  - key-s2: value
+    key-s1: 0
+\end{yamlbox}
 
-The \yaml|value| can be a string, integer, or floating point number. The type is
-inferred from the \yaml|key| when the value is read. The \ck{} format is
-designed to include all of the information necessary to simulate a given
-experiment and can be broken into two main sections. The first section details
-all of the ``meta'' information about the experiment and the \ck{} file. In this
-section, the common \yaml|author| mapping is used for any fields where an author
-must be specified. The \yaml|author| mapping contains the following fields:
+The \yabox{value} can be a string, integer, or floating point number. The \ck{}
+format is designed to include all of the information necessary to simulate a
+given experiment and can be broken into two main sections. The first section
+details all of the ``meta'' information about the experiment and the \ck{} file.
+In this section, the common \yabox{author} mapping is used for any fields where
+an author must be specified. The \yabox{author} mapping contains the following
+fields:
 %
 \begin{itemize}
-    \item \yaml|name| (required, string): The author's full name
-    \item \yaml|ORCID| (optional, string): the author's ORCID number
+    \item \yabox{name} (required, string): The author's full name
+    \item \yabox{ORCID} (optional, string): the author's ORCID number
 \end{itemize}
@@
 \begin{itemize}
-    \item \yaml|file-author| (required, \yaml|author|-type mapping): The author
-    of the \ck{} file
-    \item \yaml|file-version| (required, integer): The version of the \ck{} file
-    \item \yaml|chemked-version| (required, string): The version of the \ck{}
+    \item \yabox{file-author} (required, \yabox{author}-type mapping): The
+    author of the \ck{} file
+    \item \yabox{file-version} (required, integer): The version of the \ck{}
+    file
+    \item \yabox{chemked-version} (required, string): The version of the \ck{}
     schema that this file conforms to
-    \item \yaml|reference| (required, mapping): The reference information for
+    \item \yabox{reference} (required, mapping): The reference information for
     the article associated with the data in the file. The fields in this
@@
     \begin{itemize}
-        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
+        \item \yabox{authors} (required, sequence): A sequence of \yabox{author}
         mappings
-        \item \yaml|journal| (required, string): The name of the publishing
+        \item \yabox{journal} (required, string): The name of the publishing
         journal
-        \item \yaml|year| (required, integer): The year of publication
+        \item \yabox{year} (required, integer): The year of publication
     \end{itemize}
-    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
-    delay| or \yaml|flame speed|
-    \item \yaml|apparatus| (required, string): The type of apparatus used to
+    \item \yabox{experiment-type} (required, string): Currently, only
+    \yabox{ignition delay} is supported
+    \item \yabox{apparatus} (required, string): The type of apparatus used to
     perform the experiment
@@
 
-In addition to the meta information, the data encoded in the \ck{} file is in
-a mapping called \yaml|datapoints|. The \yaml|datapoints| mapping contains a
-sequence of nested mappings, with the following information for each element of
-the sequence:
+In the \ck{} file, the experimental data is encoded in a sequence of mappings,
+with the following information required in each element of the sequence:
 %
 \begin{itemize}
-    \item \yaml|temperature| (required, sequence): The temperature of the
+    \item \yabox{temperature} (required, sequence): The temperature of the
     experiment, with units
-    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the
-    experiment, with units, if the \yaml|experiment-type| is \yaml|ignition
-    delay|
-    \item \yaml|flame-speed| (optional, sequence): The flame speed of the
-    experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
-    \item \yaml|pressure| (required, string): The pressure of the experiment,
+    \item \yabox{ignition-delay} (required, sequence): The ignition delay of the
+    experiment, with units
+    \item \yabox{pressure} (required, string): The pressure of the experiment,
     with units
-    \item \yaml|composition| (required, sequence): The composition of the
+    \item \yabox{composition} (required, sequence): The composition of the
     mixture in the experiment, a sequence of mappings with the following fields
     \begin{itemize}
-        \item \yaml|species-name| (required, string): The name of the species
-        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent|
+        \item \yabox{species-name} (required, string): The name of the species
+        \item \yabox{mole-fraction}, \yabox{mass-fraction}, \yabox{mole-percent}
         (required, float): The mole fraction or percent, or the mass fraction.
-        Only one of the three is permitted in a given \yaml|composition|
+        Only one of the three is permitted in a given \yabox{composition}
         sequence.
-        \item \yaml|InChI|, \yaml|SMILES|, \yaml|atomic-composition|, or
-        \yaml|elemental-composition| (required, string or sequence): The InChI
+        \item \yabox{InChI}, \yabox{SMILES}, \yabox{atomic-composition}, or
+        \yabox{elemental-composition} (required, string or sequence): The InChI
         or SMILES representing the molecule, or its atomic or elemental
-        composition (\yaml|atomic-composition| and \yaml|elemental-composition|
-        are synonyms) as a sequence
+        composition (\yabox{atomic-composition} and
+        \yabox{elemental-composition} are synonyms) as a sequence
     \end{itemize}
-    \item \yaml|ignition-type| (required, mapping); The type of ignition
+    \item \yabox{ignition-type} (required, mapping); The type of ignition
     measured in the experiments. The required fields are
     \begin{itemize}
-        \item \yaml|type| (required, string): How the ignition delay was
-        measured, one of \yaml|d/dt max| to indicate the ignition point is
-        found at the maximum of the time derivative of the \yaml|target|, or
-        \yaml|max| or \yaml|min| to indicate the ignition point is at the
-        maximum or minimum of the \yaml|target|
-        \item \yaml|target| (required, string): The target for the type of the
-        ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|,
-        \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
+        \item \yabox{type} (required, string): How the ignition delay was
+        measured, one of \yabox{d/dt max} to indicate the ignition point is
+        found at the maximum of the time derivative of the \yabox{target},
+        \yabox{max} or \yabox{min} to indicate the ignition point is at the
+        maximum or minimum of the \yabox{target}, or \yabox{1/2 max} to
+        indicate the half-maximum point
+        \item \yabox{target} (required, string): The target for the type of the
+        ignition point, one of \yabox{temperature}, \yabox{pressure},
+        \yabox{OH}, \yabox{OH*}, \yabox{CH}, or \yabox{CH*}
     \end{itemize}
@@
 
-Each of the quantities in a datapoint must be specified with units. This should
-be done as a single string associated with the first element of the sequence.
-The units of the quantity are validated to ensure they have the appropriate
-dimensionality for the quanity. In addition, each of the quantities in the
-\yaml|datapoint| can optionally be assigned an uncertainty. This uncertainty can
-be either absolute or relative, and is specified as elements of the sequence of
-the associated key. For example, the absolute uncertainty of the temperature and
-the relative uncertainty of the ignition delay might be specified as follows:
+The top-level key for the sequence is called \yabox{datapoints}. Each of the
+quantities in an element must be specified with units. This should be done as a
+single string associated with the first element of the sequence. The units of
+the quantity are validated to ensure they have the appropriate dimensionality
+for the quantity. In addition, each of the quantities in an element can
+optionally be assigned an uncertainty. This uncertainty can be either absolute
+or relative, and is specified as elements of the sequence of the associated key.
+For example, the absolute uncertainty of the temperature and the relative
+uncertainty of the ignition delay might be specified as follows:
 
-\begin{minted}{yaml}
+\begin{yamlbox}
 datapoints:
@@
     ...
-\end{minted}
+\end{yamlbox}
 
 %====================================================================
-\section{PyKED Architecture}
+\section{PyKED architecture}\label{sec:pyked-architecture}
 %
@@
 of the interface to \ck{} files. \pk{} reads \ck{} files, validates their
-structure and content, and allows the user to interact with the datapoints
-contained in the \ck{} file.
+structure and content, and allows the user to interact with the data contained
+in the \ck{} file.
 
 The basic user interface to the \ck{} file in \pk{} is through the
-\python|ChemKED| class. The \python|ChemKED| class constructor takes the file
+\pybox{ChemKED} class. The \pybox{ChemKED} class constructor takes the file
 name of a \ck{} file or a Python dictionary containing the contents of a \ck{}
 file as its argument. When the file or dictionary is loaded, it is validated to
-have the correct format by Cerberus \autocite{cerberus}.
+have the correct format by the Python package Cerberus \autocite{cerberus}.
 
 The fields of the \ck{} file are stored as instance attributes of the
-\python|ChemKED| class. The following attributes are available with the
-appropriate data:
+\pybox{ChemKED} class. The following attributes are available:
 %
 \begin{itemize}
-    \item \python|chemked_version|, \python|file_version|, \python|file_author|,
-    \python|experiment_type|: Stores the values from the \ck{} file
-    \item \python|reference|: Instance of a \python|namedtuple| containing all
+    \item \pybox{chemked_version}, \pybox{file_version}, \pybox{file_author},
+    \pybox{experiment_type}: Stores the values from the \ck{} file
+    \item \pybox{reference}: Instance of a \pybox{namedtuple} containing all
     the information from the literature reference associated with the data
-    \item \python|apparatus|: Instance of a \python|namedtuple| containing all
+    \item \pybox{apparatus}: Instance of a \pybox{namedtuple} containing all
     the information about the apparatus used to perform the experiment
-    \item \python|datapoints|: A Python list of \python|DataPoint| instances
+    \item \pybox{datapoints}: A Python list of \pybox{DataPoint} instances
 \end{itemize}
 
-The \python|DataPoint| class stores the information associated with a single
-data point in the \ck{} file. Similar to the \python|ChemKED| class, the
-information from the \ck{} file are stored as attributes of class instances,
-including
+The \pybox{DataPoint} class stores the information associated with a single data
+point in the \ck{} file (i.e., a single element of the \yabox{datapoints}
+sequence). Similar to the \pybox{ChemKED} class, information in the
+\pybox{DataPoint} is stored as instance attributes:
 %
 \begin{itemize}
-    \item \python|equivalence_ratio|: The value of the equivalence ratio, if
+    \item \pybox{equivalence_ratio}: The value of the equivalence ratio, if
     present. For informational purposes only, no validation is done of the value
-    \item \python|ignition_delay|, \python|temperature|, \python|pressure|,
-    \python|pressure_rise|: The value of these data are stored as instances of
-    the \python|Quantity| class from the Pint \autocite{Grecco2016} package,
+    \item \pybox{ignition_delay}, \pybox{temperature}, \pybox{pressure},
+    \pybox{pressure_rise}: The value of these data are stored as instances of
+    the \pybox{Quantity} class from the Pint \autocite{Grecco2016} package,
     possibly with associated uncertainty
-    \item \python|composition|: A list of dictionaries of the species and their
-    respective amounts. The values are validated so that \yaml|mole-percent|,
-    \yaml|mole-fraction|, or \yaml|mass-fraction| cannot be mixed for a single
+    \item \pybox{composition}: A list of dictionaries of the species and their
+    respective amounts. The values are validated so that \yabox{mole-percent},
+    \yabox{mole-fraction}, or \yabox{mass-fraction} cannot be mixed for a single
     data point, and so that the sum of the values is approximately 1.0, or
-    100.0 for \yaml|mole-percent|.
-    \item \python|composition_type|: A string indicating the type of composition
-    information for the data point---one of \python|'mole-percent'|,
-    \python|'mole-fraction'|, or \python|'mass-fraction'|
-    \item \python|volume_history|: If the \yaml|volume-history| of an RCM
+    100.0 for \yabox{mole-percent}.
+    \item \pybox{composition_type}: A string indicating the type of composition
+    information for the data point---one of \pybox{'mole-percent'},
+    \pybox{'mole-fraction'}, or \pybox{'mass-fraction'}
+    \item \pybox{volume_history}: If the \yabox{volume-history} of an RCM
     experiment is provided in the \ck{} file, it is stored in this attribute as
-    a \python|namedtuple|, and the actual values are stored in NumPy arrays
+    a \pybox{namedtuple}, and the actual values are stored in NumPy arrays
     \autocite{vanderWalt:2011np}.
+    \item \pybox{compression_time}: If the \ck{} file encodes an RCM experiment,
+    the length of the compression stroke may be included
 \end{itemize}
 
-The \python|DataPoint| class has two instance methods defined, called
-\python|get_cantera_mole_fraction| and \python|get_cantera_mass_fraction|. With
-these methods, the composition of the reactant mixture can be output to a format
-that can be used to set the composition of a Cantera \python|Solution|. Note
-that the \python|composition| specification is not required to contain the
-molecular weights of the components, so calling
-\python|get_cantera_mass_fraction| when \yaml|mole-percent| or
-\yaml|mole-fraction| has been used in the \ck{} file will raise a
-\python|ValueError| (and vice versa for \python|get_cantera_mole_fraction|).
+The \pybox{DataPoint} class has two instance methods defined, called
+\pybox{get_cantera_mole_fraction()} and \pybox{get_cantera_mass_fraction()}.
+With these methods, the composition of the reactant mixture can be output to a
+format that can be used to set the composition of a Cantera \pybox{Solution}.
+Note that the \pybox{composition} specification does not contain the molecular
+weights of the components, so conversion between mole fractions and mass
+fractions is not possible.
 
-The \python|ChemKED| class has one instance method defined, called
-\python|get_dataframe|. Using this method, the user can get an instance of a
-Pandas~\autocite{pandas} \python|DataFrame| that contains the information in the
-list of \python|DataPoint|s. The user can specify the columns provided in the
-\python|DataFrame| by passing a list of column names to the
-\python|output_columns| argument of the \python|get_dataframe| function. The
+The \pybox{ChemKED} class has one instance method defined, called
+\pybox{get_dataframe()}. Using this method, the user can get an instance of a
+Pandas~\autocite{pandas} \pybox{DataFrame} that contains the information in the
+list of \pybox{DataPoint}s. The user can specify the columns included in the
+\pybox{DataFrame} by passing a list of column names to the
+\pybox{output_columns} argument of the \pybox{get_dataframe()} method. The
 possible columns are not case-sensitive and are:
@@
 \noindent\begin{itemize*}
-    \item \python|'Temperature'|
-    \item \python|'Pressure'|
-    \item \python|'Ignition Delay'|
-    \item \python|'Composition'|
-    \item \python|'Equivalence Ratio'|
-    \item \python|'Reference'|
-    \item \python|'Apparatus'|
-    \item \python|'Experiment Type'|
-    \item \python|'File Author'|
-    \item \python|'File Version'|
-    \item \python|'ChemKED Version'|
+    \item \pybox{'Temperature'}
+    \item \pybox{'Pressure'}
+    \item \pybox{'Ignition Delay'}
+    \item \pybox{'Composition'}
+    \item \pybox{'Equivalence Ratio'}
+    \item \pybox{'Reference'}
+    \item \pybox{'Apparatus'}
+    \item \pybox{'Experiment Type'}
+    \item \pybox{'File Author'}
+    \item \pybox{'File Version'}
+    \item \pybox{'ChemKED Version'}
 \end{itemize*}
 
-In addition, specific fields from the Reference and Apparatus attributes can
-be included by specifying the name after a colon. These options are:
+In addition, specific fields from the \pybox{Reference} and \pybox{Apparatus}
+attributes can be included by specifying the name after a colon. These options
+are:
 
 \noindent\begin{itemize*}
-    \item \python|'Reference:Volume'|
-    \item \python|'Reference:Journal'|
-    \item \python|'Reference:DOI'|
-    \item \python|'Reference:Authors'|
-    \item \python|'Reference:Detail'|
-    \item \python|'Reference:Year'|
-    \item \python|'Reference:Pages'|
-    \item \python|'Apparatus:Kind'|
+    \item \pybox{'Reference:Volume'}
+    \item \pybox{'Reference:Journal'}
+    \item \pybox{'Reference:DOI'}
+    \item \pybox{'Reference:Authors'}
+    \item \pybox{'Reference:Detail'}
+    \item \pybox{'Reference:Year'}
+    \item \pybox{'Reference:Pages'}
+    \item \pybox{'Apparatus:Kind'}
 \end{itemize*}
@@
 \noindent\begin{itemize*}
-    \item \python|'Apparatus:Facility'|
-    \item \python|'Apparatus:Institution'|
+    \item \pybox{'Apparatus:Facility'}
+    \item \pybox{'Apparatus:Institution'}
 \end{itemize*}
 
-Only the first author is printed when \python|Reference| or
-\python|Reference:Authors| is selected because the whole author list may be
+Only the first author is printed when \pybox{Reference} or
+\pybox{Reference:Authors} is selected because the whole author list may be
 quite long.
@@
 
-\todo[inline]{Need to archive version of PyKED and cite}
-\todo[inline, author=Bryan]{Should we do this even without officially releasing a version?}
-
 \pk{} is licensed under the permissive, open-source BSD 3-clause license. The
@@
 \url{https://github.com/pr-omethe-us/PyKED}. Installation of stable versions can
-be accomplished from a package repository, using either \python|pip| to install
+be accomplished from a package repository, using either \pybox{pip} to install
 from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
-\python|conda| to install from the
+\pybox{conda} to install from the
 \href{https://anaconda.org/conda-forge}{\texttt{conda-forge}} channel. In
-addition, \pk{} can be installed from the source code by running \python|python
-setup.py install| or \python|pip install .| from the root directory of the
-repository.
+addition, \pk{} can be installed from the source code by running
+\pybox|python setup.py install| or \pybox{pip install .} from the root
+directory of the repository.
 
@@
 \section{Usage example}\label{sec:usage-example}
-%
-\todo[inline]{Bryan/Kyle to do}
-\todo[inline]{Idea: CONV simulations using Cantera in a loop with conditions
-from the Pandas output for butanol?}
-\todo[inline]{Idea: VPRO simulations using Cantera from a \ck{} file for an RCM
-datapoint}
 
 The following usage examples provide a guide to the use of \pk{}. They are by no
-means an exhaustive treatement, and are meant to demonstrate the basic
+means an exhaustive treatment, and are meant to demonstrate the basic
 capabilities of the software.
@@
 The \ck{} file that will be used in this example can be found in the
-\python|tests| directory of the PyKED repository,
+\pybox{tests} directory of the PyKED repository,
 \url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
@@
 %
-The values for the \yaml|volume-history| are truncated to save space. A user
+The values for the \yabox{volume-history} are truncated to save space. A user
 might like to load the information from this file to perform a simulation to
@@
 The initial temperature, pressure, and mixture composition can be read from the
-instance of the \python|DataPoint| class. \pk{} uses Pint \python|Quantities| to
+instance of the \pybox{DataPoint} class. \pk{} uses Pint \pybox{Quantities} to
 store values with units, while Cantera expects a floating point value in SI
@@
 %
-To apply the effect of the volume trace to the \python|IdealGasReactor|, a
-\python|Wall| is installed between the reactor and the environment and assigned
+To apply the effect of the volume trace to the \pybox{IdealGasReactor}, a
+\pybox{Wall} is installed between the reactor and the environment and assigned
 a velocity. Although we do not show the details here, a reference implementation
@@
 is available in CanSen~\autocite{cansen}, in the
-\python|cansen.profile.VolumeProfile| class.
+\pybox{cansen.profile.VolumeProfile} class.
 %
@@
 Although not shown in this example, the user would typically store information
-about the \python|IdealGasReactor| (e.g., temperature, pressure, mass fractions)
+about the \pybox{IdealGasReactor} (e.g., temperature, pressure, mass fractions)
 at each time step of the integration for post-processing.
@@
 conditions, including a common mixture composition and common definition of
-ignition delay. Therefore, a \yaml|common-properties| section is specified.
+ignition delay. Therefore, a \yabox{common-properties} section is specified.
 %
@@
 This block utilizes the ability of YAML files to define an anchor using the
-\yaml|&| symbol and refer to that section later using the \yaml|*| symbol, as
-can be seen in the definition of the \yaml|datapoints| section:
+\yabox{&} symbol and refer to that section later using the \yabox{*} symbol, as
+can be seen in the definition of the \yabox{datapoints} section:
 %
@@
 In this example, we would like to run constant volume simulations at each of the
-pressure and temperature conditions in the \yaml|datapoints| list. Once again,
+pressure and temperature conditions in the \yabox{datapoints} list. Once again,
 the \ck{} file specifies all the information required for the simulations except

### 2017-02-26 pyked.tex
Edit conclusion

@@
 \section{Conclusions and future work}
-%
-\todo[inline]{both to do}
 
-In this article, we introduced the \ck{} data format for describing measurements
-taken from fundamental combustion experiments, as well as the Python-based tool
-\pk{} for working with the format. Currently, \ck{} supports ignition delay data
-from studies using shock tubes and rapid compression machines.
+Recognizing that the community has a need for an open and standardized data
+serialization format, in this paper we have presented the \ck{} data format for
+describing measurements taken from fundamental combustion experiments. \ck{}
+files are formatted using the YAML language and are plain-text, human- and
+machine-readable, and easy to construct.
 
-We are currently incorporating support for specifying uncertainty in all quantities.
-Version 1.0 of \ck{}\slash \pk{} will include ...
-Later versions will support measurements from additional fundamental experiments,
-including laminar flames and flow reactors.
+We have also presented a Python-based tool, \pk{}, for validating and working
+with \ck{} files. \pk{} provides the reference implementation of the validator
+for \ck{} files and utilizes several common packages from Python's scientific
+computing community. \pk{} and \ck{} currently support ignition delay
+measurements from rapid compression machines and shock tubes, including facility
+specific effects from each type of experiment.
+
+\ck{} and \pk{} are currently under active development, and contributions from
+the community are gratefully welcomed. All development occurs under the BSD
+3-clause license and the code is housed on GitHub at
+\url{https://github.com/pr-omethe-us/PyKED}. Future directions for development
+are outlined in the public
+\href{Roadmap}{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}. The highest
+priority issues are currently adding support for other types of fundamental
+experiments, including flame speed, speciation, and flame extinction
+measurements.
 

### 2017-02-26 pyked.tex
edits and additions to intro

--- 2017-02-26 pyked.tex
@@
 the combustion\slash chemical kinetics community lacks an accepted, commonly used
-standard for recording and sharing data from fundamental combustion experiments.
-
+standard for recording and sharing data from fundamental combustion experiments.\footnote{
 Interestingly, this contrasts the situation for chemical kinetic models, where the
@@
 open availability.
-\todo[inline, author=Bryan]{This reads more as a footnote, maybe consider putting it there instead?}
 % Similarly, % with more space, discuss JANAF/NASA polynomial standard ?
+}
 
@@
 rather than as supplementary material associated with a paper.
-In the worst case, data is only available in figures and must be digitized by the use of software such as WebPlotDigitizer.
+In the worst case, data is only available in figures and must be digitized by the
+use of software such as WebPlotDigitizer.
 \todo{give a citation here}
@@
 benefits of standardized, widely available data for combustion.
-\todo[author=Bryan,inline]{It would maybe be good to give a bit of Frenklach's perspective here, rather than just throwing in the citation}
+In brief, much of combustion\slash chemical kinetics research drives towards the ultimate goal
+of developing predictive kinetic models---and ultimately this depends on a community
+infrastructure for data and methods.
 
-While the community has not widely adopted a standard format or the practice of openly \todo{this second bit feels like a slightly separate point, and it makes the second clause here a bit awkward}
-sharing data, some researchers and groups attempted to solve the first problem.
+The combustion community has not yet widely adopted a standard format for experimental data,
+but some researchers and groups attempted to solve this problem.\todo[author=Kyle]{this still doesn't read quite right}{}
 Most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
@@
 
-PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represented an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
-prevented wide adoption: the PrIMe standard does not require or support all information
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represents an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
+prevent wide adoption: the PrIMe standard does not require or support all information
 needed to simulate an experiment, including a machine-readable definition of ignition or a
@@
 to creating and working with files.
-\todo{Also because the standard was closed?}
+Furthermore, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
+limit contributions and community development of these and compatible tools.
 
 More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
-an XML-based format that builds on PrIMe. While ReSpecTh adds important features, it still
-suffers from the same usability issues as PrIMe.
-\todo{slightly more detail?}
+an XML-based format that builds on PrIMe. ReSpecTh adds important features that make files
+better standalone representations of experimental data---i.e., more informative by themselves
+rather than in concert with a larger system.
+For example, where PrIMe used internal bibliographic references, ReSpecTh added a field for
+typical bibliographic data. The ReSpecTh standard also provides machine-readable formats
+for better specifying ignition experiments, including a definition of autoignition and
+either volume histories for RCMs or pressure-rise in shock tubes.
+However, being XML-based, ReSpecTh experimental files still suffer from the same usability
+issues as PrIMe, and the standard remains closed.
+\todo[author=Kyle,inline]{how's this?}
+
+Olm and colleagues used their ReSpecTh-based database to quantify the performance of literature
+hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models, and more recently
+develop improved models for hydrogen\slash syngas~\autocite{Varga:2016gj} and
+ethanol~\autocite{Olm:2016et}. In all four cases, they converted numerous experimental datasets
+from the literature into the ReSpecTh standard.
+
+Building on the lack of an accepted data standard, limited examples exist of open sharing
+of such data. The Clean Combustion Research Center at King Abdullah University of Science
+and Technology (KAUST) started the CloudFlame project in 2013~\autocite{Goteng:2013cf}\todo{a few more refs needed}{}
+(\url{https://cloudflame.kaust.edu.sa}). CloudFlame does serve as an openly accessible database
+for experimental data, available in a standard CSV format, but also goes beyond this by
+providing a cloud infrastructure for running simulations based on stored models and data.
+While an admirable (and useful) effort, CloudFlame represents another closed system controlled
+by a specific institution rather than the community.
 
@@
 Python software tool for validating and working with \ck{} files: \pk{}.
-Our motivations are similar those of the PrIMe and ReSpecTh teams, although
-less ambitious in scope\todo{its not clear to me how this is less ambitious than the other projects}: the \ck{} project will enable easy sharing and use of fundamental combustion data, for the (primary) purposes of
-developing and validating predictive chemical kinetic models.
-Furthermore, we plan to share all data and software openly, and cultivate a community of
-user-developers.
-
-\todo[inline, author=Bryan]{We should probably also mention the CloudFlame project from KAUST.}
+Our motivations are similar those of the PrIMe and ReSpecTh teams:
+the \ck{} project will enable easy sharing and use of fundamental combustion data,
+for the (primary) purposes of developing and validating predictive chemical kinetic models.
+However, we make usability a major design focus, and plan to share all data and software openly
+to cultivate a community of user-contributors.
 
@@
 are:
+\todo[author=Kyle,inline]{one difference ReSpecTh called out from PrIMe is that
+the latter added a copyright line in all files, claiming copyright on the data...
+do we want to be incendiary and point out that you can't copyright data?}
 %
@@
 Examining that file, we find the first section specifies the information about
-the \ck{} file itself:
+the \ck{} file itself
 %
-\begin{yamlbox}
+\begin{minted}{yaml}
 file-author:
@@
 chemked-version: 0.0.1
-\end{yamlbox}
+fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
+\end{minted}
 %
 Then, we find the information regarding the article in the literature from which
-this data was taken. In this case, the data point is from the work of
-\textcite{Mittal2006a}:
+this data was taken. In this case, the data point is from the work of \textcite{Mittal2006a}:
 %
-\begin{yamlbox}
+\begin{minted}{yaml}
 reference:
@@
   facility: CWRU RCM
-\end{yamlbox}
-%
+  fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
+\end{minted}
+
 Finally, there is just a single datapoint in this file, which codes the ignition
 delay for the experiment, the mixture composition, the initial temperature,
-pressure, compression time, ignition type, and the volume history that specifies
-how the volume of the reactor varies with time, simulating the compression
-stroke and post-compression processes.
+pressure,
 %
-\begin{yamlbox}
+\begin{minted}{yaml}
 datapoints:
@@
         - [3.00E-003, 5.38124109043E+002]
+        - 0 # Dammit Atom
         ...
-\end{yamlbox}
-%
-The values for the \yaml|volume-history| are truncated to save space. A user
-might like to load the information from this file to perform a simulation to
-calculate the ignition delay using Cantera. All of the information required to
-perform this simulation is present in the \ck{} file, with the exception of a
-chemical kinetic mechanism for H$_2$/CO combustion.
-%
-\begin{pythonbox}
+\end{minted}
+
+The values for the \yaml|volume-history| are truncated to save space. The user
+can load this file into \pk{} and run a simulation in Cantera with the following
+code:
+
+\begin{minted}{python}
 import cantera as ct
@@
 
-# Load the ChemKED file and retrieve the first element of the
-# datapoints list, which is an instance of the DataPoint class
 ck = ChemKED('testfile_rcm.yaml')
 dp = ck.datapoints[0]
-\end{pythonbox}
-%
-The initial temperature, pressure, and mixture composition can be read from the
-instance of the \python|DataPoint| class. \pk{} uses Pint \python|Quantities| to
-store values with units, while Cantera expects a floating point value in SI
-units as input. Therefore, we use the built-in capabilities of Pint to convert
-the units from those specified in the \ck{} file to SI units:
-%
-\begin{pythonbox}
+
 T_0 = dp.temperature.to('K').magnitude
@@
 
-# Load the mechanism and set the initial state of the mixture
 gas = ct.Solution('h2-co-mechanism.cti')
@@
 
-# Create the reactor and the outside environment
 reac = ct.IdealGasReactor(gas)
 env = ct.Reservoir(ct.Solution('air.xml'))
-\end{pythonbox}
-%
-To apply the effect of the volume trace to the \python|IdealGasReactor|, a
-\python|Wall| is installed between the reactor and the environment and assigned
-a velocity. Although we do not show the details here, a reference implementation
-of a class that computes a wall velocity given the volume history of the reactor
-is available in CanSen~\autocite{cansen}, in the
-\python|cansen.profile.VolumeProfile| class.
-%
-\begin{pythonbox}
+
 time = dp.volume_history.time
@@
 netw = ct.ReactorNet([reac])
-
-# Integrate for 50 ms
 while netw.time < 0.05:
     netw.step()
-\end{pythonbox}
-%
-Although not shown in this example, the user would typically store information
-about the \python|IdealGasReactor| (e.g., temperature, pressure, mass fractions)
-at each time step of the integration for post-processing.
+
+\end{minted}
 
@@
 
-The \ck{} file used in this example can be found online.\todo{add url}{} The
-data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
-shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
-meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
+The \ck{} file used in this example can be found online.\todo{add url}{}
+The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
+shock-tube ignition delays for \textit{tert}-butanol.
+Examining the file, we first find the section that specifies the information about the \ck{} file itself:
+%
+\begin{minted}{yaml}
+file-author:
+  name: Morgan Mayer
+  ORCID: 0000-0001-7137-5721
+file-version: (0,1)
+chemked-version: 0.0.1
+screw-you-atom: 0
+\end{minted}
+%
+Then the reference information:
+%
+\begin{minted}{yaml}
+reference:
+  doi: 10.1016/j.combustflame.2011.08.014
+  authors:
+    - name: Ivo Stranic
+    - name: Deanna P. Chase
+    - name: Joseph T. Harmon
+    - name: Sheng Yang
+    - name: David F. Davidson
+    - name: Ronald K. Hanson
+  journal: Combustion and Flame
+  year: 2011
+  volume: 159
+  pages: 516-527
+  detail:
+experiment-type: ignition delay
+apparatus:
+  kind: shock tube
+  institution: High Temperature Gasdynamics Laboratory, Stanford University
+  facility: stainless steel shock tube
+dammit-atom: 0
+\end{minted}
 %
 In this \ck{} file, multiple data points are specified with some common
-conditions, including a common mixture composition and common definition of
+conditions, including a common mixture composition and common defintion of
 ignition delay. Therefore, a \yaml|common-properties| section is specified.
 %
-\begin{yamlbox}
+\begin{minted}{yaml}
 common-properties:
@@
     type: 1/2 max
-\end{yamlbox}
+atom-is-bad-at-syntax-highlighting: 0
+\end{minted}
 %
@@
 %
-\begin{yamlbox}
+\begin{minted}{yaml}
 datapoints:
@@
     equivalence-ratio: 0.5
-\end{yamlbox}
+\end{minted}
 
 In this example, we would like to run constant volume simulations at each of the
-pressure and temperature conditions in the \yaml|datapoints| list. Once again,
-the \ck{} file specifies all the information required for the simulations except
-for the chemical kinetic mechanism.
-%
-\begin{pythonbox}
+pressure and temperature conditions in the \yaml|datapoints| list.
+
+\begin{minted}{python}
 import cantera as ct
@@
 
-# Since the composition is specified by the common-properties,
-# just take the composition of the first DataPoint
 X_0 = ck.datapoints[0].get_cantera_mole_fraction()
 
-# Loop through each of the DataPoints
 for dp in ck.datapoints:
@@
     netw = ct.ReactorNet([reac])
-    # Define ignition delay as T_0 + 400 K
     while reac.T < T_0 + 400:
@@
     ))
-\end{pythonbox}
-%
-Note that the ignition delay in this example is defined as equal to $T_0 +
-400\,\mathrm{K}$, a simplified definition used for this example. In general, the
-user could post-process the concentration information from the simulation to
-determine one-half the maximum OH* concentration to match the experimental
-definition of ignition delay.
+\end{minted}
 %====================================================================

### 2017-02-26 pyked.tex
fixed some \yaml|ignition delay| formatting
If there is a line break inside the ||, extra space is added in the
compiled document.

@@
     \end{itemize}
-    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
-    delay| or \yaml|flame speed|
+    \item \yaml|experiment-type| (required, string): One of
+    \yaml|ignition delay| or \yaml|flame speed|
     \item \yaml|apparatus| (required, string): The type of apparatus used to
@@
     \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the
-    experiment, with units, if the \yaml|experiment-type| is \yaml|ignition
-    delay|
+    experiment, with units, if the \yaml|experiment-type| is
+    \yaml|ignition delay|
     \item \yaml|flame-speed| (optional, sequence): The flame speed of the
@@
 %====================================================================
-\section{PyKED Architecture}
+\section{PyKED architecture}
 %

### 2017-02-26 pyked.tex
edited discussion of PrIMe flaws a bit

@@
 
-PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represents an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
-prevent wide adoption: the PrIMe standard does not require or support all information
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make it
+a strong standardization format, but suffers from two primary flaws that
+prevent wide adoption: (1) the PrIMe standard does not require or support all information
 needed to simulate an experiment, including a machine-readable definition of ignition or a
-standard way to express the rapid compression machine (RCM) compression process; and
+standard way to express the rapid compression machine (RCM) compression process; (2) and
 the XML format, intended more as a markup language than data format, presents a barrier
 to creating and working with files.
-Furthermore, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
+Furthermore, the focus on using internal identifiers for bibliographic references and
+species\slash reactions, rather than standard identifiers (e.g., DOIs for scholarly products
+and InCHI or SMILES for species), limits usability outside the PrIMe ecosystem.
+Finally, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
 limit contributions and community development of these and compatible tools.
@@
 issues as PrIMe, and the standard remains closed.
-\todo[author=Kyle,inline]{how's this?}
-
 Olm and colleagues used their ReSpecTh-based database to quantify the performance of literature
@@
 from the literature into the ReSpecTh standard.
+\todo[author=Kyle,inline]{how's this?}
 

### 2017-02-26 pyked.tex

@@
     \begin{itemize}
-        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
+        \item \yabox{authors} (required, sequence): A sequence of \yabox{author}
         mappings
-        \item \yaml|journal| (required, string): The name of the publishing
+        \item \yabox{journal} (required, string): The name of the publishing
         journal
-        \item \yaml|year| (required, integer): The year of publication
+        \item \yabox{year} (required, integer): The year of publication
     \end{itemize}
-    \item \yaml|experiment-type| (required, string): One of
-    \yaml|ignition delay| or \yaml|flame speed|
-    \item \yaml|apparatus| (required, string): The type of apparatus used to
+    \item \yabox{experiment-type} (required, string): Currently, only
+    \yabox{ignition delay} is supported
+    \item \yabox{apparatus} (required, string): The type of apparatus used to
     perform the experiment
@@
 
-In addition to the meta information, the data encoded in the \ck{} file is in
-a mapping called \yaml|datapoints|. The \yaml|datapoints| mapping contains a
-sequence of nested mappings, with the following information for each element of
-the sequence:
+In the \ck{} file, the experimental data is encoded in a sequence of mappings,
+with the following information required in each element of the sequence:
 %
 \begin{itemize}
-    \item \yaml|temperature| (required, sequence): The temperature of the
+    \item \yabox{temperature} (required, sequence): The temperature of the
     experiment, with units
-    \item \yaml|ignition-delay| (optional, sequence): The ignition delay of the
-    experiment, with units, if the \yaml|experiment-type| is
-    \yaml|ignition delay|
-    \item \yaml|flame-speed| (optional, sequence): The flame speed of the
-    experiment, with units, if the \yaml|experiment-type| is \yaml|flame speed|
-    \item \yaml|pressure| (required, string): The pressure of the experiment,
+    \item \yabox{ignition-delay} (required, sequence): The ignition delay of the
+    experiment, with units
+    \item \yabox{pressure} (required, string): The pressure of the experiment,
     with units
-    \item \yaml|composition| (required, sequence): The composition of the
+    \item \yabox{composition} (required, sequence): The composition of the
     mixture in the experiment, a sequence of mappings with the following fields
     \begin{itemize}
-        \item \yaml|species-name| (required, string): The name of the species
-        \item \yaml|mole-fraction|, \yaml|mass-fraction|, \yaml|mole-percent|
+        \item \yabox{species-name} (required, string): The name of the species
+        \item \yabox{mole-fraction}, \yabox{mass-fraction}, \yabox{mole-percent}
         (required, float): The mole fraction or percent, or the mass fraction.
-        Only one of the three is permitted in a given \yaml|composition|
+        Only one of the three is permitted in a given \yabox{composition}
         sequence.
-        \item \yaml|InChI|, \yaml|SMILES|, \yaml|atomic-composition|, or
-        \yaml|elemental-composition| (required, string or sequence): The InChI
+        \item \yabox{InChI}, \yabox{SMILES}, \yabox{atomic-composition}, or
+        \yabox{elemental-composition} (required, string or sequence): The InChI
         or SMILES representing the molecule, or its atomic or elemental
-        composition (\yaml|atomic-composition| and \yaml|elemental-composition|
-        are synonyms) as a sequence
+        composition (\yabox{atomic-composition} and
+        \yabox{elemental-composition} are synonyms) as a sequence
     \end{itemize}
-    \item \yaml|ignition-type| (required, mapping); The type of ignition
+    \item \yabox{ignition-type} (required, mapping); The type of ignition
     measured in the experiments. The required fields are
     \begin{itemize}
-        \item \yaml|type| (required, string): How the ignition delay was
-        measured, one of \yaml|d/dt max| to indicate the ignition point is
-        found at the maximum of the time derivative of the \yaml|target|, or
-        \yaml|max| or \yaml|min| to indicate the ignition point is at the
-        maximum or minimum of the \yaml|target|
-        \item \yaml|target| (required, string): The target for the type of the
-        ignition point, one of \yaml|temperature|, \yaml|pressure|, \yaml|OH|,
-        \yaml|OH*|, \yaml|CH|, or \yaml|CH*|
+        \item \yabox{type} (required, string): How the ignition delay was
+        measured, one of \yabox{d/dt max} to indicate the ignition point is
+        found at the maximum of the time derivative of the \yabox{target},
+        \yabox{max} or \yabox{min} to indicate the ignition point is at the
+        maximum or minimum of the \yabox{target}, or \yabox{1/2 max} to
+        indicate the half-maximum point
+        \item \yabox{target} (required, string): The target for the type of the
+        ignition point, one of \yabox{temperature}, \yabox{pressure},
+        \yabox{OH}, \yabox{OH*}, \yabox{CH}, or \yabox{CH*}
     \end{itemize}
@@
     ...
-\end{minted}
+\end{yamlbox}
 
 %====================================================================
-\section{PyKED architecture}
+\section{PyKED architecture}\label{sec:pyked-architecture}
 %
@@
 The \ck{} file that will be used in this example can be found in the
-\python|tests| directory of the PyKED repository,
+\pybox{tests} directory of the PyKED repository,
 \url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
 Examining that file, we find the first section specifies the information about
-the \ck{} file itself
+the \ck{} file itself:
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 file-author:
@@
         - [3.00E-003, 5.38124109043E+002]
-        - 0 # Dammit Atom
         ...
-\end{minted}
-
-The values for the \yaml|volume-history| are truncated to save space. The user
-can load this file into \pk{} and run a simulation in Cantera with the following
-code:
-
-\begin{minted}{python}
+\end{yamlbox}
+%
+The values for the \yabox{volume-history} are truncated to save space. A user
+might like to load the information from this file to perform a simulation to
+calculate the ignition delay using Cantera. All of the information required to
+perform this simulation is present in the \ck{} file, with the exception of a
+chemical kinetic mechanism for H$_2$/CO combustion.
+%
+\begin{pythonbox}
 import cantera as ct
@@
 
+# Load the ChemKED file and retrieve the first element of the
+# datapoints list, which is an instance of the DataPoint class
 ck = ChemKED('testfile_rcm.yaml')
 dp = ck.datapoints[0]
-
+\end{pythonbox}
+%
+The initial temperature, pressure, and mixture composition can be read from the
+instance of the \pybox{DataPoint} class. \pk{} uses Pint \pybox{Quantities} to
+store values with units, while Cantera expects a floating point value in SI
+units as input. Therefore, we use the built-in capabilities of Pint to convert
+the units from those specified in the \ck{} file to SI units:
+%
+\begin{pythonbox}
 T_0 = dp.temperature.to('K').magnitude
@@
 
+# Create the reactor and the outside environment
 reac = ct.IdealGasReactor(gas)
 env = ct.Reservoir(ct.Solution('air.xml'))
-
+\end{pythonbox}
+%
+To apply the effect of the volume trace to the \pybox{IdealGasReactor}, a
+\pybox{Wall} is installed between the reactor and the environment and assigned
+a velocity. Although we do not show the details here, a reference implementation
+of a class that computes a wall velocity given the volume history of the reactor
+is available in CanSen~\autocite{cansen}, in the
+\pybox{cansen.profile.VolumeProfile} class.
+%
+\begin{pythonbox}
 time = dp.volume_history.time
@@
 netw = ct.ReactorNet([reac])
+
+# Integrate for 50 ms
 while netw.time < 0.05:
     netw.step()
-
-\end{minted}
+\end{pythonbox}
+%
+Although not shown in this example, the user would typically store information
+about the \pybox{IdealGasReactor} (e.g., temperature, pressure, mass fractions)
+at each time step of the integration for post-processing.
 
@@
 
-The \ck{} file used in this example can be found online.\todo{add url}{}
-The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
-shock-tube ignition delays for \textit{tert}-butanol.
-Examining the file, we first find the section that specifies the information about the \ck{} file itself:
-%
-\begin{minted}{yaml}
-file-author:
-  name: Morgan Mayer
-  ORCID: 0000-0001-7137-5721
-file-version: (0,1)
-chemked-version: 0.0.1
-screw-you-atom: 0
-\end{minted}
-%
-Then the reference information:
-%
-\begin{minted}{yaml}
-reference:
-  doi: 10.1016/j.combustflame.2011.08.014
-  authors:
-    - name: Ivo Stranic
-    - name: Deanna P. Chase
-    - name: Joseph T. Harmon
-    - name: Sheng Yang
-    - name: David F. Davidson
-    - name: Ronald K. Hanson
-  journal: Combustion and Flame
-  year: 2011
-  volume: 159
-  pages: 516-527
-  detail:
-experiment-type: ignition delay
-apparatus:
-  kind: shock tube
-  institution: High Temperature Gasdynamics Laboratory, Stanford University
-  facility: stainless steel shock tube
-dammit-atom: 0
-\end{minted}
+The \ck{} file used in this example can be found online.\todo{add url}{} The
+data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
+shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
+meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
 %
 In this \ck{} file, multiple data points are specified with some common
-conditions, including a common mixture composition and common defintion of
-ignition delay. Therefore, a \yaml|common-properties| section is specified.
+conditions, including a common mixture composition and common definition of
+ignition delay. Therefore, a \yabox{common-properties} section is specified.
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 common-properties:
@@
     type: 1/2 max
-atom-is-bad-at-syntax-highlighting: 0
-\end{minted}
+\end{yamlbox}
 %
 This block utilizes the ability of YAML files to define an anchor using the
-\yaml|&| symbol and refer to that section later using the \yaml|*| symbol, as
-can be seen in the definition of the \yaml|datapoints| section:
+\yabox{&} symbol and refer to that section later using the \yabox{*} symbol, as
+can be seen in the definition of the \yabox{datapoints} section:
 %
-\begin{minted}{yaml}
+\begin{yamlbox}
 datapoints:
@@
     equivalence-ratio: 0.5
-\end{minted}
+\end{yamlbox}
 
 In this example, we would like to run constant volume simulations at each of the
-pressure and temperature conditions in the \yaml|datapoints| list.
-
-\begin{minted}{python}
+pressure and temperature conditions in the \yabox{datapoints} list. Once again,
+the \ck{} file specifies all the information required for the simulations except
+for the chemical kinetic mechanism.
+%
+\begin{pythonbox}
 import cantera as ct
@@
     ))
-\end{minted}
+\end{pythonbox}
+%
+Note that the ignition delay in this example is defined as equal to $T_0 +
+400\,\mathrm{K}$, a simplified definition used for this example. In general, the
+user could post-process the concentration information from the simulation to
+determine one-half the maximum OH* concentration to match the experimental
+definition of ignition delay.
 %====================================================================
 \section{Conclusions and future work}
-%
-\todo[inline]{both to do}
 
-In this article, we introduced the \ck{} data format for describing measurements
-taken from fundamental combustion experiments, as well as the Python-based tool
-\pk{} for working with the format. Currently, \ck{} supports ignition delay data
-from studies using shock tubes and rapid compression machines.
+Recognizing that the community has a need for an open and standardized data
+serialization format, in this paper we have presented the \ck{} data format for
+describing measurements taken from fundamental combustion experiments. \ck{}
+files are formatted using the YAML language and are plain-text, human- and
+machine-readable, and easy to construct.
 
-We are currently incorporating support for specifying uncertainty in all quantities.
-Version 1.0 of \ck{}\slash \pk{} will include ...
-Later versions will support measurements from additional fundamental experiments,
-including laminar flames and flow reactors.
+We have also presented a Python-based tool, \pk{}, for validating and working
+with \ck{} files. \pk{} provides the reference implementation of the validator
+for \ck{} files and utilizes several common packages from Python's scientific
+computing community. \pk{} and \ck{} currently support ignition delay
+measurements from rapid compression machines and shock tubes, including facility
+specific effects from each type of experiment.
+
+\ck{} and \pk{} are currently under active development, and contributions from
+the community are gratefully welcomed. All development occurs under the BSD
+3-clause license and the code is housed on GitHub at
+\url{https://github.com/pr-omethe-us/PyKED}. Future directions for development
+are outlined in the public
+\href{Roadmap}{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}. The highest
+priority issues are currently adding support for other types of fundamental
+experiments, including flame speed, speciation, and flame extinction
+measurements.
 

### 2017-02-26 pyked.tex
added citation for WebPlotDigitizer
some edits to conclusions

@@
 In the worst case, data is only available in figures and must be digitized by the
-use of software such as WebPlotDigitizer.
-\todo{give a citation here}
+use of software such as WebPlotDigitizer~\autocite{WebPlotDigitizer}.
 % I think we're already stirring the pot enough, no need to call out
@@
 
-Recognizing that the community has a need for an open and standardized data
-serialization format, in this paper we have presented the \ck{} data format for
-describing measurements taken from fundamental combustion experiments. \ck{}
-files are formatted using the YAML language and are plain-text, human- and
+In this paper we presented the \ck{} data format for describing measurements
+taken from fundamental combustion experiments, recognizing
+that the community has a need for an open and standardized data serialization format.
+\ck{} files are formatted using the YAML language and are plain-text, human- and
 machine-readable, and easy to construct.
 
-We have also presented a Python-based tool, \pk{}, for validating and working
+We also presented a Python-based tool, \pk{}, for validating and working
 with \ck{} files. \pk{} provides the reference implementation of the validator
@@
 computing community. \pk{} and \ck{} currently support ignition delay
-measurements from rapid compression machines and shock tubes, including facility
-specific effects from each type of experiment.
+measurements from rapid compression machines and shock tubes, including
+facility-specific effects from each type of experiment.
 

### 2017-02-27 pyked.tex
Update and reflow intro

@@
 %
-Fundamental combustion experiments provide vital data for understanding
-fuel chemistry and validating chemical kinetic models. Important measured quantities
-include autoignition delay times, laminar flame speeds, and species profiles.
-However, despite both the importance and abundance of such information in the literature,
-the combustion\slash chemical kinetics community lacks an accepted, commonly used
-standard for recording and sharing data from fundamental combustion experiments.\footnote{
-Interestingly, this contrasts the situation for chemical kinetic models, where the
-CHEMKIN~\autocite{Kee:1996ck} format dominates. Competing standards such as the
-Cantera~\autocite{Cantera:2.3.0} CTI format or FlameMaster~\autocite{FlameMaster:ref,FlameMaster:code}
-lag behind considerably, although use of the former continually grows due to Cantera's
-open availability.
+Fundamental combustion experiments provide vital data for understanding fuel
+chemistry and validating chemical kinetic models. Important measured quantities
+include autoignition delay times, laminar flame speeds, and species profiles,
+among others. However, despite both the importance and abundance of such
+information in the literature, the combustion\slash chemical kinetics community
+lacks an accepted, commonly used standard for recording and sharing data from
+fundamental combustion experiments.\footnote{ Interestingly, this contrasts the
+situation for chemical kinetic models, where the CHEMKIN~\autocite{Kee:1996ck}
+format dominates. Competing standards such as the
+Cantera~\autocite{Cantera:2.3.0} CTI format or
+FlameMaster~\autocite{FlameMaster:ref,FlameMaster:code} lag behind considerably,
+although use of the former continually grows due to Cantera's open availability.
 % Similarly, % with more space, discuss JANAF/NASA polynomial standard ?
@@
 
-Instead, such data is typically found in under-documented
-comma-separated value files and Excel spreadsheets, or contained in PDF tables
-rather than as supplementary material associated with a paper.
-In the worst case, data is only available in figures and must be digitized by the
-use of software such as WebPlotDigitizer~\autocite{WebPlotDigitizer}.
-% I think we're already stirring the pot enough, no need to call out
-% specific people
-% \todo{call out specific examples of these...?}{}
-These practices limit the wide use of the valuable, extensive experimental data
-available in the literature. Frenklach~\autocite{Frenklach:2007bm} further explained the
-benefits of standardized, widely available data for combustion.
-In brief, much of combustion\slash chemical kinetics research drives towards the ultimate goal
-of developing predictive kinetic models---and ultimately this depends on a community
-infrastructure for data and methods.
+Instead, such data is typically found in under-documented comma-separated value
+files and Excel spreadsheets, or contained in PDF tables rather than as
+supplementary material associated with a paper. In the worst case, data is only
+available in figures and must be digitized by the use of software such as
+WebPlotDigitizer~\autocite{WebPlotDigitizer}. These practices limit the wide use
+of the valuable, extensive experimental data available in the literature.
+Frenklach~\autocite{Frenklach:2007bm} further explained the benefits of
+standardized, widely available data for combustion. In brief, much of
+fundamental combustion\slash chemical kinetics research drives towards the
+ultimate goal of developing predictive kinetic models---and ultimately this
+depends on a \emph{community} infrastructure for data and methods.
 
-The combustion community has not yet widely adopted a standard format for experimental data,
-but some researchers and groups attempted to solve this problem.\todo[author=Kyle]{this still doesn't read quite right}{}
-Most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
-\textbf{M}od\textbf{e}l) data format~\autocite{Frenklach:2007bm,You:2011hy} and an associated
-online database, the PrIMe Data Warehouse~\autocite{PrIMe}. PrIMe files encode fundamental
-combustion experimental data, along with kinetic models and calculated quantities,
-using the Extensible Markup Language (XML) standard.
+The combustion community has not yet widely adopted a standard format for
+experimental data, but some researchers and groups have proposed solutions to
+this problem. Most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess
+\textbf{I}nformatics \textbf{M}od\textbf{e}l) data
+format~\autocite{Frenklach:2007bm,You:2011hy} and an associated online database,
+the PrIMe Data Warehouse~\autocite{PrIMe}. PrIMe files encode fundamental
+combustion experimental data, along with kinetic models and calculated
+quantities, using the Extensible Markup Language (XML) standard.
 
-PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make it
-a strong standardization format, but suffers from two primary flaws that
-prevent wide adoption: (1) the PrIMe standard does not require or support all information
-needed to simulate an experiment, including a machine-readable definition of ignition or a
-standard way to express the rapid compression machine (RCM) compression process; (2) and
-the XML format, intended more as a markup language than data format, presents a barrier
-to creating and working with files.
-Furthermore, the focus on using internal identifiers for bibliographic references and
-species\slash reactions, rather than standard identifiers (e.g., DOIs for scholarly products
-and InCHI or SMILES for species), limits usability outside the PrIMe ecosystem.
-Finally, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
-limit contributions and community development of these and compatible tools.
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make
+it a strong standardization format, but it suffers from several flaws that
+prevent wide adoption. First, the PrIMe standard does not require or support all
+information needed to simulate an experiment, including a machine-readable
+definition of ignition or a standard way to express detailed facility specific
+effects important to properly simulating certain experiments. Second, the XML
+format is intended to be a markup language rather than a data format, and
+presents a barrier to creating and working with database files. Furthermore, the
+focus on using internal identifiers for bibliographic references and
+species\slash reactions, rather than standard identifiers (e.g., DOIs for
+scholarly products and InCHI or SMILES for species), limits usability outside
+the PrIMe ecosystem. Finally, the closed and opaque nature of the PrIMe standard
+and associated Data Warehouse limit contributions and community development of
+the PrIMe standard and compatible data files and tools.
 
-More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
-an XML-based format that builds on PrIMe. ReSpecTh adds important features that make files
-better standalone representations of experimental data---i.e., more informative by themselves
-rather than in concert with a larger system.
-For example, where PrIMe used internal bibliographic references, ReSpecTh added a field for
-typical bibliographic data. The ReSpecTh standard also provides machine-readable formats
-for better specifying ignition experiments, including a definition of autoignition and
-either volume histories for RCMs or pressure-rise in shock tubes.
-However, being XML-based, ReSpecTh experimental files still suffer from the same usability
-issues as PrIMe, and the standard remains closed.
-Olm and colleagues used their ReSpecTh-based database to quantify the performance of literature
-hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models, and more recently
-develop improved models for hydrogen\slash syngas~\autocite{Varga:2016gj} and
-ethanol~\autocite{Olm:2016et}. In all four cases, they converted numerous experimental datasets
-from the literature into the ReSpecTh standard.
-\todo[author=Kyle,inline]{how's this?}
+More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the
+ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
+features that make files better standalone representations of experimental
+data---i.e., more informative by themselves rather than in concert with a larger
+system. For example, where PrIMe uses internal bibliographic references,
+ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
+provides machine-readable formats for specifying ignition experiments, including
+a field for the definition of autoignition and the ability to specify
+facility-specific effects. However, ReSpecTh experimental files are another
+XML-based format, and as such, suffer from the same usability issues as PrIMe.
+Moreover, the standard remains closed, requiring potential users to register
+with the standard's authors to access the specification.
 
-Building on the lack of an accepted data standard, limited examples exist of open sharing
-of such data. The Clean Combustion Research Center at King Abdullah University of Science
-and Technology (KAUST) started the CloudFlame project in 2013~\autocite{Goteng:2013cf}\todo{a few more refs needed}{}
-(\url{https://cloudflame.kaust.edu.sa}). CloudFlame does serve as an openly accessible database
-for experimental data, available in a standard CSV format, but also goes beyond this by
-providing a cloud infrastructure for running simulations based on stored models and data.
-While an admirable (and useful) effort, CloudFlame represents another closed system controlled
-by a specific institution rather than the community.
+Nonetheless, the usefulness of databases of fundamental combustion experiments
+has been shown to be quite useful. For instance, Olm and colleagues used their
+ReSpecTh-based database to quantify the performance of literature
+hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models,
+and more recently develop improved models for hydrogen\slash
+syngas~\autocite{Varga:2016gj} and ethanol~\autocite{Olm:2016et}. In all four
+cases, they converted numerous experimental datasets from the literature into
+the ReSpecTh standard.
 
-In this work, we present a new, open-source, human- and machine-readable data standard for fundamental combustion
-experiments, \ck{}, and offer tools for easily working with data encoded in this format.
-Niemeyer recently introduced an initial version of the \ck{} format~\autocite{Niemeyer:2016wf},
-and in this work we further formalize and develop the standard. We also discuss a related
-Python software tool for validating and working with \ck{} files: \pk{}.
-Our motivations are similar those of the PrIMe and ReSpecTh teams:
-the \ck{} project will enable easy sharing and use of fundamental combustion data,
-for the (primary) purposes of developing and validating predictive chemical kinetic models.
-However, we make usability a major design focus, and plan to share all data and software openly
-to cultivate a community of user-contributors.
+Despite the development of these standards, limited examples of open data
+\emph{sharing} utilizing the formats is available in the literature. To the best
+of our knowledge, the largest such such publicly available database is hosted by
+the Clean Combustion Research Center at King Abdullah University of Science and
+Technology (KAUST) and is known as CloudFlame
+(\url{https://cloudflame.kaust.edu.sa}). Started in
+2013~\autocite{Goteng:2013cf},\todo{a few more refs needed}{} CloudFlame serves
+as an openly accessible database for experimental data, available in a standard
+CSV format, and also provides a cloud infrastructure for running simulations
+based on stored models and data. While an admirable and useful effort,
+CloudFlame, like ReSpecTh and PrIMe, is a closed system controlled by a single
+institution rather than by the community at large.
+
+Therefore, we believe there is a need for an open, community focused and
+developed, combustion\slash chemical kinetics data format. In this work, we
+present a new, open-source, human- and machine-readable data standard for
+fundamental combustion experiments, \ck{}, and offer tools for easily working
+with data encoded in this format. Niemeyer recently introduced an initial
+version of the \ck{} format~\autocite{Niemeyer:2016wf}, and in this work we
+further formalize and develop the standard. We also discuss a related Python
+software tool for validating and working with \ck{} files: \pk{}. Our
+motivations are similar those of the PrIMe and ReSpecTh teams: the \ck{} project
+will enable easy sharing and use of fundamental combustion data, for the
+(primary) purposes of developing and validating predictive chemical kinetic
+models. However, we make usability a major design focus, and plan to share all
+data and software openly to cultivate a community of user-contributors.
 
@@
 \ck{} and \pk{} are currently under active development, and contributions from
-the community are gratefully welcomed. All development occurs under the BSD
+the community are welcomed. All development occurs under the BSD
 3-clause license and the code is housed on GitHub at

### 2017-02-27 pyked.tex
cleaned up a few things

@@
 lacks an accepted, commonly used standard for recording and sharing data from
-fundamental combustion experiments.\footnote{ Interestingly, this contrasts the
+fundamental combustion experiments.\footnote{Interestingly, this contrasts the
 situation for chemical kinetic models, where the CHEMKIN~\autocite{Kee:1996ck}
@@
 
-Instead, such data is typically found in under-documented comma-separated value
+Instead, such data is typically found in under-documented comma-separated value (CSV)
 files and Excel spreadsheets, or contained in PDF tables rather than as
@@
 Despite the development of these standards, limited examples of open data
-\emph{sharing} utilizing the formats is available in the literature. To the best
-of our knowledge, the largest such such publicly available database is hosted by
+\emph{sharing} using the formats exist in the literature. To the best
+of our knowledge, the largest such publicly available database is hosted by
 the Clean Combustion Research Center at King Abdullah University of Science and
@@
 
-Therefore, we believe there is a need for an open, community focused and
+Therefore, we believe there remains a need for an open, community-focused and
 developed, combustion\slash chemical kinetics data format. In this work, we
@@
 are:
-\todo[author=Kyle,inline]{one difference ReSpecTh called out from PrIMe is that
-the latter added a copyright line in all files, claiming copyright on the data...
-do we want to be incendiary and point out that you can't copyright data?}
 %
@@
 \url{https://github.com/pr-omethe-us/PyKED}. Future directions for development
-are outlined in the public
-\href{Roadmap}{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}. The highest
-priority issues are currently adding support for other types of fundamental
+are outlined in the public roadmap\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}.
+The highest priority issues are currently adding support for other types of fundamental
 experiments, including flame speed, speciation, and flame extinction

### 2017-02-27 pyked.tex
added additional CloudFlame refs

@@
 (\url{https://cloudflame.kaust.edu.sa}). Started in
-2013~\autocite{Goteng:2013cf},\todo{a few more refs needed}{} CloudFlame serves
+2013~\autocite{Goteng:2013cf,Goteng:2014,ReynoChiasson:2015}, CloudFlame serves
 as an openly accessible database for experimental data, available in a standard

### 2017-02-27 pyked.tex
added PyKED ref, and moved long URLs to footnotes

@@
 \pk{} is a Python package that currently provides the reference implementation
-of the interface to \ck{} files. \pk{} reads \ck{} files, validates their
-structure and content, and allows the user to interact with the data contained
+of the interface to \ck{} files~\autocite{PyKED}. \pk{} reads \ck{} files, validates
+their structure and content, and allows the user to interact with the data contained
 in the \ck{} file.
@@
 
-\pk{} relies on well-established scientific Python software tools. These include
-NumPy~\autocite{vanderWalt:2011np} and Pandas~\autocite{pandas,McKinney2010} for
-array manipulation, Pint~\autocite{Grecco2016} for interpreting and converting
+\pk{}~\autocite{PyKED} relies on well-established scientific Python software tools.
+These include NumPy~\autocite{vanderWalt:2011np} and Pandas~\autocite{pandas,McKinney2010}
+for array manipulation, Pint~\autocite{Grecco2016} for interpreting and converting
 between units, \texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML
@@
 source code is publicly available on GitHub at
-\url{https://github.com/pr-omethe-us/PyKED}. Installation of stable versions can
-be accomplished from a package repository, using either \pybox{pip} to install
+\url{https://github.com/pr-omethe-us/PyKED}~\autocite{PyKED}. Installation of stable versions
+can be accomplished from a package repository, using either \pybox{pip} to install
 from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
@@
 The \ck{} file that will be used in this example can be found in the
-\pybox{tests} directory of the PyKED repository,
-\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
+\pybox{tests} directory of the PyKED
+repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}.
 Examining that file, we find the first section specifies the information about
@@
 
-The \ck{} file used in this example can be found online.\todo{add url}{} The
-data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
+The \ck{} file used in this example can be found in the \pybox{tests} directory of the \pk{} 
+repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}.
+The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
 shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file

### 2017-02-27 pyked.tex
Last updates from me!

@@
 
-Instead, such data is typically found in under-documented comma-separated value (CSV)
-files and Excel spreadsheets, or contained in PDF tables rather than as
+Instead, such data is typically found in under-documented comma-separated value
+(CSV) files and Excel spreadsheets, or contained in PDF tables rather than as
 supplementary material associated with a paper. In the worst case, data is only
@@
 of the valuable, extensive experimental data available in the literature.
-Frenklach~\autocite{Frenklach:2007bm} further explained the benefits of
-standardized, widely available data for combustion. In brief, much of
-fundamental combustion\slash chemical kinetics research drives towards the
-ultimate goal of developing predictive kinetic models---and ultimately this
-depends on a \emph{community} infrastructure for data and methods.
+\textcite{Frenklach:2007bm} further explained the benefits of standardized,
+widely available data for combustion. In brief, much of fundamental
+combustion\slash chemical kinetics research drives towards the ultimate goal of
+developing predictive kinetic models---and ultimately this depends on a
+\emph{community} infrastructure for data and methods.
 
@@
 prevent wide adoption. First, the PrIMe standard does not require or support all
-information needed to simulate an experiment, including a machine-readable
+the information needed to simulate an experiment, including a machine-readable
 definition of ignition or a standard way to express detailed facility specific
@@
 
-In the \ck{} file, the experimental data is encoded in a sequence of mappings,
-with the following information required in each element of the sequence:
+The second section of the file encodes the experimental data. In the \ck{} file,
+this is done in a sequence of mappings, with the following information required
+in each element of the sequence:
 %
@@
 The basic user interface to the \ck{} file in \pk{} is through the
-\pybox{ChemKED} class. The \pybox{ChemKED} class constructor takes the file
-name of a \ck{} file or a Python dictionary containing the contents of a \ck{}
-file as its argument. When the file or dictionary is loaded, it is validated to
-have the correct format by the Python package Cerberus \autocite{cerberus}.
+\pybox{ChemKED} class. The \pybox{ChemKED} class constructor takes the file name
+of a \ck{} file or a Python dictionary containing the contents of a \ck{} file
+as its argument. When the file or dictionary is loaded, it is validated to have
+the correct format by the Python package Cerberus \autocite{cerberus}. The
+schema used for validation of the \ck{} files is available publicly in the
+source code repository for
+PyKED\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/chemked_schema.yaml}}.
 
@@
     \item \pybox{equivalence_ratio}: The value of the equivalence ratio, if
-    present. For informational purposes only, no validation is done of the value
+    present. For informational purposes only, no validation is done of the
+    value.
     \item \pybox{ignition_delay}, \pybox{temperature}, \pybox{pressure},
@@
     a \pybox{namedtuple}, and the actual values are stored in NumPy arrays
-    \autocite{vanderWalt:2011np}.
+    \autocite{vanderWalt:2011np}
     \item \pybox{compression_time}: If the \ck{} file encodes an RCM experiment,
     the length of the compression stroke may be included
+    \item \pybox{ignition_type}: A dictionary specifying the method of the
+    measurement of ignition delay
 \end{itemize}
@@
 
-In this paper we presented the \ck{} data format for describing measurements
-taken from fundamental combustion experiments, recognizing
-that the community has a need for an open and standardized data serialization format.
-\ck{} files are formatted using the YAML language and are plain-text, human- and
+In this paper, we presented the \ck{} data format for describing measurements
+taken from fundamental combustion experiments, recognizing that the community
+has a need for an open and standardized data serialization format. \ck{} files
+are formatted using the YAML language and are plain-text, human- and
 machine-readable, and easy to construct.
@@
 \ck{} and \pk{} are currently under active development, and contributions from
-the community are welcomed. All development occurs under the BSD
-3-clause license and the code is housed on GitHub at
+the community are welcomed. All development occurs under the BSD 3-clause
+license and the code is housed on GitHub at
 \url{https://github.com/pr-omethe-us/PyKED}. Future directions for development
-are outlined in the public roadmap\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}.
-The highest priority issues are currently adding support for other types of fundamental
-experiments, including flame speed, speciation, and flame extinction
+are outlined in the public
+roadmap\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}. The
+highest priority issues are currently adding support for other types of
+fundamental experiments, including flame speed, speciation, and flame extinction
 measurements.

### 2017-02-28 pyked.tex
talk -> paper

@@
 the community lacks a widely adopted standard format for this data. This impedes
-both sharing and wide use by the community. In this talk, we will introduce a
+both sharing and wide use by the community. In this paper, we will introduce a
 new Chemical Kinetics Experimental Data format, ChemKED, and the related

### 2017-02-28 pyked.tex
fix for minted formatting

@@
 With these methods, the composition of the reactant mixture can be output to a
-format that can be used to set the composition of a Cantera \pybox{Solution}.
+format that can be used to set the composition of a Cantera~\pybox{Solution}.
 Note that the \pybox{composition} specification does not contain the molecular
@@
 
-The \pybox{ChemKED} class has one instance method defined, called
+The \pybox{ChemKED} class has one instance method defined:
 \pybox{get_dataframe()}. Using this method, the user can get an instance of a

### 2017-02-28 pyked.tex
minor wording fix

@@
 %
-This block utilizes the ability of YAML files to define an anchor using the
-\yabox{&} symbol and refer to that section later using the \yabox{*} symbol, as
+This block uses the ability of YAML files to define an anchor with the
+\yabox{&} symbol and refer to that section later with the \yabox{*} symbol, as
 can be seen in the definition of the \yabox{datapoints} section:

### 2017-03-30 pyked.tex
changed style to ICCK

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{ussci}
+\documentclass[12pt]{icck}
 
@@
 %======================================================================
-\newcommand\papertopic{Reaction Kinetics}
 %======================================================================
@@
 
-\author[1*]{Bryan W. Weber}
-\author[2]{Kyle E. Niemeyer}
+\author[1]{Bryan W.~Weber}
+\author[2]{Kyle E.~Niemeyer\thanks{Corresponding author: \email{kyle.niemeyer@oregonstate.edu}}}
 
@@
 \affil[2]{School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
-\affil[*]{Corresponding author: \email{bryan.weber@uconn.edu}}
 
@@
 
+\begin{keyword}
+    Chemical kinetics\sep Experimental data\sep Autoignition\sep Data standard \\
+\end{keyword}
+
 %====================================================================
 \begin{abstract} % not to exceed 200 words
-Fundamental experimental measurements of quantities including ignition delay
-times, laminar flame speeds, and species profiles serve important roles in
-understanding fuel chemistry and validating chemical kinetic models. However,
-despite both the importance and abundance of such information in the literature,
-the community lacks a widely adopted standard format for this data. This impedes
-both sharing and wide use by the community. In this paper, we will introduce a
-new Chemical Kinetics Experimental Data format, ChemKED, and the related
-Python-based package for creating and validating ChemKED-formatted files called
-PyKED. We will also review past and related efforts, and motivate the need for a
-new solution. ChemKED currently supports the representation of autoignition
-delay time and laminar flame speed measurements. ChemKED-formatted files contain
-all of the information needed to simulate experimental data points, including
-uncertainty. ChemKED is based on the YAML data serialization language, and is
-intended as a human- and machine-readable standard for easy creation and
-automated use. Development of ChemKED and PyKED occurs openly on GitHub under
-the BSD 3-clause license, and contributions from the community are welcome.
-Plans for future development include support for experimental data from jet
-stirred reactor, extinction, and speciation measurements.
+Fundamental experimental measurements of quantities including ignition delay times,
+laminar flame speeds, and species profiles serve important roles in understanding
+fuel chemistry and validating chemical kinetic models. However, despite both the
+importance and abundance of such information in the literature, the community lacks
+a widely adopted standard format for this data. This impedes both sharing and wide
+use by the community. In this talk, we will introduce a new Chemical Kinetics
+Experimental Data format, ChemKED, and the related Python-based package for creating
+and validating ChemKED-formatted files called PyKED. We will also review past and
+related efforts, and motivate the need for a new solution. ChemKED currently supports
+the representation of autoignition delay time and laminar flame speed measurements.
+ChemKED-formatted files contain all of the information needed to simulate experimental
+data points, including the uncertainty of the data. ChemKED is based on the YAML data
+serialization language, and is intended as a human- and machine-readable standard for
+easy creation and automated use.
+Development of ChemKED and PyKED occurs openly on GitHub under the
+BSD 3-clause license, and contributions from the community are welcome. Plans for
+future development include support for experimental data from jet stirred reactor,
+extinction, and speciation measurements.
 \end{abstract}
-
-% (Provide 2-4 keywords describing your research. Only abbreviations firmly
-% established in the field may be used. These keywords will be used for
-% sessioning/indexing purposes.)
-\begin{keyword}
-    chemical kinetics\sep data standard\sep software\sep database
-\end{keyword}
 

### 2017-03-30 pyked.tex
updated references and style

@@
 Cantera~\autocite{Cantera:2.3.0} CTI format or
-FlameMaster~\autocite{FlameMaster:ref,FlameMaster:code} lag behind considerably,
+FlameMaster~\autocite{FlameMaster:ref} lag behind considerably,
 although use of the former continually grows due to Cantera's open availability.

### 2017-03-31 pyked.tex
edits through the end of the papre

@@
 \documentclass[12pt]{icck}
+
+\usepackage[version=4]{mhchem}
+
+\usepackage{siunitx}
+\sisetup{group-separator={,},
+     detect-all,
+     binary-units,
+     list-units = single,
+     range-units = single,
+     tophrase = --,
+     per-mode = symbol-or-fraction,
+     separate-uncertainty = true,
+     list-final-separator = {, and }
+}
+\DeclareSIUnit\atm{atm}
 
@@
 format~\autocite{Frenklach:2007bm,You:2011hy} and an associated online database,
-the PrIMe Data Warehouse~\autocite{PrIMe}. PrIMe files encode fundamental
+the PrIMe Data Warehouse.\footnote{\url{http://primekinetics.org}}
+PrIMe files encode fundamental
 combustion experimental data, along with kinetic models and calculated
@@
 facility-specific effects. However, ReSpecTh experimental files are another
-XML-based format, and as such, suffer from the same usability issues as PrIMe.
+XML-based format, and, as such, suffer from the same usability issues as PrIMe.
 Moreover, the standard remains closed, requiring potential users to register
@@
 addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
-files. Note that in the following, text with gray background is a \yabox{yaml}
-keyword or value, while text with a yellow background is \pybox{Python} code.
+files. We indicate \yabox{yaml} keywords or values using text with gray background,
+and \pybox{Python} code using text with a yellow background.
 
-\ck{} files use the YAML data serialization format. This format offers the
-advantages of being human readable, written in plain text, and having parsers in
-most common programming languages, including Python, C++, Java, Perl, MATLAB,
-and many others. The YAML syntax is quite simple. The basic file structure is
-made of mappings, delimited by a colon. The key for the mapping is on the left
-of the colon, and the value on the right can be a single value, a sequence of
-values, a nested mapping, or some combination of these:
+\ck{} files use the YAML data serialization format~\autocite{yaml:1.2}. This
+format offers the advantages of being human readable, written in plain text, and
+having parsers in most common programming languages, including Python, C++, Java,
+Perl, MATLAB, and many others. The YAML syntax is quite simple. The basic file
+structure is made of mappings, delimited by a colon. The key for the mapping is
+on the left of the colon, and the value on the right can be a single value,
+a sequence of values, a nested mapping, or some combination of these:
 %
@@
         \item \yabox{year} (required, integer): The year of publication
+        \item \yabox{doi} (optional, string): The article DOI
+        \item \yabox{volume} (required, integer): The journal volume number
+        \item \yabox{pages} (optional, string): The article pages
+        \item \yabox{detail} (optional, string): A description of from where the
+        data originated (e.g., figure or table number)
     \end{itemize}
@@
     \item \yabox{apparatus} (required, string): The type of apparatus used to
-    perform the experiment
+    perform the experiment (\yabox{shock tube} or \yabox{rapid compression machine})
 \end{itemize}
@@
     \end{itemize}
-    \item \yabox{ignition-type} (required, mapping); The type of ignition
+    \item \yabox{ignition-type} (required, mapping): The type of ignition
     measured in the experiments. The required fields are
@@
         \item \yabox{type} (required, string): How the ignition delay was
-        measured, one of \yabox{d/dt max} to indicate the ignition point is
+        measured; one of \yabox{d/dt max} to indicate the ignition point is
         found at the maximum of the time derivative of the \yabox{target},
@@
         \item \yabox{target} (required, string): The target for the type of the
-        ignition point, one of \yabox{temperature}, \yabox{pressure},
+        ignition point; one of \yabox{temperature}, \yabox{pressure},
         \yabox{OH}, \yabox{OH*}, \yabox{CH}, or \yabox{CH*}
@@
 The basic user interface to the \ck{} file in \pk{} is through the
-\pybox{ChemKED} class. The \pybox{ChemKED} class constructor takes the file name
+\pybox{ChemKED} class. The \pybox{ChemKED} class constructor takes the filename
 of a \ck{} file or a Python dictionary containing the contents of a \ck{} file
@@
 
-The \pybox{DataPoint} class has two instance methods defined, called
+The \pybox{DataPoint} class defines two instance methods:
 \pybox{get_cantera_mole_fraction()} and \pybox{get_cantera_mass_fraction()}.
 With these methods, the composition of the reactant mixture can be output to a
-format that can be used to set the composition of a Cantera~\pybox{Solution}.
-Note that the \pybox{composition} specification does not contain the molecular
+format that can be used to set the composition of a
+Cantera~\pybox{Solution}~\autocite{Cantera:2.3.0}.
+The \pybox{composition} specification does not contain the molecular
 weights of the components, so conversion between mole fractions and mass
-fractions is not possible.
+fractions is not currently possible.
 
-The \pybox{ChemKED} class has one instance method defined:
+The \pybox{ChemKED} class defines one instance method:
 \pybox{get_dataframe()}. Using this method, the user can get an instance of a
-Pandas~\autocite{pandas} \pybox{DataFrame} that contains the information in the
+Pandas \pybox{DataFrame}~\autocite{pandas} that contains the information in the
 list of \pybox{DataPoint}s. The user can specify the columns included in the
@@
 files, Cerberus for validating \ck{} files~\autocite{cerberus}, and
-pytest~\autocite{pytest:3.0.1} for unit testing. Travis-CI~\autocite{Travis2016}
-provides continuous integration testing.
+pytest~\autocite{pytest:3.0.1} for unit testing.
+Travis-CI\footnote{\url{https://travis-ci.org/}} provides continuous integration
+testing.
 
@@
 Then, we find the information regarding the article in the literature from which
-this data was taken. In this case, the data point is from the work of
+this data was taken. In this case, the dataset comes from the work of
 \textcite{Mittal2006a}:
@@
 %
-Finally, there is just a single datapoint in this file, which codes the ignition
+Finally, there is just a single datapoint in this file, which describes the ignition
 delay for the experiment, the mixture composition, the initial temperature,
 pressure, compression time, ignition type, and the volume history that specifies
-how the volume of the reactor varies with time, simulating the compression
+how the volume of the reactor varies with time, for simulating the compression
 stroke and post-compression processes.
@@
 %
-The values for the \yabox{volume-history} are truncated to save space. A user
-might like to load the information from this file to perform a simulation to
-calculate the ignition delay using Cantera. All of the information required to
+The values for the \yabox{volume-history} are truncated here to save space. A user
+might want to load the information from this file to perform a simulation using
+Cantera to calculate the ignition delay. All of the information required to
 perform this simulation is present in the \ck{} file, with the exception of a
-chemical kinetic mechanism for H$_2$/CO combustion.
+chemical kinetic model for \ce{H2}\slash {CO} combustion.
 %
@@
 \begin{pythonbox}
-T_0 = dp.temperature.to('K').magnitude
-P_0 = dp.pressure.to('Pa').magnitude
-X_0 = dp.get_cantera_mole_fraction()
+T_initial = dp.temperature.to('K').magnitude
+P_initial = dp.pressure.to('Pa').magnitude
+X_initial = dp.get_cantera_mole_fraction()
 
@@
 gas = ct.Solution('h2-co-mechanism.cti')
-gas.TPX = T_0, P_0, X_0
+gas.TPX = T_initial, P_initial, X_initial
 
@@
 is available in CanSen~\autocite{cansen}, in the
-\pybox{cansen.profile.VolumeProfile} class.
+\pybox{cansen.profile.VolumeProfile} class. Then, a Cantera \pybox{ReactorNet}
+can be used to advance the state through autoignition, in this case to an end
+time of \SI{50}{\milli\second}:
 %
@@
 %
-In this \ck{} file, multiple data points are specified with some common
+This \ck{} file specifies multiple data points with some common
 conditions, including a common mixture composition and common definition of
@@
 \end{yamlbox}
+The \yabox{common-properties} section is not required, but can save space and
+help avoid errors when all data points share some common values.
 
-In this example, we would like to run constant volume simulations at each of the
+In this example, we would like to run constant-volume simulations at each of the
 pressure and temperature conditions in the \yabox{datapoints} list. Once again,
 the \ck{} file specifies all the information required for the simulations except
-for the chemical kinetic mechanism.
+for the chemical kinetic model, and Cantera can be used to simulate autoignition:
 %
@@
 # just take the composition of the first DataPoint
-X_0 = ck.datapoints[0].get_cantera_mole_fraction()
+X_initial = ck.datapoints[0].get_cantera_mole_fraction()
 
@@
 for dp in ck.datapoints:
-    T_0 = dp.temperature.to('K').magnitude
-    P_0 = dp.pressure.to('Pa').magnitude
-    gas.TPX = T_0, P_0, X_0
+    T_initial = dp.temperature.to('K').magnitude
+    P_initial = dp.pressure.to('Pa').magnitude
+    gas.TPX = T_initial, P_initial, X_initial
     reac = ct.IdealGasReactor(gas)
     netw = ct.ReactorNet([reac])
-    # Define ignition delay as T_0 + 400 K
-    while reac.T < T_0 + 400:
+    # Define ignition delay as T_initial + 400 K
+    while reac.T < T_initial + 400:
         netw.step()
 
-    print('The ignition delay for T_0={}, P_0={} is: {} seconds'.format(
-        T_0, P_0, netw.time
-    ))
+    print('The ignition delay for T_initial={}, P_initial={} is: '
+          '{} seconds'.format(T_initial, P_initial, netw.time)
+          )
 \end{pythonbox}
 %
-Note that the ignition delay in this example is defined as equal to $T_0 +
-400\,\mathrm{K}$, a simplified definition used for this example. In general, the
-user could post-process the concentration information from the simulation to
-determine one-half the maximum OH* concentration to match the experimental
-definition of ignition delay.
+The ignition delay in this example is defined as equal to the \SI{400}{\kelvin}
+above the initial temperature, a simplified definition used for this example.
+In general, the user could post-process the concentration information from the
+simulation to determine one-half the maximum \ce{OH^*} concentration to match
+the experimental definition of ignition delay.
 %====================================================================

### 2017-06-06 pyked.tex
Update with BY license and PDF metadata

--- 2017-03-30 pyked.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{icck}
+\documentclass[12pt]{ussci}
 
 \usepackage{todonotes}
+
+\usepackage{xmpincl}
+\includexmp{CC_Attribution_4.0_International}
+
+\hypersetup{
+    pdftitle={ChemKED: a human- and machine-readable data standard for chemical kinetics experiments},%
+    pdfauthor={Bryan W. Weber},%
+}
 
@@
 %======================================================================
+\newcommand\papertopic{Reaction Kinetics}
 %======================================================================
@@
 
-\author[1]{Bryan W.~Weber}
-\author[2]{Kyle E.~Niemeyer\thanks{Corresponding author: \email{kyle.niemeyer@oregonstate.edu}}}
+\author[1*]{Bryan W. Weber}
+\author[2]{Kyle E. Niemeyer}
 
@@
 \affil[2]{School of Mechanical, Industrial, and Manufacturing Engineering, Oregon State University, Corvallis, OR, USA}
+\affil[*]{Corresponding author: \email{bryan.weber@uconn.edu}}
 
@@
 
-\begin{keyword}
-    Chemical kinetics\sep Experimental data\sep Autoignition\sep Data standard \\
-\end{keyword}
-
 %====================================================================
 \begin{abstract} % not to exceed 200 words
-Fundamental experimental measurements of quantities including ignition delay times,
-laminar flame speeds, and species profiles serve important roles in understanding
-fuel chemistry and validating chemical kinetic models. However, despite both the
-importance and abundance of such information in the literature, the community lacks
-a widely adopted standard format for this data. This impedes both sharing and wide
-use by the community. In this talk, we will introduce a new Chemical Kinetics
-Experimental Data format, ChemKED, and the related Python-based package for creating
-and validating ChemKED-formatted files called PyKED. We will also review past and
-related efforts, and motivate the need for a new solution. ChemKED currently supports
-the representation of autoignition delay time and laminar flame speed measurements.
-ChemKED-formatted files contain all of the information needed to simulate experimental
-data points, including the uncertainty of the data. ChemKED is based on the YAML data
-serialization language, and is intended as a human- and machine-readable standard for
-easy creation and automated use.
-Development of ChemKED and PyKED occurs openly on GitHub under the
-BSD 3-clause license, and contributions from the community are welcome. Plans for
-future development include support for experimental data from jet stirred reactor,
-extinction, and speciation measurements.
+Fundamental experimental measurements of quantities including ignition delay
+times, laminar flame speeds, and species profiles serve important roles in
+understanding fuel chemistry and validating chemical kinetic models. However,
+despite both the importance and abundance of such information in the literature,
+the community lacks a widely adopted standard format for this data. This impedes
+both sharing and wide use by the community. In this paper, we will introduce a
+new Chemical Kinetics Experimental Data format, ChemKED, and the related
+Python-based package for creating and validating ChemKED-formatted files called
+PyKED. We will also review past and related efforts, and motivate the need for a
+new solution. ChemKED currently supports the representation of autoignition
+delay time and laminar flame speed measurements. ChemKED-formatted files contain
+all of the information needed to simulate experimental data points, including
+uncertainty. ChemKED is based on the YAML data serialization language, and is
+intended as a human- and machine-readable standard for easy creation and
+automated use. Development of ChemKED and PyKED occurs openly on GitHub under
+the BSD 3-clause license, and contributions from the community are welcome.
+Plans for future development include support for experimental data from jet
+stirred reactor, extinction, and speciation measurements.
 \end{abstract}
+
+% (Provide 2-4 keywords describing your research. Only abbreviations firmly
+% established in the field may be used. These keywords will be used for
+% sessioning/indexing purposes.)
+\begin{keyword}
+    chemical kinetics\sep data standard\sep software\sep database
+\end{keyword}
 

### 2017-07-18 pyked.tex
style name -> ijck
replace inline code highlighting with listings

--- 2017-03-31 pyked.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{icck}
+\documentclass[12pt]{ijck}
+
+%\usepackage{listings}
+
+\usepackage{fancyvrb,newverbs,xcolor}
+
+\definecolor{cverbbg}{gray}{0.93}
+\colorlet{pyyellow}{yellow!10!white}
+\colorlet{yamlblack}{black!10!white}
+
+% \newcommand{\yabox}[1]{\colorbox{yamlblack}{\texttt{#1}}}
+% \newcommand{\pybox}[1]{\colorbox{pyyellow}{\texttt{#1}}}
+\newverbcommand{\yabox}
+  {\setbox\verbbox\hbox\bgroup}
+  {\egroup\colorbox{yamlblack}{\box\verbbox}}
+\newverbcommand{\pybox}
+  {\setbox\verbbox\hbox\bgroup}
+  {\egroup\colorbox{pyyellow}{\box\verbbox}}
 
@@
 This is not a complete listing of available fields and more complete examples of
-\ck{} files can be found in the \pybox{tests} directory of the \pk{} package. In
+\ck{} files can be found in the \pybox|tests| directory of the \pk{} package. In
 addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
-files. We indicate \yabox{yaml} keywords or values using text with gray background,
-and \pybox{Python} code using text with a yellow background.
+files. We indicate \yabox|yaml| keywords or values using text with gray background,
+and \pybox|Python| code using text with a yellow background.
 
@@
 
-The \yabox{value} can be a string, integer, or floating point number. The \ck{}
+The \yabox|value| can be a string, integer, or floating point number. The \ck{}
 format is designed to include all of the information necessary to simulate a
@@
 details all of the ``meta'' information about the experiment and the \ck{} file.
-In this section, the common \yabox{author} mapping is used for any fields where
-an author must be specified. The \yabox{author} mapping contains the following
+In this section, the common \yabox|author| mapping is used for any fields where
+an author must be specified. The \yabox|author| mapping contains the following
 fields:
@@
 \begin{itemize}
-    \item \yabox{name} (required, string): The author's full name
-    \item \yabox{ORCID} (optional, string): the author's ORCID number
+    \item \yabox|name| (required, string): The author's full name
+    \item \yabox|ORCID| (optional, string): the author's ORCID number
 \end{itemize}
@@
 \begin{itemize}
-    \item \yabox{file-author} (required, \yabox{author}-type mapping): The
+    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
     author of the \ck{} file
-    \item \yabox{file-version} (required, integer): The version of the \ck{}
+    \item \yabox|file-version| (required, integer): The version of the \ck{}
     file
-    \item \yabox{chemked-version} (required, string): The version of the \ck{}
+    \item \yabox|chemked-version| (required, string): The version of the \ck{}
     schema that this file conforms to
-    \item \yabox{reference} (required, mapping): The reference information for
+    \item \yabox|reference| (required, mapping): The reference information for
     the article associated with the data in the file. The fields in this
@@
     \begin{itemize}
-        \item \yabox{authors} (required, sequence): A sequence of \yabox{author}
+        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
         mappings
-        \item \yabox{journal} (required, string): The name of the publishing
+        \item \yabox|journal| (required, string): The name of the publishing
         journal
-        \item \yabox{year} (required, integer): The year of publication
-        \item \yabox{doi} (optional, string): The article DOI
-        \item \yabox{volume} (required, integer): The journal volume number
-        \item \yabox{pages} (optional, string): The article pages
-        \item \yabox{detail} (optional, string): A description of from where the
+        \item \yabox|year| (required, integer): The year of publication
+        \item \yabox|doi| (optional, string): The article DOI
+        \item \yabox|volume| (required, integer): The journal volume number
+        \item \yabox|pages| (optional, string): The article pages
+        \item \yabox|detail| (optional, string): A description of from where the
         data originated (e.g., figure or table number)
     \end{itemize}
-    \item \yabox{experiment-type} (required, string): Currently, only
-    \yabox{ignition delay} is supported
-    \item \yabox{apparatus} (required, string): The type of apparatus used to
-    perform the experiment (\yabox{shock tube} or \yabox{rapid compression machine})
+    \item \yabox|experiment-type| (required, string): Currently, only
+    \yabox|ignition delay| is supported
+    \item \yabox|apparatus| (required, string): The type of apparatus used to
+    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
 \end{itemize}
@@
 \begin{itemize}
-    \item \yabox{temperature} (required, sequence): The temperature of the
+    \item \yabox|temperature| (required, sequence): The temperature of the
     experiment, with units
-    \item \yabox{ignition-delay} (required, sequence): The ignition delay of the
+    \item \yabox|ignition-delay| (required, sequence): The ignition delay of the
     experiment, with units
-    \item \yabox{pressure} (required, string): The pressure of the experiment,
+    \item \yabox|pressure| (required, string): The pressure of the experiment,
     with units
-    \item \yabox{composition} (required, sequence): The composition of the
+    \item \yabox|composition| (required, sequence): The composition of the
     mixture in the experiment, a sequence of mappings with the following fields
     \begin{itemize}
-        \item \yabox{species-name} (required, string): The name of the species
-        \item \yabox{mole-fraction}, \yabox{mass-fraction}, \yabox{mole-percent}
+        \item \yabox|species-name| (required, string): The name of the species
+        \item \yabox|mole-fraction|, \yabox|mass-fraction|, \yabox|mole-percent|
         (required, float): The mole fraction or percent, or the mass fraction.
-        Only one of the three is permitted in a given \yabox{composition}
+        Only one of the three is permitted in a given \yabox|composition|
         sequence.
-        \item \yabox{InChI}, \yabox{SMILES}, \yabox{atomic-composition}, or
-        \yabox{elemental-composition} (required, string or sequence): The InChI
+        \item \yabox|InChI|, \yabox|SMILES|, \yabox|atomic-composition|, or
+        \yabox|elemental-composition| (required, string or sequence): The InChI
         or SMILES representing the molecule, or its atomic or elemental
-        composition (\yabox{atomic-composition} and
-        \yabox{elemental-composition} are synonyms) as a sequence
+        composition (\yabox|atomic-composition| and
+        \yabox|elemental-composition| are synonyms) as a sequence
     \end{itemize}
-    \item \yabox{ignition-type} (required, mapping): The type of ignition
+    \item \yabox|ignition-type| (required, mapping): The type of ignition
     measured in the experiments. The required fields are
     \begin{itemize}
-        \item \yabox{type} (required, string): How the ignition delay was
-        measured; one of \yabox{d/dt max} to indicate the ignition point is
-        found at the maximum of the time derivative of the \yabox{target},
-        \yabox{max} or \yabox{min} to indicate the ignition point is at the
-        maximum or minimum of the \yabox{target}, or \yabox{1/2 max} to
+        \item \yabox|type| (required, string): How the ignition delay was
+        measured; one of \yabox|d/dt max| to indicate the ignition point is
+        found at the maximum of the time derivative of the \yabox|target|,
+        \yabox|max| or \yabox|min| to indicate the ignition point is at the
+        maximum or minimum of the \yabox|target|, or \yabox|1/2 max| to
         indicate the half-maximum point
-        \item \yabox{target} (required, string): The target for the type of the
-        ignition point; one of \yabox{temperature}, \yabox{pressure},
-        \yabox{OH}, \yabox{OH*}, \yabox{CH}, or \yabox{CH*}
+        \item \yabox|target| (required, string): The target for the type of the
+        ignition point; one of \yabox|temperature|, \yabox|pressure|,
+        \yabox|OH|, \yabox|OH*|, \yabox|CH|, or \yabox|CH*|
     \end{itemize}
@@
 
-The top-level key for the sequence is called \yabox{datapoints}. Each of the
+The top-level key for the sequence is called \yabox|datapoints|. Each of the
 quantities in an element must be specified with units. This should be done as a
@@
 The basic user interface to the \ck{} file in \pk{} is through the
-\pybox{ChemKED} class. The \pybox{ChemKED} class constructor takes the filename
+\pybox|ChemKED| class. The \pybox|ChemKED| class constructor takes the filename
 of a \ck{} file or a Python dictionary containing the contents of a \ck{} file
@@
 The fields of the \ck{} file are stored as instance attributes of the
-\pybox{ChemKED} class. The following attributes are available:
+\pybox|ChemKED| class. The following attributes are available:
 %
 \begin{itemize}
-    \item \pybox{chemked_version}, \pybox{file_version}, \pybox{file_author},
-    \pybox{experiment_type}: Stores the values from the \ck{} file
-    \item \pybox{reference}: Instance of a \pybox{namedtuple} containing all
+    \item \pybox|chemked_version|, \pybox|file_version|, \pybox|file_author|,
+    \pybox|experiment_type|: Stores the values from the \ck{} file
+    \item \pybox|reference|: Instance of a \pybox|namedtuple| containing all
     the information from the literature reference associated with the data
-    \item \pybox{apparatus}: Instance of a \pybox{namedtuple} containing all
+    \item \pybox|apparatus|: Instance of a \pybox|namedtuple| containing all
     the information about the apparatus used to perform the experiment
-    \item \pybox{datapoints}: A Python list of \pybox{DataPoint} instances
+    \item \pybox|datapoints|: A Python list of \pybox|DataPoint| instances
 \end{itemize}
 
-The \pybox{DataPoint} class stores the information associated with a single data
-point in the \ck{} file (i.e., a single element of the \yabox{datapoints}
-sequence). Similar to the \pybox{ChemKED} class, information in the
-\pybox{DataPoint} is stored as instance attributes:
+The \pybox|DataPoint| class stores the information associated with a single data
+point in the \ck{} file (i.e., a single element of the \yabox|datapoints|
+sequence). Similar to the \pybox|ChemKED| class, information in the
+\pybox|DataPoint| is stored as instance attributes:
 %
 \begin{itemize}
-    \item \pybox{equivalence_ratio}: The value of the equivalence ratio, if
+    \item \pybox|equivalence_ratio|: The value of the equivalence ratio, if
     present. For informational purposes only, no validation is done of the
     value.
-    \item \pybox{ignition_delay}, \pybox{temperature}, \pybox{pressure},
-    \pybox{pressure_rise}: The value of these data are stored as instances of
-    the \pybox{Quantity} class from the Pint \autocite{Grecco2016} package,
+    \item \pybox|ignition_delay|, \pybox|temperature|, \pybox|pressure|,
+    \pybox|pressure_rise|: The value of these data are stored as instances of
+    the \pybox|Quantity| class from the Pint \autocite{Grecco2016} package,
     possibly with associated uncertainty
-    \item \pybox{composition}: A list of dictionaries of the species and their
-    respective amounts. The values are validated so that \yabox{mole-percent},
-    \yabox{mole-fraction}, or \yabox{mass-fraction} cannot be mixed for a single
+    \item \pybox|composition|: A list of dictionaries of the species and their
+    respective amounts. The values are validated so that \yabox|mole-percent|,
+    \yabox|mole-fraction|, or \yabox|mass-fraction| cannot be mixed for a single
     data point, and so that the sum of the values is approximately 1.0, or
-    100.0 for \yabox{mole-percent}.
-    \item \pybox{composition_type}: A string indicating the type of composition
-    information for the data point---one of \pybox{'mole-percent'},
-    \pybox{'mole-fraction'}, or \pybox{'mass-fraction'}
-    \item \pybox{volume_history}: If the \yabox{volume-history} of an RCM
+    100.0 for \yabox|mole-percent|.
+    \item \pybox|composition_type|: A string indicating the type of composition
+    information for the data point---one of \pybox|'mole-percent'|,
+    \pybox|'mole-fraction'|, or \pybox|'mass-fraction'|
+    \item \pybox|volume_history|: If the \yabox|volume-history| of an RCM
     experiment is provided in the \ck{} file, it is stored in this attribute as
-    a \pybox{namedtuple}, and the actual values are stored in NumPy arrays
+    a \pybox|namedtuple|, and the actual values are stored in NumPy arrays
     \autocite{vanderWalt:2011np}
-    \item \pybox{compression_time}: If the \ck{} file encodes an RCM experiment,
+    \item \pybox|compression_time|: If the \ck{} file encodes an RCM experiment,
     the length of the compression stroke may be included
-    \item \pybox{ignition_type}: A dictionary specifying the method of the
+    \item \pybox|ignition_type|: A dictionary specifying the method of the
     measurement of ignition delay
@@
 
-The \pybox{DataPoint} class defines two instance methods:
-\pybox{get_cantera_mole_fraction()} and \pybox{get_cantera_mass_fraction()}.
+The \pybox|DataPoint| class defines two instance methods:
+\pybox|get_cantera_mole_fraction()| and \pybox|get_cantera_mass_fraction()|.
 With these methods, the composition of the reactant mixture can be output to a
 format that can be used to set the composition of a
-Cantera~\pybox{Solution}~\autocite{Cantera:2.3.0}.
-The \pybox{composition} specification does not contain the molecular
+Cantera~\pybox|Solution|~\autocite{Cantera:2.3.0}.
+The \pybox|composition| specification does not contain the molecular
 weights of the components, so conversion between mole fractions and mass
@@
 
-The \pybox{ChemKED} class defines one instance method:
-\pybox{get_dataframe()}. Using this method, the user can get an instance of a
-Pandas \pybox{DataFrame}~\autocite{pandas} that contains the information in the
-list of \pybox{DataPoint}s. The user can specify the columns included in the
-\pybox{DataFrame} by passing a list of column names to the
-\pybox{output_columns} argument of the \pybox{get_dataframe()} method. The
+The \pybox|ChemKED| class defines one instance method:
+\pybox|get_dataframe()|. Using this method, the user can get an instance of a
+Pandas \pybox|DataFrame|~\autocite{pandas} that contains the information in the
+list of \pybox|DataPoint|s. The user can specify the columns included in the
+\pybox|DataFrame| by passing a list of column names to the
+\pybox|output_columns| argument of the \pybox|get_dataframe()| method. The
 possible columns are not case-sensitive and are:
@@
 \noindent\begin{itemize*}
-    \item \pybox{'Temperature'}
-    \item \pybox{'Pressure'}
-    \item \pybox{'Ignition Delay'}
-    \item \pybox{'Composition'}
-    \item \pybox{'Equivalence Ratio'}
-    \item \pybox{'Reference'}
-    \item \pybox{'Apparatus'}
-    \item \pybox{'Experiment Type'}
-    \item \pybox{'File Author'}
-    \item \pybox{'File Version'}
-    \item \pybox{'ChemKED Version'}
+    \item \pybox|'Temperature'|
+    \item \pybox|'Pressure'|
+    \item \pybox|'Ignition Delay'|
+    \item \pybox|'Composition'|
+    \item \pybox|'Equivalence Ratio'|
+    \item \pybox|'Reference'|
+    \item \pybox|'Apparatus'|
+    \item \pybox|'Experiment Type'|
+    \item \pybox|'File Author'|
+    \item \pybox|'File Version'|
+    \item \pybox|'ChemKED Version'|
 \end{itemize*}
 
-In addition, specific fields from the \pybox{Reference} and \pybox{Apparatus}
+In addition, specific fields from the \pybox|Reference| and \pybox|Apparatus|
 attributes can be included by specifying the name after a colon. These options
@@
 \noindent\begin{itemize*}
-    \item \pybox{'Reference:Volume'}
-    \item \pybox{'Reference:Journal'}
-    \item \pybox{'Reference:DOI'}
-    \item \pybox{'Reference:Authors'}
-    \item \pybox{'Reference:Detail'}
-    \item \pybox{'Reference:Year'}
-    \item \pybox{'Reference:Pages'}
-    \item \pybox{'Apparatus:Kind'}
+    \item \pybox|'Reference:Volume'|
+    \item \pybox|'Reference:Journal'|
+    \item \pybox|'Reference:DOI'|
+    \item \pybox|'Reference:Authors'|
+    \item \pybox|'Reference:Detail'|
+    \item \pybox|'Reference:Year'|
+    \item \pybox|'Reference:Pages'|
+    \item \pybox|'Apparatus:Kind'|
 \end{itemize*}
@@
 \noindent\begin{itemize*}
-    \item \pybox{'Apparatus:Facility'}
-    \item \pybox{'Apparatus:Institution'}
+    \item \pybox|'Apparatus:Facility'|
+    \item \pybox|'Apparatus:Institution'|
 \end{itemize*}
 
-Only the first author is printed when \pybox{Reference} or
-\pybox{Reference:Authors} is selected because the whole author list may be
+Only the first author is printed when \pybox|Reference| or
+\pybox|Reference:Authors| is selected because the whole author list may be
 quite long.
@@
 \url{https://github.com/pr-omethe-us/PyKED}~\autocite{PyKED}. Installation of stable versions
-can be accomplished from a package repository, using either \pybox{pip} to install
+can be accomplished from a package repository, using either \pybox|pip| to install
 from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
-\pybox{conda} to install from the
+\pybox|conda| to install from the
 \href{https://anaconda.org/conda-forge}{\texttt{conda-forge}} channel. In
 addition, \pk{} can be installed from the source code by running
-\pybox|python setup.py install| or \pybox{pip install .} from the root
+\pybox|python setup.py install| or \pybox|pip install .| from the root
 directory of the repository.
@@
 The \ck{} file that will be used in this example can be found in the
-\pybox{tests} directory of the PyKED
+\pybox|tests| directory of the PyKED
 repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}.
@@
 %
-The values for the \yabox{volume-history} are truncated here to save space. A user
+The values for the \yabox|volume-history| are truncated here to save space. A user
 might want to load the information from this file to perform a simulation using
@@
 The initial temperature, pressure, and mixture composition can be read from the
-instance of the \pybox{DataPoint} class. \pk{} uses Pint \pybox{Quantities} to
+instance of the \pybox|DataPoint| class. \pk{} uses Pint \pybox|Quantities| to
 store values with units, while Cantera expects a floating point value in SI
@@
 %
-To apply the effect of the volume trace to the \pybox{IdealGasReactor}, a
-\pybox{Wall} is installed between the reactor and the environment and assigned
+To apply the effect of the volume trace to the \pybox|IdealGasReactor|, a
+\pybox|Wall| is installed between the reactor and the environment and assigned
 a velocity. Although we do not show the details here, a reference implementation
@@
 is available in CanSen~\autocite{cansen}, in the
-\pybox{cansen.profile.VolumeProfile} class. Then, a Cantera \pybox{ReactorNet}
+\pybox|cansen.profile.VolumeProfile| class. Then, a Cantera \pybox|ReactorNet|
 can be used to advance the state through autoignition, in this case to an end
@@
 Although not shown in this example, the user would typically store information
-about the \pybox{IdealGasReactor} (e.g., temperature, pressure, mass fractions)
+about the \pybox|IdealGasReactor| (e.g., temperature, pressure, mass fractions)
 at each time step of the integration for post-processing.
@@
 
-The \ck{} file used in this example can be found in the \pybox{tests} directory of the \pk{}
+The \ck{} file used in this example can be found in the \pybox|tests| directory of the \pk{}
 repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}.
@@
 conditions, including a common mixture composition and common definition of
-ignition delay. Therefore, a \yabox{common-properties} section is specified.
+ignition delay. Therefore, a \yabox|common-properties| section is specified.
 %
@@
 This block uses the ability of YAML files to define an anchor with the
-\yabox{&} symbol and refer to that section later with the \yabox{*} symbol, as
-can be seen in the definition of the \yabox{datapoints} section:
+\yabox|&| symbol and refer to that section later with the \yabox|*| symbol, as
+can be seen in the definition of the \yabox|datapoints| section:
 %
@@
 \end{yamlbox}
-The \yabox{common-properties} section is not required, but can save space and
+The \yabox|common-properties| section is not required, but can save space and
 help avoid errors when all data points share some common values.
@@
 In this example, we would like to run constant-volume simulations at each of the
-pressure and temperature conditions in the \yabox{datapoints} list. Once again,
+pressure and temperature conditions in the \yabox|datapoints| list. Once again,
 the \ck{} file specifies all the information required for the simulations except

### 2017-07-18 pyked.tex
fix schema link, add text about testing

@@
 schema used for validation of the \ck{} files is available publicly in the
-source code repository for
-PyKED\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/chemked_schema.yaml}}.
+\texttt{schemas} directory of the PyKED source code
+repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/tree/master/pyked/schemas}}.
+To ease extension of the schema to describe other experimental measurements, it
+comprises multiple YAML files.
 
@@
 between units, \texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML
-files, Cerberus for validating \ck{} files~\autocite{cerberus}, and
-pytest~\autocite{pytest:3.0.1} for unit testing.
+files, and Cerberus for validating \ck{} files~\autocite{cerberus}.
+Furthermore, \pk{} includes extensive unit, integration, and functional tests
+that ensure all aspects of the software operate as intended; others have pointed
+out the importance of automated testing in scientific software~\autocite{Wilson:bestpractices}.
+\pk{} relies on pytest~\autocite{pytest:3.0.1} for unit testing, and
 Travis-CI\footnote{\url{https://travis-ci.org/}} provides continuous integration
-testing.
+testing (i.e., runs the test suite after any new changes are made or pull-requested).
 
@@
 %====================================================================
-\section{Usage example}\label{sec:usage-example}
+\section{Usage examples}\label{sec:usage-example}
 

### 2017-07-18 pyked.tex
clarified current version

@@
 
-This section will provide a brief introduction to the \ck{} database format.
+This section will provide a brief introduction to the current version of the
+\ck{} database format, v0.1.6, which represents a functional standard
+for documenting autoignition measurements in shock tubes and rapid compression machines.
 This is not a complete listing of available fields and more complete examples of
@@
 having parsers in most common programming languages, including Python, C++, Java,
-Perl, MATLAB, and many others. The YAML syntax is quite simple. The basic file
+Perl, MATLAB, and many others. The YAML syntax is quite simple: the basic file
 structure is made of mappings, delimited by a colon. The key for the mapping is

### 2017-07-18 pyked.tex
add link to github proposal issue

@@
 
-In this paper, we presented the \ck{} data format for describing measurements
+In this article, we presented the \ck{} data format for describing measurements
 taken from fundamental combustion experiments, recognizing that the community
@@
 highest priority issues are currently adding support for other types of
-fundamental experiments, including flame speed, speciation, and flame extinction
-measurements.
+fundamental experiments, including speciation in shock tubes, flame speeds,
+and flame extinction measurements. We also welcome proposals for new experimental
+types; see Issue \#60 in the \pk{} repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
 

### 2017-07-18 pyked.tex
completed final sentence

@@
 \texttt{schemas} directory of the PyKED source code
-repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/tree/master/pyked/schemas}}.
+repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/tree/master/pyked/schemas}}
 To ease extension of the schema to describe other experimental measurements, it
@@
 \pybox|tests| directory of the PyKED
-repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}.
+repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
 Examining that file, we find the first section specifies the information about
@@
 The \ck{} file used in this example can be found in the \pybox|tests| directory of the \pk{}
-repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}.
+repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
 The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
@@
 are outlined in the public
-roadmap\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}. The
-highest priority issues are currently adding support for other types of
+roadmap.\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}
+The highest priority issues are currently adding support for other types of
 fundamental experiments, including speciation in shock tubes, flame speeds,
 and flame extinction measurements. We also welcome proposals for new experimental
-types; see Issue \#60 in the \pk{} repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
+types; see Issue \#60 in the \pk{} repository for an example of how this may be
+done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
 

### 2017-07-18 pyked.tex
ruamel.yaml -> PyYAML in text

@@
 for array manipulation, Pint~\autocite{Grecco2016} for interpreting and converting
-between units, \texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML
+between units, PyYAML~\autocite{pyyaml} for parsing YAML
 files, and Cerberus for validating \ck{} files~\autocite{cerberus}.

### 2017-07-20 pyked.tex
Update examples to v0.1.6 format

@@
 for documenting autoignition measurements in shock tubes and rapid compression machines.
-This is not a complete listing of available fields and more complete examples of
-\ck{} files can be found in the \pybox|tests| directory of the \pk{} package. In
-addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
+This is not a complete listing of available fields and a complete reference for the
+schema can be found in the online documentation.
+\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
+In addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
 files. We indicate \yabox|yaml| keywords or values using text with gray background,
@@
 format is designed to include all of the information necessary to simulate a
-given experiment and can be broken into two main sections. The first section
-details all of the ``meta'' information about the experiment and the \ck{} file.
-In this section, the common \yabox|author| mapping is used for any fields where
-an author must be specified. The \yabox|author| mapping contains the following
-fields:
+given experiment. A \ck{} file can generally be broken into two main sections.
+The first section details all of the ``meta'' information about the experiment
+and the \ck{} file. In this section, the common \yabox|author| mapping is used
+for any fields where an author must be specified. The \yabox|author| mapping
+contains the following fields:
 %
@@
   ORCID: 0000-0003-4425-7097
-file-version: (1, 0)
-chemked-version: 0.0.1
+file-version: 0
+chemked-version: 0.1.6
 \end{yamlbox}
@@
 datapoints:
-  - temperature: 297.4 kelvin
-    ignition-delay: 1.0 ms
-    pressure: 958.0 torr
+- temperature:
+    - 297.4 kelvin
+  ignition-delay:
+    - 1.0 ms
+  pressure:
+    - 958.0 torr
   composition:
+    kind: mole fraction
+    species:
       - species-name: H2
         InChI: 1S/H2/h1H
-        mole-fraction: 0.12500
+        amount:
+          - 0.12500
       - species-name: O2
         InChI: 1S/O2/c1-2
-        mole-fraction: 0.06250
+        amount:
+          - 0.06250
       - species-name: N2
         InChI: 1S/N2/c1-2
-        mole-fraction: 0.18125
+        amount:
+          - 0.18125
       - species-name: Ar
         InChI: 1S/Ar
-        mole-fraction: 0.63125
+        amount:
+          - 0.63125
   ignition-type:
@@
     type: d/dt max
-    compression-time: 38.0 ms
+  compression-time:
+    - 38.0 ms
   volume-history:
@@
       - [2.00E-003, 5.43427034574E+002]
-        - [3.00E-003, 5.38124109043E+002]
       ...
@@
   composition: &comp
+    kind: mole fraction
+    species:
       - species-name: t-butanol
         InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
-      mole-fraction: 0.003333333
+        amount:
+          - 0.003333333
       - species-name: O2
         InChI:  1S/O2/c1-2
-      mole-fraction: 0.04
+        amount:
+          - 0.04
       - species-name: Ar
         InChI:  1S/Ar
-      mole-fraction: 0.956666667
+        amount:
+          - 0.956666667
   ignition-type: &ign
-    target: OH* concentration
+    target: OH*
     type: 1/2 max
@@
 datapoints:
-  - temperature: 1459 kelvin
-    ignition-delay: 347 us
-    pressure: 1.60 atm
+  - temperature:
+      - 1459 kelvin
+    ignition-delay:
+      - 347 us
+    pressure:
+      - 1.60 atm
     composition: *comp
@@
     equivalence-ratio: 0.5
-  - temperature: 1389 kelvin
-    ignition-delay: 756 us
-    pressure: 1.67 atm
+  - temperature:
+      - 1389 kelvin
+    ignition-delay:
+      - 756 us
+    pressure:
+      - 1.67 atm
     composition: *comp
@@
     equivalence-ratio: 0.5
-  - temperature: 1497 kelvin
-    ignition-delay: 212 us
-    pressure: 1.55 atm
+  - temperature:
+      - 1497 kelvin
+    ignition-delay:
+      - 212 us
+    pressure:
+      - 1.55 atm
     composition: *comp
@@
     equivalence-ratio: 0.5
-  - temperature: 1562 kelvin
-    ignition-delay: 105 us
-    pressure: 1.50 atm
+  - temperature:
+      - 1562 kelvin
+    ignition-delay:
+      - 105 us
+    pressure:
+      - 1.50 atm
     composition: *comp
@@
 The \yabox|common-properties| section is not required, but can save space and
-help avoid errors when all data points share some common values.
+help avoid errors when several data points share some common values. However, even if
+properties are specified in the \yabox|common-properties| section, they must also be
+explicitly specified in the \yabox|datapoints| section for each data point, as can be
+seen in the previous code.
 

### 2017-07-20 pyked.tex
Add text about database, repo locations, and docs

@@
 
+ChemKED files are stored online in a git repository hosted on GitHub.
+\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
+Submissions from the community are welcomed and can be made via the standard Pull Request
+on GitHub. The authors also encourage questions and problems to be submitted either via
+email or as Issues on GitHub. In particular, researchers interested in converters for
+internal data formats to the ChemKED format are encouraged to reach out. In addition,
+converters to and from the ReSpecTh format are available in the PyKED repository.
+
 %====================================================================
@@
 out the importance of automated testing in scientific software~\autocite{Wilson:bestpractices}.
-\pk{} relies on pytest~\autocite{pytest:3.0.1} for unit testing, and
-Travis-CI\footnote{\url{https://travis-ci.org/}} provides continuous integration
-testing (i.e., runs the test suite after any new changes are made or pull-requested).
+\pk{} relies on pytest~\autocite{pytest:3.0.1} for unit testing, and currently has
+100\% line and branch coverage for all of the executable code in the repository. This
+means that the tests execute every line of code at least once, and every branch statement
+(e.g., if-then-else statement) is tested so that every condition is executed at least
+once.
+
+Travis-CI\footnote{\url{https://travis-ci.org/pr-omethe-us/PyKED}} provides continuous
+integration service, including running the test suite, building and distributing binary
+installer packages, and building and publishing the online documentation. This automation
+ensures that the test suite is run after every change to the source code, the
+documentation is up-to-date with the latest changes, and releasing new versions is as
+streamlined as possible.
 
@@
 source code is publicly available on GitHub at
-\url{https://github.com/pr-omethe-us/PyKED}~\autocite{PyKED}. Installation of stable versions
-can be accomplished from a package repository, using either \pybox|pip| to install
-from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
-\pybox|conda| to install from the
-\href{https://anaconda.org/conda-forge}{\texttt{conda-forge}} channel. In
-addition, \pk{} can be installed from the source code by running
-\pybox|python setup.py install| or \pybox|pip install .| from the root
-directory of the repository.
+\url{https://github.com/pr-omethe-us/PyKED}~\autocite{PyKED}. Installation instructions
+are available in the online documentation at
+\url{https://pr-omethe-us.github.io/PyKED/install.html} and can be done via binary
+packages or from the source code.
 
@@
 \ck{} and \pk{} are currently under active development, and contributions from
-the community are welcomed. All development occurs under the BSD 3-clause
+the community are welcomed. All development occurs under the BSD 3-clause open source
 license and the code is housed on GitHub at
@@
 fundamental experiments, including speciation in shock tubes, flame speeds,
-and flame extinction measurements. We also welcome proposals for new experimental
-types; see Issue \#60 in the \pk{} repository for an example of how this may be
-done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
+and flame extinction measurements.
+Questions, comments, or suggestions are welcomed and can be posted as Issues in the
+GitHub repository or emailed to the authors. In particular, we welcome proposals for new
+experimental types; see Issue \#60 in the \pk{} repository for an example of how this may
+be done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
 

### 2017-07-24 pyked.tex
minor wording fix

@@
 out the importance of automated testing in scientific software~\autocite{Wilson:bestpractices}.
-\pk{} relies on pytest~\autocite{pytest:3.0.1} for unit testing, and currently has
+\pk{} relies on pytest~\autocite{pytest:3.0.1} for automated testing, and currently has
 100\% line and branch coverage for all of the executable code in the repository. This
-means that the tests execute every line of code at least once, and every branch statement
-(e.g., if-then-else statement) is tested so that every condition is executed at least
-once.
+means that the tests execute every line of code at least once, and evaluate
+every branch statement (e.g., if-then-else statements) to execute every condition
+at least once.
 
@@
 installer packages, and building and publishing the online documentation. This automation
-ensures that the test suite is run after every change to the source code, the
-documentation is up-to-date with the latest changes, and releasing new versions is as
-streamlined as possible.
+runs the test suite after every change to the source code, keeps the documentation
+up-to-date with the latest changes, and streamlines the release of new versions.
 

### 2017-07-24 pyked.tex
more minor fixes

@@
 \pk{} is licensed under the permissive, open-source BSD 3-clause license. The
-source code is publicly available on GitHub at
-\url{https://github.com/pr-omethe-us/PyKED}~\autocite{PyKED}. Installation instructions
-are available in the online documentation at
-\url{https://pr-omethe-us.github.io/PyKED/install.html} and can be done via binary
-packages or from the source code.
+source code is publicly available on GitHub at \url{https://github.com/pr-omethe-us/PyKED},
+and versioned releases are automatically archived via Zenodo~\autocite{PyKED}.
+\pk{} can be installed via binary packages or the source code; the online documentation
+provides detailed installation instructions at
+\url{https://pr-omethe-us.github.io/PyKED/install.html}.
 

### 2017-07-24 pyked.tex
fixed abstract

@@
 a widely adopted standard format for this data. This impedes both sharing and wide
-use by the community. In this talk, we will introduce a new Chemical Kinetics
-Experimental Data format, ChemKED, and the related Python-based package for creating
-and validating ChemKED-formatted files called PyKED. We will also review past and
+use by the community. Here we introduce a new chemical kinetics
+experimental data format, ChemKED, and the related Python-based package for creating
+and validating ChemKED-formatted files called PyKED. We also review past and
 related efforts, and motivate the need for a new solution. ChemKED currently supports
-the representation of autoignition delay time and laminar flame speed measurements.
+the representation of autoignition delay time measurements from shock tubes and rapid
+compression machines.
 ChemKED-formatted files contain all of the information needed to simulate experimental
@@
 BSD 3-clause license, and contributions from the community are welcome. Plans for
-future development include support for experimental data from jet stirred reactor,
-extinction, and speciation measurements.
+future development include support for experimental data from laminar flame,
+jet stirred reactor, and speciation measurements.
 \end{abstract}
@@
 
-ChemKED files are stored online in a git repository hosted on GitHub.
+\ck{} files are stored online in a git repository hosted on GitHub.
 \footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
@@
 email or as Issues on GitHub. In particular, researchers interested in converters for
-internal data formats to the ChemKED format are encouraged to reach out. In addition,
-converters to and from the ReSpecTh format are available in the PyKED repository.
+internal data formats to the \ck{} format are encouraged to reach out. In addition,
+the \pk{} package (described in more detail in Sec.~\label{sec:pyked-architecture})
+provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
 

### 2017-07-24 pyked.tex
edits through beginning of section 2

@@
 available in figures and must be digitized by the use of software such as
-WebPlotDigitizer~\autocite{WebPlotDigitizer}. These practices limit the wide use
+WebPlotDigitizer~\autocite{WebPlotDigitizer}. These practices limit wide use
 of the valuable, extensive experimental data available in the literature.
@@
 the information needed to simulate an experiment, including a machine-readable
-definition of ignition or a standard way to express detailed facility specific
-effects important to properly simulating certain experiments. Second, the XML
+definition of ignition or a standard way to express detailed facility-specific
+effects necessary to properly simulate certain experiments. Second, the XML
 format is intended to be a markup language rather than a data format, and
-presents a barrier to creating and working with database files. Furthermore, the
+its lack of human readibility presents a barrier to creating and working with
+database files. Furthermore, the
 focus on using internal identifiers for bibliographic references and
 species\slash reactions, rather than standard identifiers (e.g., DOIs for
-scholarly products and InCHI or SMILES for species), limits usability outside
-the PrIMe ecosystem. Finally, the closed and opaque nature of the PrIMe standard
+scholarly products, InCHI or SMILES for species), limits usability outside
+the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
 and associated Data Warehouse limit contributions and community development of
@@
 ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
-provides machine-readable formats for specifying ignition experiments, including
+provides machine-readable formats for describing ignition experiments, including
 a field for the definition of autoignition and the ability to specify
@@
 Moreover, the standard remains closed, requiring potential users to register
-with the standard's authors to access the specification.
+with the standard's authors to access the specification and the growing database.
 
@@
 
-Despite the development of these standards, limited examples of open data
-\emph{sharing} using the formats exist in the literature. To the best
+Despite the development of these standards, limited examples exist in the
+literature of open data \emph{sharing} using the formats. To the best
 of our knowledge, the largest such publicly available database is hosted by
 the Clean Combustion Research Center at King Abdullah University of Science and
-Technology (KAUST) and is known as CloudFlame
-(\url{https://cloudflame.kaust.edu.sa}). Started in
-2013~\autocite{Goteng:2013cf,Goteng:2014,ReynoChiasson:2015}, CloudFlame serves
-as an openly accessible database for experimental data, available in a standard
+Technology (KAUST): CloudFlame.\footnote{\url{https://cloudflame.kaust.edu.sa}}
+Started in 2013~\autocite{Goteng:2013cf,Goteng:2014,ReynoChiasson:2015}, CloudFlame
+serves as an openly accessible database for experimental data, available in a standard
 CSV format, and also provides a cloud infrastructure for running simulations
@@
 for documenting autoignition measurements in shock tubes and rapid compression machines.
-This is not a complete listing of available fields and a complete reference for the
+Here we do not include a complete listing of available fields; a complete reference for the
 schema can be found in the online documentation.
@@
 internal data formats to the \ck{} format are encouraged to reach out. In addition,
-the \pk{} package (described in more detail in Sec.~\label{sec:pyked-architecture})
+the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
 provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.

### 2017-07-24 pyked.tex
edits through the end of section 3

@@
 
-Therefore, we believe there remains a need for an open, community-focused and
+We therefore believe there remains a need for an open, community-focused and
 developed, combustion\slash chemical kinetics data format. In this work, we
@@
 
-This section will provide a brief introduction to the current version of the
+This section provides a brief introduction to the current version of the
 \ck{} database format, v0.1.6, which represents a functional standard
@@
 Here we do not include a complete listing of available fields; a complete reference for the
-schema can be found in the online documentation.
-\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
-In addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
-files. We indicate \yabox|yaml| keywords or values using text with gray background,
-and \pybox|Python| code using text with a yellow background.
+schema can be found in the online
+documentation.\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
+In Section~\ref{sec:usage-example}, we provide examples of \ck{} files and
+representative use cases. We indicate \yabox|yaml| keywords or values using text with
+gray background, and \pybox|Python| code using text with a yellow background.
 
@@
 Perl, MATLAB, and many others. The YAML syntax is quite simple: the basic file
-structure is made of mappings, delimited by a colon. The key for the mapping is
+structure consists of mappings, delimited by a colon. The key for the mapping is
 on the left of the colon, and the value on the right can be a single value,
@@
 
-The \yabox|value| can be a string, integer, or floating point number. The \ck{}
-format is designed to include all of the information necessary to simulate a
-given experiment. A \ck{} file can generally be broken into two main sections.
-The first section details all of the ``meta'' information about the experiment
-and the \ck{} file. In this section, the common \yabox|author| mapping is used
-for any fields where an author must be specified. The \yabox|author| mapping
+The \yabox|value| can be a string, integer, or floating-point number. We designed
+the \ck{} format to include all information necessary to simulate a
+given experiment. A \ck{} file is generally broken into two main sections.
+The first section details all the ``meta'' information about the experiment
+and the \ck{} file itself. This section uses the common \yabox|author| mapping
+for any fields that specify an author. The \yabox|author| mapping
 contains the following fields:
@@
     \item \yabox|name| (required, string): The author's full name
-    \item \yabox|ORCID| (optional, string): the author's ORCID number
+    \item \yabox|ORCID| (optional, string): the author's unique ORCID code
 \end{itemize}
@@
     \item \yabox|chemked-version| (required, string): The version of the \ck{}
-    schema that this file conforms to
+    schema to which this file conforms
     \item \yabox|reference| (required, mapping): The reference information for
-    the article associated with the data in the file. The fields in this
+    the published article associated with the data in the file. The fields in this
     mapping are:
@@
 
-The second section of the file encodes the experimental data. In the \ck{} file,
-this is done in a sequence of mappings, with the following information required
+The second section of the file encodes the experimental data. The \ck{} file
+describes these using a sequence of mappings, with the following information required
 in each element of the sequence:
@@
     \item \yabox|composition| (required, sequence): The composition of the
-    mixture in the experiment, a sequence of mappings with the following fields
+    mixture in the experiment, described via a sequence of mappings with the fields
     \begin{itemize}
         \item \yabox|species-name| (required, string): The name of the species
-        \item \yabox|mole-fraction|, \yabox|mass-fraction|, \yabox|mole-percent|
+        \item \yabox|mole-fraction|, \yabox|mole-percent|, \yabox|mass-fraction|
         (required, float): The mole fraction or percent, or the mass fraction.
@@
     \end{itemize}
-    \item \yabox|ignition-type| (required, mapping): The type of ignition
-    measured in the experiments. The required fields are
+    \item \yabox|ignition-type| (required, mapping): The method used to detect
+    ignition in the experiments. The required fields are
     \begin{itemize}
-        \item \yabox|type| (required, string): How the ignition delay was
-        measured; one of \yabox|d/dt max| to indicate the ignition point is
-        found at the maximum of the time derivative of the \yabox|target|,
-        \yabox|max| or \yabox|min| to indicate the ignition point is at the
-        maximum or minimum of the \yabox|target|, or \yabox|1/2 max| to
-        indicate the half-maximum point
-        \item \yabox|target| (required, string): The target for the type of the
-        ignition point; one of \yabox|temperature|, \yabox|pressure|,
+        \item \yabox|type| (required, string): How ignition delay was
+        measured; one of \yabox|d/dt max| (indicates the ignition point was
+        found at the maximum of the time derivative of the \yabox|target|),
+        \yabox|max| or \yabox|min| (indicates the ignition point was at the
+        maximum or minimum of the \yabox|target|), or \yabox|1/2 max| (
+        indicates the half-maximum point of \yabox|target|)
+        \item \yabox|target| (required, string): The target for the ignition
+        \yabox|type| measurement; one of \yabox|temperature|, \yabox|pressure|,
         \yabox|OH|, \yabox|OH*|, \yabox|CH|, or \yabox|CH*|
@@
 single string associated with the first element of the sequence. The units of
-the quantity are validated to ensure they have the appropriate dimensionality
+the quantity are validated to ensure appropriate dimensionality
 for the quantity. In addition, each of the quantities in an element can
 optionally be assigned an uncertainty. This uncertainty can be either absolute
-or relative, and is specified as elements of the sequence of the associated key.
+or relative, and is specified as an element of the sequence of the associated key.
 For example, the absolute uncertainty of the temperature and the relative
-uncertainty of the ignition delay might be specified as follows:
-
+uncertainty of the ignition delay might be specified as:
+%
 \begin{yamlbox}
@@
 
-\ck{} files are stored online in a git repository hosted on GitHub.
-\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
-Submissions from the community are welcomed and can be made via the standard Pull Request
+We have begun building an open repository of \ck{} files, available online at
+GitHub.\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
+We welcome submissions from the community; these can be made via the standard pull request
 on GitHub. The authors also encourage questions and problems to be submitted either via
-email or as Issues on GitHub. In particular, researchers interested in converters for
-internal data formats to the \ck{} format are encouraged to reach out. In addition,
+email or as issues on GitHub. In particular, researchers interested in converters for
+internal data formats to the \ck{} format are encouraged to reach out;
 the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
-provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
+currently provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
+While we plan to continue growing this database---and hope for submissions from
+the community---others can freely download files, or in fact copy (``fork,'' in GitHub
+parlance) the entire database for their own purposes.
 
@@
 %
-\pk{} is a Python package that currently provides the reference implementation
+\pk{} is a Python package that provides the reference implementation
 of the interface to \ck{} files~\autocite{PyKED}. \pk{} reads \ck{} files, validates
@@
 
-The basic user interface to the \ck{} file in \pk{} is through the
-\pybox|ChemKED| class. The \pybox|ChemKED| class constructor takes the filename
+\pk{} provides the basic user interface to a \ck{} file through the
+\pybox|ChemKED| class. The \pybox|ChemKED| class constructor takes the name
 of a \ck{} file or a Python dictionary containing the contents of a \ck{} file
-as its argument. When the file or dictionary is loaded, it is validated to have
-the correct format by the Python package Cerberus \autocite{cerberus}. The
+as its argument. When the file or dictionary is loaded, \pk{} validates its
+format and contents using the Python package Cerberus \autocite{cerberus}. The
 schema used for validation of the \ck{} files is available publicly in the
@@
 repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/tree/master/pyked/schemas}}
-To ease extension of the schema to describe other experimental measurements, it
-comprises multiple YAML files.
+The schema comprises multiple YAML files to ease its extension to describe other
+experimental measurements.
 
@@
 point in the \ck{} file (i.e., a single element of the \yabox|datapoints|
-sequence). Similar to the \pybox|ChemKED| class, information in the
-\pybox|DataPoint| is stored as instance attributes:
+sequence). Similar to the \pybox|ChemKED| class,
+the \pybox|DataPoint| stores information as instance attributes:
 %
@@
     \item \pybox|equivalence_ratio|: The value of the equivalence ratio, if
-    present. For informational purposes only, no validation is done of the
+    present. For informational purposes only---no validation is done of the
     value.
@@
 \pybox|get_cantera_mole_fraction()| and \pybox|get_cantera_mass_fraction()|.
-With these methods, the composition of the reactant mixture can be output to a
+These methods output the composition of the reactant mixture to a
 format that can be used to set the composition of a
@@
 weights of the components, so conversion between mole fractions and mass
-fractions is not currently possible.
+fractions is not currently possible. (Future versions of \pk{} may support this
+feature via, e.g., online lookup of species using their InChI\slash SMILES
+identifiers.)
 
-The \pybox|ChemKED| class defines one instance method:
-\pybox|get_dataframe()|. Using this method, the user can get an instance of a
+The \pybox|ChemKED| class defines three instance methods:
+\pybox|get_dataframe()|, \pybox|write_file()|, and \pybox|convert_to_ReSpecTh()|.
+The \pybox|get_dataframe()| method provides an instance of a
 Pandas \pybox|DataFrame|~\autocite{pandas} that contains the information in the
@@
 possible columns are not case-sensitive and are:
-
+%
 \noindent\begin{itemize*}
@@
 are:
-
+%
 \noindent\begin{itemize*}
@@
     \item \pybox|'Apparatus:Kind'|
+    \item \pybox|'Apparatus:Facility'|
 \end{itemize*}
@@
 \noindent\begin{itemize*}
-    \item \pybox|'Apparatus:Facility'|
     \item \pybox|'Apparatus:Institution'|
@@
 \pybox|Reference:Authors| is selected because the whole author list may be
-quite long.
+quite long. The \pybox|write_file()| method writes another \ck{} YAML file based on
+the internal attributes, while \pybox|convert_to_ReSpecTh()| converts the \ck{}
+to a ReSpecTh XML file. In the latter case, some information may be lost, as \ck{}
+files both encode more details and specify more granularity of some information than
+ReSpecTh files.
 
@@
 \pk{} relies on pytest~\autocite{pytest:3.0.1} for automated testing, and currently has
-100\% line and branch coverage for all of the executable code in the repository. This
+100\% line and branch coverage for all executable code in the repository. This
 means that the tests execute every line of code at least once, and evaluate
@@
 at least once.
-
 Travis-CI\footnote{\url{https://travis-ci.org/pr-omethe-us/PyKED}} provides continuous

### 2017-07-24 pyked.tex
edits thru rest of paper. moved common-properties description up

@@
 
+Frequently, experimental series hold certain properties constant, such as initial
+mixture composition or pressure, and use the same method to detect ignition delay.
+\ck{} files support a \yabox|common-properties| convenience section where these
+common details may be defined once, and referenced in each datapoint.
+This block uses the ability of YAML files to define an anchor with the
+\yabox|&| symbol and refer to that section later with the \yabox|*| symbol:
+%
+\begin{yamlbox}
+common-properties:
+  composition: &comp
+    kind: mole fraction
+    species:
+      ...
+  ignition-type: &ign
+    ...
+datapoints:
+  composition: *comp
+  ignition-type: *ign
+  temperature: 1000 kelvin
+  ...
+\end{yamlbox}
+%
+The \yabox|common-properties| section is not required, but can save space and
+help avoid errors when several data points share some common values. However, even if
+properties are specified in the \yabox|common-properties| section, they must also be
+explicitly specified and referenced in the \yabox|datapoints| section for each data point,
+as seen in the previous code.
+
 We have begun building an open repository of \ck{} files, available online at
@@
 The \ck{} file that will be used in this example can be found in the
-\pybox|tests| directory of the PyKED
+\texttt{tests} directory of the PyKED
 repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
@@
 %
-Finally, there is just a single datapoint in this file, which describes the ignition
-delay for the experiment, the mixture composition, the initial temperature,
-pressure, compression time, ignition type, and the volume history that specifies
+Finally, this file contains just a single datapoint, which describes the experimental
+ignition delay, initial mixture composition, initial temperature, initial pressure,
+compression time, ignition type, and volume history that specifies
 how the volume of the reactor varies with time, for simulating the compression
@@
 might want to load the information from this file to perform a simulation using
-Cantera to calculate the ignition delay. All of the information required to
-perform this simulation is present in the \ck{} file, with the exception of a
-chemical kinetic model for \ce{H2}\slash {CO} combustion.
+Cantera~\autocite{Cantera:2.3.0} to calculate the ignition delay. All information
+required to perform this simulation is present in the \ck{} file, with the exception
+of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
 %
@@
 instance of the \pybox|DataPoint| class. \pk{} uses Pint \pybox|Quantities| to
-store values with units, while Cantera expects a floating point value in SI
+store values with units, while Cantera expects a floating-point value in SI
 units as input. Therefore, we use the built-in capabilities of Pint to convert
@@
 To apply the effect of the volume trace to the \pybox|IdealGasReactor|, a
-\pybox|Wall| is installed between the reactor and the environment and assigned
+\pybox|Wall| is installed between the reactor and environment, and assigned
 a velocity. Although we do not show the details here, a reference implementation
@@
 
-The \ck{} file used in this example can be found in the \pybox|tests| directory of the \pk{}
+The \ck{} file used in this example can be found in the \texttt{tests} directory of the \pk{}
 repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
@@
     type: 1/2 max
-\end{yamlbox}
-%
-This block uses the ability of YAML files to define an anchor with the
-\yabox|&| symbol and refer to that section later with the \yabox|*| symbol, as
-can be seen in the definition of the \yabox|datapoints| section:
-%
-\begin{yamlbox}
 datapoints:
@@
 \end{yamlbox}
-The \yabox|common-properties| section is not required, but can save space and
-help avoid errors when several data points share some common values. However, even if
-properties are specified in the \yabox|common-properties| section, they must also be
-explicitly specified in the \yabox|datapoints| section for each data point, as can be
-seen in the previous code.
 
-In this example, we would like to run constant-volume simulations at each of the
-pressure and temperature conditions in the \yabox|datapoints| list. Once again,
+In this example, we would like to run constant-volume simulations at each
+pressure and temperature condition in the \yabox|datapoints| list. Once again,
 the \ck{} file specifies all the information required for the simulations except
@@
 The highest priority issues are currently adding support for other types of
-fundamental experiments, including speciation in shock tubes, flame speeds,
+fundamental experiments, including speciation in shock tubes, laminar flame speeds,
 and flame extinction measurements.
-Questions, comments, or suggestions are welcomed and can be posted as Issues in the
+Questions, comments, or suggestions are welcomed and can be posted as issues in the
 GitHub repository or emailed to the authors. In particular, we welcome proposals for new
-experimental types; see Issue \#60 in the \pk{} repository for an example of how this may
+experimental types; see issue \#60 in the \pk{} repository for an example of how this may
 be done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}

### 2017-07-25 pyked.tex
Minor updates to wording and text

@@
 \begin{abstract} % not to exceed 200 words
-Fundamental experimental measurements of quantities including ignition delay times,
-laminar flame speeds, and species profiles serve important roles in understanding
+Fundamental experimental measurements of quantities such as ignition delay times,
+laminar flame speeds, and species profiles (among others) serve important roles in understanding
 fuel chemistry and validating chemical kinetic models. However, despite both the
@@
 combustion experimental data, along with kinetic models and calculated
-quantities, using the Extensible Markup Language (XML) standard.
+quantities, using the eXtensible Markup Language (XML) standard.
 
@@
 effects necessary to properly simulate certain experiments. Second, the XML
-format is intended to be a markup language rather than a data format, and
+format is intended to be a machine-readable markup language rather than a data format, and
 its lack of human readibility presents a barrier to creating and working with
@@
 
-More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the
+More recently, Varga and coworkers~\autocite{Varga2015a,Varga2015b} developed the
 ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
@@
 
-Nonetheless, the usefulness of databases of fundamental combustion experiments
-has been shown to be quite useful. For instance, Olm and colleagues used their
+Nonetheless, databases of fundamental combustion experiments
+have been shown to be quite useful. For instance, Olm and colleagues used their
 ReSpecTh-based database to quantify the performance of literature
@@
 \pybox|get_dataframe()|, \pybox|write_file()|, and \pybox|convert_to_ReSpecTh()|.
-The \pybox|get_dataframe()| method provides an instance of a
+The \pybox|get_dataframe()| method returns an instance of a
 Pandas \pybox|DataFrame|~\autocite{pandas} that contains the information in the
@@
 \pybox|Reference:Authors| is selected because the whole author list may be
-quite long. The \pybox|write_file()| method writes another \ck{} YAML file based on
-the internal attributes, while \pybox|convert_to_ReSpecTh()| converts the \ck{}
-to a ReSpecTh XML file. In the latter case, some information may be lost, as \ck{}
+quite long.
+
+The \pybox|write_file()| method writes a new \ck{} YAML file based on
+the class' instance attributes, while \pybox|convert_to_ReSpecTh()| converts the \pybox|ChemKED|
+instance to a ReSpecTh XML file. In the latter case, some information may be lost, as \ck{}
 files both encode more details and specify more granularity of some information than
@@
 The initial temperature, pressure, and mixture composition can be read from the
-instance of the \pybox|DataPoint| class. \pk{} uses Pint \pybox|Quantities| to
+instance of the \pybox|DataPoint| class. \pk{} uses instances of the Pint \pybox|Quantity| class to
 store values with units, while Cantera expects a floating-point value in SI

### 2017-07-25 pyked.tex
Update schema and API description with new format

@@
     \item \yabox|temperature| (required, sequence): The temperature of the
-    experiment, with units
+    experiment, with units and optionally uncertainty
     \item \yabox|ignition-delay| (required, sequence): The ignition delay of the
-    experiment, with units
-    \item \yabox|pressure| (required, string): The pressure of the experiment,
-    with units
-    \item \yabox|composition| (required, sequence): The composition of the
-    mixture in the experiment, described via a sequence of mappings with the fields
+    experiment, with units and optionally uncertainty
+    \item \yabox|pressure| (required, sequence): The pressure of the experiment,
+    with units and optionally uncertainty
+    \item \yabox|composition| (required, mapping): The composition of the
+    mixture in the experiment, described via a nested mapping with the fields
+    \begin{itemize}
+        \item \yabox|kind| (required, string): one of \yabox|mole fraction|, \yabox|mole percent|, or \yabox|mass fraction|
+        \item \yabox|species|: sequence of mappings, with the fields
         \begin{itemize}
             \item \yabox|species-name| (required, string): The name of the species
-        \item \yabox|mole-fraction|, \yabox|mole-percent|, \yabox|mass-fraction|
-        (required, float): The mole fraction or percent, or the mass fraction.
-        Only one of the three is permitted in a given \yabox|composition|
-        sequence.
+            \item \yabox|amount| (required, sequence): The mole fraction or percent, or the mass fraction, optionally with uncertainty
             \item \yabox|InChI|, \yabox|SMILES|, \yabox|atomic-composition|, or
             \yabox|elemental-composition| (required, string or sequence): The InChI
-        or SMILES representing the molecule, or its atomic or elemental
-        composition (\yabox|atomic-composition| and
+            or SMILES string representing the molecule, or its atomic or elemental
+            composition (\yabox|atomic-composition| and\\
             \yabox|elemental-composition| are synonyms) as a sequence
+        \end{itemize}
     \end{itemize}
@@
     \item \pybox|chemked_version|, \pybox|file_version|, \pybox|file_author|,
-    \pybox|experiment_type|: Stores the values from the \ck{} file
+    \pybox|experiment_type|: Store the ``meta'' values from the \ck{} file
     \item \pybox|reference|: Instance of a \pybox|namedtuple| containing all
@@
     the information about the apparatus used to perform the experiment
-    \item \pybox|datapoints|: A Python list of \pybox|DataPoint| instances
+    \item \pybox|datapoints|: A Python \pybox|list| of \pybox|DataPoint| instances
 \end{itemize}
@@
     the length of the compression stroke may be included
+    \item \pybox|compressed_temperature|: If the \ck{} file encodes an RCM experiment,
+    the temperature at the end of compression may be included
+    \item \pybox|compressed_pressure|: If the \ck{} file encodes an RCM experiment,
+    the pressure at the end of compression may be included
     \item \pybox|ignition_type|: A dictionary specifying the method of the

### 2017-07-25 pyked.tex
few final edits

@@
 Nonetheless, databases of fundamental combustion experiments
-have been shown to be quite useful. For instance, Olm and colleagues used their
+have proved quite useful. For instance, Olm and colleagues used their
 ReSpecTh-based database to quantify the performance of literature
@@
 
-The \pybox|write_file()| method writes a new \ck{} YAML file based on
-the class' instance attributes, while \pybox|convert_to_ReSpecTh()| converts the \pybox|ChemKED|
+The \pybox|write_file()| method writes a new \ck{} YAML file based on the instance attributes
+of the class, while \pybox|convert_to_ReSpecTh()| converts the \pybox|ChemKED|
 instance to a ReSpecTh XML file. In the latter case, some information may be lost, as \ck{}

### 2017-09-25 rebuttal.tex
added initial rebuttal; needs reviewer 2 comments

\documentclass[a4paper,10pt]{elsarticle}
\usepackage[utf8]{inputenc}

\biboptions{sort&compress, square, comma}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}

\usepackage[usenames, dvipsnames]{xcolor}
\usepackage{siunitx}
\sisetup{group-separator={,},
     detect-all,
     binary-units,
     list-units = single,
     range-units = single,
     range-phrase = --,
     per-mode = symbol-or-fraction,
     separate-uncertainty = true,
     multi-part-units = single,
     list-final-separator = {, and }
%    scientific-notation = fixed
}
\providecommand{\hly}[1]{\colorbox{yellow}{#1}}
\providecommand{\hlb}[1]{\colorbox{SkyBlue}{#1}}
\providecommand{\hlg}[1]{\colorbox{green}{#1}}

% bump up section font size
\usepackage{sectsty}
\sectionfont{\fontsize{12}{15}\selectfont}

\usepackage[colorlinks=true]{hyperref}
\usepackage{todonotes}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.4cm}

\begin{document}

We thank the reviewers for their helpful and detailed comments, and have addressed their remarks here.
The manuscript has been significantly revised, and we marked significant changes corresponding to suggestions from Reviewers \#1 and \#2 in \hlg{yellow} and \hly{green}, respectively.

\section*{Reviewer \#1:}

\textit{Currently four database formats are offered for storing combustion experimental data. These are CloudFlame (Excel/CSV format), PrIMe (XML-based format), ReSpecTh (also XML-based format), and ChemKED (YAML-based format). Elaboration of the last format is reported in the submitted manuscript. The motivation for the elaboration of this new format was based on the statement that YAML is more human readable than XML. I think that this is simply not true. Even a first-time user can read the raw datafiles of both formats with equal efficiency. Also, currently the suggested new format is much more limited (e.g. it is able to encode ignition delay time experiments only). The presently planned features of ChemKED are already available in the other three formats.}

\textit{However, I suggest the publication of an improved and shortened version of the manuscript in IJCK. The reason is that a permanent and laboratory-independent storage of experimental data is an important issue and I agree that a public discussion promotes the acceptance of a widely used common format.}

\textbf{Reviewer comments:}
\textit{I think that the paper contains too much information science details. Still, as the authors note, the description is not comprehensive and they refer to the user manual. I suggest the creation of a significantly abridged version that contains the basic information technology points only and does not discuss the details. For example, Chapter 4 could be deleted and Chapters 2 and 3 may also be shortened.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind considerably, although use of the former continually grows due to Cantera?s open availability.''
Note that FlameMaster is also freely available from the GitLab repository of the developers.}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{p.~2 ``Second, the XML format is intended to be a machine-readable markup language rather than a data format, and its lack of human readibility presents a barrier to creating and working with database files.''
XML is a text format and it is fully human readable, similarly to YMAL. Even a first-time user is able to read the chemical content of a  PrIMe XML file and an experienced user is able to write such a file.}

\textit{I suggest adding to following block to this para: ``The original PrIMe was on an Amazon cloud server. Recently it was shifted to the computer center of DLR Stuttgart. At the new location it is called ``PrIMe 2.0'' [\url{https://prime.cki-know.org/}]. This is identical to the original PrIMe apart from technical changes due to the different computation background. Professor Frenklach plans to develop further PrIMe (using name PrIMe 3.0) and this version will be availabe at [\url{http://primekinetics.org/}].''
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for typical bibliographic data.'' 
I suggest the following extension:
``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for typical bibliographic data, including the DOI of the cited arecicle. Each datafile has a unique own DOI and unique URL address. [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}]''. Also species can be identified by InCHI or SMILES.
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{``The ReSpecTh standard also provides machine-readable formats for describing ignition experiments, including a field for the definition of autoignition and the ability to specify facility-specific effects.''
Modified text:
``The ReSpecTh standard provides formats that are both human and machine-readable for describing ignition experiments, including a field for the definition of autoignition and the ability to specify facility-specific effects. Also, it describes PSR and flow reactor speciation data, burner stabilized flame speciation data and burning velocity data.''
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{``However, ReSpecTh experimental files are another XML-based format, and, as such, suffer from the same usability issues as PrIMe.''
I think that the limits of an XML database is set by the actual set of tags and not the format.
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{``Moreover, the standard remains closed, requiring potential users to register with the standard's authors to access the specification and the growing database.''
I suggest instead:
``Registration is open for everyone who provides a valid institutional e-mail address.''
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{
``... hydrogen/syngas [11] and ethanol [12]. In all four cases, they converted numerous experimental datasets from the literature into the ReSpecTh standard.''
Extended version:
``... hydrogen/syngas [11], methanol and formaldehyde [C. Olm, T. Varga, E. Valko, H. J. Curran, T. Turanyi: Uncertainty quantification of a newly optimized methanol and formaldehyde combustion mechanism, Combustion and Flame, in press] and ethanol [12]. In all five cases, they converted numerous experimental datasets from the literature into the ReSpecTh standard. The database currently contains  93240 experimental data points in 1376 XML-format data files [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}]''.
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{``type (required, string): How ignition delay was measured;''
The options described here is a very limited subset of the ignition delay time definitions used in experimental articles. For example, backward or forward interpolated values of a measured quantity are further possibilities.  
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{p.~8 ``In the latter case, some information may be lost, as ChemKED files both encode more details and specify more granularity of some information than ReSpecTh files.''
This was true for ReSpecTh format 1.0, but no longer true for ReSpecTh Kinetic Dataformat 2.0. The latter covers all info of the ChemKED files (and more). 
}

\textbf{Our response:}

\textbf{Reviewer comment:}
\textit{Missing data in the references: Year is missing is refs. [5], [6], [9], [10], [11], [12], [21],[25],[27].
}

\textit{[3] Please add reference: \url{https://www.itv.rwth-aachen.de/index.php?id=13}
Also, the latest version is from 2017.
}

\textit{[7] Further info:  Paper P1-04, ISBN 978-963-12-1257-0}

\textit{[8] Please refer to ReSpecTh Kinetics Data Format Specification 2.0 instead.}

\textbf{Our response:}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2:}

\section*{References}
\bibliography{rebuttal}
\bibliographystyle{elsarticle-num}

\end{document}

### 2017-09-26 rebuttal.tex
added reviewer 2 comments to rebuttal

@@
 
+\textit{My main concern with the creation of a publically available database, formatted with ChemKED, is that it seems that the creation of the files has to be done manually, i.e.
+susceptible for errors.
+This is especially true if the experiment comprises a big set of datapoints and/or hundreds of measured species.
+Furthermore, the YAML data serialization format is indeed human-readable, but it does not provide an easy summary/overview of the experimental data, such as an excel or CSV file would give, and therefore, people might be not willing to put the extra effort.}
+
+\textbf{Reviewer comment:}
+\textit{1.~How are the authors planning on ‘forcing’ the experimentalists to publish their data in this new ChemKED file format?}
+
+\textbf{Our response:}
+
+\textit{2.~Am I correct to state that ChemKED is a YAML file with a predefined structure?}
+
+\textbf{Our response:}
+
+\textit{3.~I think the article would be more appealing if it would have at least one figure as an example.}
+
+\textbf{Our response:}
+
+\subsection*{Abstract and introduction}
+\textit{4.
+In the abstract, it is mentioned that the Python-based package PyKED creates ChemKED- formatted files.
+However in the article there is no evidence that PyKED aides in creating ChemKED-formatted files.}
+
+\textbf{Our response:}
+
+\textit{5.~Page 1, line 31: The future development plans include support for jet stirred reactors, however, this is not mentioned in section 5, Conclusions and future work, nor is there any reference to it on the GitHub wiki roadmap of PyKED.}
+
+\textbf{Our response:}
+
+\textit{6.
+I don't completely agree with the second part of the statement on page 2 line 16-18: ``Second, the XML format is intended to be a machine-readable mark-up language rather than a data format, and its lack of human readability presents a barrier to creating and working with database files.''}
+
+\textbf{Our response:}
+
+\textit{In my opinion the YAML format resembles the XML format strongly.
+XML format is human- readable, however, because of the html-tags and `\texttt{<>} characters, the YAML format might be easier to read.
+If XML would lack human readability as it is stated in the article, it would not be such a popular format, and I don't think PrIMe and ReSpecTh would be using it.}
+
+\textbf{Our response:}
+
+\textit{7.~An initial version of the ChemKED format was introduced by Niemeyer, as is mentioned in the introduction.
+The mentioned article describes a Python-package called PyTeCK.
+How does this PyTeCK package relate to the PyKED package that is presented in the submitted article?}
+
+\textbf{Our response:}
+
+\textit{8.~Page 3, line 10: ``Our motivations are similar to those of '', `to' is missing.}
+
+\textbf{Our response:}
+
+\subsection*{Overview of ChemKED format}
+
+\textit{9.~A reference should not be a required mapping of an experimental dataset.
+If the intention is to make the community to adopt the ChemKED file structure, it might be easier to do so if experimentalists can use the structure internally as well, i.e.
+to read, store, and share non-published internal data.}
+
+\textbf{Our response:}
+
+\textit{10.~Page 4, line 32: ``The second section of the file encodes the experimental data''.
+Perhaps it is easier to follow if this is briefly mentioned before the authors explain the details of the first section of the ChemKED format.}
+
+\textbf{Our response:}
+
+\textit{11.~Why do both \texttt{atomic-composition} and \texttt{elemental-composition} exist as synonyms? Wouldn't it be clearer if only one of the two can be used as a valid input?}
+
+\textbf{Our response:}
+
+\textit{12.~Is it possible to create for example two \texttt{common-properties} \texttt{composition} blocks A and B and refer to them in the \texttt{DataPoint} keys?}
+
+\textbf{Our response:}
+
+\textit{13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED files, on GitHub.
+Is this the most efficient way for making a publically available database?
+One issue with this is the fact that adding files is done via pull requests, which have to be verified and approved by the owners of the repository.
+Therefore a certain delay is inevitable.}
+
+\textbf{Our response:}
+
+\textit{14.~Page 6, line 18 states that a converter exists from ReSpecTh to ChemKED. But isn't this feature added when going from v0.1.6 to v0.2.0, which makes page 3 line 21: v0.1.6 not correct.}
+
+\textbf{Our response:}
+
+\textit{15.~Page 6, line 21, is the `,' necessary after fork?}
+
+\textbf{Our response:}
+
+\subsection*{PyKED architecture}
+
+\textit{16.~The Python package Cerberus is used to validate the format and content of the given ChemKED file.
+How does this package handles/reports the inconsistencies with the schemas?
+As far as my local test went with giving a corrupted YAML file to the ChemKED initiator, the package just spits out a Python error, which might be not the most user-friendly solution.}
+
+\textbf{Our response:}
+
+\textit{17.~A recurring issue with actively developing codes is backwards compatibility. How are the authors planning on addressing this issue?}
+
+\textbf{Our response:}
+
+\textit{18.~Page 7, line 12, \texttt{mole-percent}, \texttt{mole-fraction}, and \texttt{mass-fraction} include a dash in their name, whereas this is not the case for the example on page 5 line 48 and in the description of the composition element on page 4, line 47.}
+
+\textbf{Our response:}
+
+\textit{19.~Same issue as \#18 holds for page 7, line 17.}
+
+\textbf{Our response:}
+
+\textit{20.~I think it would be more clear if \texttt{volume\_history}, \texttt{compression\_time}, \texttt{compressed\_temperature},
+and \texttt{compressed\_pressure} would be combined. The line: `if the ChemKED file encodes an
+RCM experiment' is unnecessarily repeated.}
+
+\textbf{Our response:}
+
+\textit{21.~Page 7, line 39: ``Future versions of PyKED may support this feature via, e.g. online lookup''.
+There are two main caveats with online lookups: a) It will make the code slower if the internet
+connection is not strong, b) an internet connection is necessary for the code to be executed.}
+
+\textbf{Our response:}
+
+\textit{22.~Page 7, line 40: Why is it interesting to create the YAML file again with the \texttt{write\_file()} function.
+Creating a summary excel/CSV file would be more useful.}
+
+\subsection*{Usage examples}
+\subsubsection*{RCM modelling with varying reactor volume}
+
+\textit{23.~Page 10, line 29--30: ``A user might want to load the information'' Avoid the term `might'.
+It's one of the main goals of using ChemKED and its associated Python-package PyKED to load the information necessary for simulating the experiment.}
+
+\textbf{Our response:}
+
+\textit{24.~The Python code snippets are not always initiated in the text.}
+
+\textbf{Our response:}
+
+\textit{25.~Page 10, line 55: add more commentary text. It is not clear that those lines create the
+necessary instances for Cantera, rather than ChemKED.}
+
+\textbf{Our response:}
+
+\textit{26.~Page 11, line 7: Where is this `air.xml' file coming from. This is not clear from the text.}
+
+\textbf{Our response:}
+
+\subsubsection*{Shock tube modelling with constant volume}
+
+\textit{27.~The table overlapping page 11 and 12 can be reduced by truncating the \texttt{DataPoints}, similarly as is done with the table starting at page 9, line 31 and ending at page 10, line 27.}
+
+\textbf{Our response:}
+
+\textit{28.
+Python code snippet on page 13: (a.) Add more commentary text.
+(b.) Changing the example to enable multithreading might make it more appealing to start using ChemKED as a data format in addition with Cantera as a simulation tool.}
+
+\textbf{Our response:}
+
+\subsection*{Conclusions and future work}
+
+\textit{29.~No conclusion is mentioned for the usage examples.}
+
+\textbf{Our response:}
+
 \section*{References}

### 2017-09-26 pyked.tex
Updated remaining refs in response to #2

@@
 
-More recently, Varga and coworkers~\autocite{Varga2015a,Varga2015b} developed the
+More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
 ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
@@
 the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
-currently provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
+currently provides converters to and from the ReSpecTh format~\autocite{Varga2017}.
 While we plan to continue growing this database---and hope for submissions from

### 2017-09-26 rebuttal.tex
added new ref in response to #2

@@
 The reason is that a permanent and laboratory-independent storage of experimental data is an important issue and I agree that a public discussion promotes the acceptance of a widely used common format.}
+
+\textbf{Our response:}
 
@@
 
-\textbf{Our response:}
-
 \textit{In my opinion the YAML format resembles the XML format strongly.
-XML format is human- readable, however, because of the html-tags and `\texttt{<>} characters, the YAML format might be easier to read.
+XML format is human-readable, however, because of the html-tags and `\texttt{<>} characters, the YAML format might be easier to read.
 If XML would lack human readability as it is stated in the article, it would not be such a popular format, and I don't think PrIMe and ReSpecTh would be using it.}
@@
 \textbf{Our response:}
+We agree with the reviewer that XML is technically human readable by an expert (as opposed to, say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really mean human-\emph{understandable}.
 

### 2017-09-26 pyked.tex
added new ref in response to #2

@@
 and more recently develop improved models for hydrogen\slash
-syngas~\autocite{Varga:2016gj} and ethanol~\autocite{Olm:2016et}. In all four
+syngas~\autocite{Varga:2016gj}, methanol\slash formaldehyde~\autocite{Olm:2017a},
+ and ethanol~\autocite{Olm:2016et}. In all four
 cases, they converted numerous experimental datasets from the literature into

### 2017-09-26 rebuttal.tex
some responses to #3 points 1 and 2

@@
 \textbf{Reviewer comment:}
-\textit{1.~How are the authors planning on ‘forcing’ the experimentalists to publish their data in this new ChemKED file format?}
+\textit{1.~How are the authors planning on `forcing' the experimentalists to publish their data in this new ChemKED file format?}
 
 \textbf{Our response:}
+While we cannot force anyone to use a particular format, we hope that by defining a format and ecosystem with desirable features---including inherent openness and usability---experimentalists will begin adopting a community-standard format.
+In fact, we have already received interest from various groups, and hope this will grow as we continue to improve and add features.
+We also hope that at a certain point, with wide adoption, sharing experimental data in a standard format will become the expected norm (e.g., when submitting an article for publication) like sharing a Chemkin-format chemical kinetic model.
 
@@
 \textbf{Our response:}
+Yes, ChemKED files are files in YAML format, with a prescribed structure and set of contents.
 

### 2017-09-26 pyked.tex
fixed sentence in response to #3

@@
 software tool for validating and working with \ck{} files: \pk{}. Our
-motivations are similar those of the PrIMe and ReSpecTh teams: the \ck{} project
+motivations resemble those of the PrIMe and ReSpecTh teams: the \ck{} project
 will enable easy sharing and use of fundamental combustion data, for the

### 2017-09-27 pyked.tex
removed comma for #3

@@
 While we plan to continue growing this database---and hope for submissions from
-the community---others can freely download files, or in fact copy (``fork,'' in GitHub
+the community---others can freely download files, or in fact copy (``fork'' in GitHub
 parlance) the entire database for their own purposes.

### 2017-09-27 rebuttal.tex
fixed sentence in response to #3
removed comma for #3
Change rebuttal format to use environments

--- 2017-09-26 rebuttal.tex
@@
 
+\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup\vspace{0.5\baselineskip}}
+\newenvironment{response}{\vspace{0.5\baselineskip}\begin{response}
+
+\end{response}}{\vspace{0.5\baselineskip}}
+\newcommand{\review}[1]{{\itshape#1}}
+\newcommand{\respond}[1]{\begin{response}
+
+\end{response} #1}
+
 \begin{document}
@@
 
-\textit{Currently four database formats are offered for storing combustion experimental data.
-These are CloudFlame (Excel/CSV format), PrIMe (XML-based format), ReSpecTh (also XML-based format), and ChemKED (YAML-based format).
-Elaboration of the last format is reported in the submitted manuscript.
-The motivation for the elaboration of this new format was based on the statement that YAML is more human readable than XML.
-I think that this is simply not true.
-Even a first-time user can read the raw datafiles of both formats with equal efficiency.
-Also, currently the suggested new format is much more limited (e.g.
-it is able to encode ignition delay time experiments only).
-The presently planned features of ChemKED are already available in the other three formats.}
+\begin{reviewer}
+    %
+    Currently four database formats are offered for storing combustion experimental data. These are
+    CloudFlame (Excel/CSV format), PrIMe (XML-based format), ReSpecTh (also XML-based format), and
+    ChemKED (YAML-based format). Elaboration of the last format is reported in the submitted
+    manuscript. The motivation for the elaboration of this new format was based on the statement
+    that YAML is more human readable than XML. I think that this is simply not true. Even a
+    first-time user can read the raw datafiles of both formats with equal efficiency. Also,
+    currently the suggested new format is much more limited (e.g. it is able to encode ignition
+    delay time experiments only). The presently planned features of ChemKED are already available in
+    the other three formats.
 
-\textit{However, I suggest the publication of an improved and shortened version of the manuscript in IJCK.
-The reason is that a permanent and laboratory-independent storage of experimental data is an important issue and I agree that a public discussion promotes the acceptance of a widely used common format.}
+    However, I suggest the publication of an improved and shortened version of the manuscript in
+    IJCK. The reason is that a permanent and laboratory-independent storage of experimental data is
+    an important issue and I agree that a public discussion promotes the acceptance of a widely used
+    common format.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+    %
+    We would like to thank the reviewer for their comments. Detailed responses to each comment are
+    below.
+    %
+\end{response}
 
-\textbf{Reviewer comments:}
-\textit{I think that the paper contains too much information science details.
-Still, as the authors note, the description is not comprehensive and they refer to the user manual.
-I suggest the creation of a significantly abridged version that contains the basic information technology points only and does not discuss the details.
-For example, Chapter 4 could be deleted and Chapters 2 and 3 may also be shortened.}
+\begin{reviewer}
+    %
+    I think that the paper contains too much information science details. Still, as the authors
+    note, the description is not comprehensive and they refer to the user manual. I suggest the
+    creation of a significantly abridged version that contains the basic information technology
+    points only and does not discuss the details. For example, Chapter 4 could be deleted and
+    Chapters 2 and 3 may also be shortened.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textbf{Reviewer comment:}
-\textit{p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind considerably, although use of the former continually grows due to Cantera?s open availability.''
-Note that FlameMaster is also freely available from the GitLab repository of the developers.}
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind
+    considerably, although use of the former continually grows due to Cantera?s open availability.''
 
-\textbf{Reviewer comment:}
-\textit{p.~2 ``Second, the XML format is intended to be a machine-readable markup language rather than a data format, and its lack of human readibility presents a barrier to creating and working with database files.''
-XML is a text format and it is fully human readable, similarly to YMAL.
-Even a first-time user is able to read the chemical content of a  PrIMe XML file and an experienced user is able to write such a file.}
+    Note that FlameMaster is also freely available from the GitLab repository of the developers.
+    %
+\end{reviewer}
 
-\textit{I suggest adding to following block to this para: ``The original PrIMe was on an Amazon cloud server.
-Recently it was shifted to the computer center of DLR Stuttgart.
-At the new location it is called ``PrIMe 2.0'' [\url{https://prime.cki-know.org/}].
-This is identical to the original PrIMe apart from technical changes due to the different computation background.
-Professor Frenklach plans to develop further PrIMe (using name PrIMe 3.0) and this version will be availabe at [\url{http://primekinetics.org/}].''
-}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textbf{Reviewer comment:}
-\textit{``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for typical bibliographic data.''
+\begin{reviewer}
+    %
+    p.~2 ``Second, the XML format is intended to be a machine-readable markup language rather than a
+    data format, and its lack of human readibility presents a barrier to creating and working with
+    database files.''
+
+    XML is a text format and it is fully human readable, similarly to YMAL. Even a
+    first-time user is able to read the chemical content of a  PrIMe XML file and an experienced
+    user is able to write such a file.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    I suggest adding to following block to this para:
+
+    ``The original PrIMe was on an Amazon cloud server. Recently it was shifted to the computer
+    center of DLR Stuttgart. At the new location it is called ``PrIMe 2.0''
+    [\url{https://prime.cki-know.org/}]. This is identical to the original PrIMe apart from
+    technical changes due to the different computation background. Professor Frenklach plans to
+    develop further PrIMe (using name PrIMe 3.0) and this version will be availabe at
+    [\url{http://primekinetics.org/}].''
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    ``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for
+    typical bibliographic data.''
+
     I suggest the following extension:
-``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for typical bibliographic data, including the DOI of the cited arecicle.
-Each datafile has a unique own DOI and unique URL address.
-[\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}]''.
-Also species can be identified by InCHI or SMILES.
-}
 
-\textbf{Our response:}
+    ``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for
+    typical bibliographic data, including the DOI of the cited arecicle. Each datafile has a unique
+    own DOI and unique URL address.
+    [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}]''. Also species can be
+    identified by InCHI or SMILES.
+    %
+\end{reviewer}
 
-\textbf{Reviewer comment:}
-\textit{``The ReSpecTh standard also provides machine-readable formats for describing ignition experiments, including a field for the definition of autoignition and the ability to specify facility-specific effects.''
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    ``The ReSpecTh standard also provides machine-readable formats for describing ignition
+    experiments, including a field for the definition of autoignition and the ability to specify
+    facility-specific effects.''
+
     Modified text:
-``The ReSpecTh standard provides formats that are both human and machine-readable for describing ignition experiments, including a field for the definition of autoignition and the ability to specify facility-specific effects.
-Also, it describes PSR and flow reactor speciation data, burner stabilized flame speciation data and burning velocity data.''
-}
 
-\textbf{Our response:}
+    ``The ReSpecTh standard provides formats that are both human and machine-readable for describing
+    ignition experiments, including a field for the definition of autoignition and the ability to
+    specify facility-specific effects. Also, it describes PSR and flow reactor speciation data,
+    burner stabilized flame speciation data and burning velocity data.''
+    %
+\end{reviewer}
 
-\textbf{Reviewer comment:}
-\textit{``However, ReSpecTh experimental files are another XML-based format, and, as such, suffer from the same usability issues as PrIMe.''
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    ``However, ReSpecTh experimental files are another XML-based format, and, as such, suffer from
+    the same usability issues as PrIMe.''
+
     I think that the limits of an XML database is set by the actual set of tags and not the format.
-}
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textbf{Reviewer comment:}
-\textit{``Moreover, the standard remains closed, requiring potential users to register with the standard's authors to access the specification and the growing database.''
+\end{response}
+
+\begin{reviewer}
+    %
+    ``Moreover, the standard remains closed, requiring potential users to register with the
+    standard's authors to access the specification and the growing database.''
+
     I suggest instead:
+
     ``Registration is open for everyone who provides a valid institutional e-mail address.''
-}
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textbf{Reviewer comment:}
-\textit{
-``... hydrogen/syngas [11] and ethanol [12]. In all four cases, they converted numerous experimental datasets from the literature into the ReSpecTh standard.''
+\end{response}
+
+\begin{reviewer}
+    %
+    ``... hydrogen/syngas [11] and ethanol [12]. In all four cases, they converted numerous
+    experimental datasets from the literature into the ReSpecTh standard.''
+
     Extended version:
-``...
-hydrogen/syngas [11], methanol and formaldehyde [C.
-Olm, T.
-Varga, E.
-Valko, H.
-J.
-Curran, T.
-Turanyi: Uncertainty quantification of a newly optimized methanol and formaldehyde combustion mechanism, Combustion and Flame, in press] and ethanol [12].
-In all five cases, they converted numerous experimental datasets from the literature into the ReSpecTh standard.
-The database currently contains  93240 experimental data points in 1376 XML-format data files [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}]''.
-}
 
-\textbf{Our response:}
+    ``... hydrogen/syngas [11], methanol and formaldehyde [C. Olm, T. Varga, E. Valko, H. J. Curran,
+    T. Turanyi: Uncertainty quantification of a newly optimized methanol and formaldehyde combustion
+    mechanism, Combustion and Flame, in press] and ethanol [12]. In all five cases, they converted
+    numerous experimental datasets from the literature into the ReSpecTh standard. The database
+    currently contains  93240 experimental data points in 1376 XML-format data files
+    [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}].''
+    %
+\end{reviewer}
 
-\textbf{Reviewer comment:}
-\textit{``type (required, string): How ignition delay was measured;''
-The options described here is a very limited subset of the ignition delay time definitions used in experimental articles.
-For example, backward or forward interpolated values of a measured quantity are further possibilities.
-}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textbf{Reviewer comment:}
-\textit{p.~8 ``In the latter case, some information may be lost, as ChemKED files both encode more details and specify more granularity of some information than ReSpecTh files.''
-This was true for ReSpecTh format 1.0, but no longer true for ReSpecTh Kinetic Dataformat 2.0. The latter covers all info of the ChemKED files (and more).
-}
+\begin{reviewer}
+    %
+    ``type (required, string): How ignition delay was measured;''
 
-\textbf{Our response:}
+    The options described here is a very limited subset of the ignition delay time definitions used
+    in experimental articles. For example, backward or forward interpolated values of a measured
+    quantity are further possibilities.
+    %
+\end{reviewer}
 
-\textbf{Reviewer comment:}
-\textit{Missing data in the references: Year is missing is refs. [5], [6], [9], [10], [11], [12], [21],[25],[27].
-}
+\begin{response}
 
-\textit{[3] Please add reference: \url{https://www.itv.rwth-aachen.de/index.php?id=13}
-Also, the latest version is from 2017.
-}
+\end{response}
 
-\textit{[7] Further info:  Paper P1-04, ISBN 978-963-12-1257-0}
+\begin{reviewer}
+    %
+    p.~8 ``In the latter case, some information may be lost, as ChemKED files both encode more
+    details and specify more granularity of some information than ReSpecTh files.''
 
-\textit{[8] Please refer to ReSpecTh Kinetics Data Format Specification 2.0 instead.}
+    This was true for ReSpecTh format 1.0, but no longer true for ReSpecTh Kinetic Dataformat 2.0.
+    The latter covers all info of the ChemKED files (and more).
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    Missing data in the references: Year is missing is refs. [5], [6], [9], [10], [11], [12],
+    [21],[25],[27].
+    %
+\end{reviewer}
+
+\begin{reviewer}
+    %
+    [3] Please add reference: \url{https://www.itv.rwth-aachen.de/index.php?id=13}  Also, the
+    latest version is from 2017.
+
+    [7] Further info:  Paper P1-04, ISBN 978-963-12-1257-0
+
+    [8] Please refer to ReSpecTh Kinetics Data Format Specification 2.0 instead.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
 
@@
 
-\textit{My main concern with the creation of a publically available database, formatted with ChemKED, is that it seems that the creation of the files has to be done manually, i.e.
-susceptible for errors.
-This is especially true if the experiment comprises a big set of datapoints and/or hundreds of measured species.
-Furthermore, the YAML data serialization format is indeed human-readable, but it does not provide an easy summary/overview of the experimental data, such as an excel or CSV file would give, and therefore, people might be not willing to put the extra effort.}
+\begin{reviewer}
+    %
+    My main concern with the creation of a publically available database, formatted with ChemKED, is
+    that it seems that the creation of the files has to be done manually, i.e. susceptible for
+    errors. This is especially true if the experiment comprises a big set of datapoints and/or
+    hundreds of measured species. Furthermore, the YAML data serialization format is indeed
+    human-readable, but it does not provide an easy summary/overview of the experimental data, such
+    as an excel or CSV file would give, and therefore, people might be not willing to put the extra
+    effort.
+    %
+\end{reviewer}
 
-\textbf{Reviewer comment:}
-\textit{1.~How are the authors planning on ‘forcing’ the experimentalists to publish their data in this new ChemKED file format?}
+\begin{reviewer}
+    %
+    1.~How are the authors planning on `forcing' the experimentalists to publish their data in this
+    new ChemKED file format?
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+    %
+    While we cannot force anyone to use a particular format, we hope that by defining a format and
+    ecosystem with desirable features---including inherent openness and usability---experimentalists
+    will begin adopting a community-standard format. In fact, we have already received interest from
+    various groups, and hope this will grow as we continue to improve and add features. We also hope
+    that at a certain point, with wide adoption, sharing experimental data in a standard format will
+    become the expected norm (e.g., when submitting an article for publication) like sharing a
+    Chemkin-format chemical kinetic model.
+    %
+\end{response}
 
-\textit{2.~Am I correct to state that ChemKED is a YAML file with a predefined structure?}
+\begin{reviewer}
+    %
+    2.~Am I correct to state that ChemKED is a YAML file with a predefined structure?
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+    %
+    Yes, ChemKED files are files in YAML format, with a prescribed structure and set of contents.
+    %
+\end{response}
 
-\textit{3.~I think the article would be more appealing if it would have at least one figure as an example.}
+\begin{reviewer}
+    %
+    3.~I think the article would be more appealing if it would have at least one figure as an
+    example.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+
+\end{response}
 
 \subsection*{Abstract and introduction}
-\textit{4.
-In the abstract, it is mentioned that the Python-based package PyKED creates ChemKED- formatted files.
-However in the article there is no evidence that PyKED aides in creating ChemKED-formatted files.}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    4. In the abstract, it is mentioned that the Python-based package PyKED creates ChemKED-
+    formatted files. However in the article there is no evidence that PyKED aides in creating
+    ChemKED-formatted files.
+    %
+\end{reviewer}
 
-\textit{5.~Page 1, line 31: The future development plans include support for jet stirred reactors, however, this is not mentioned in section 5, Conclusions and future work, nor is there any reference to it on the GitHub wiki roadmap of PyKED.}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textit{6.
-I don't completely agree with the second part of the statement on page 2 line 16-18: ``Second, the XML format is intended to be a machine-readable mark-up language rather than a data format, and its lack of human readability presents a barrier to creating and working with database files.''}
+\begin{reviewer}
+    %
+    5.~Page 1, line 31: The future development plans include support for jet stirred reactors,
+    however, this is not mentioned in section 5, Conclusions and future work, nor is there any
+    reference to it on the GitHub wiki roadmap of PyKED.
+    %
+\end{reviewer}
 
-\textit{In my opinion the YAML format resembles the XML format strongly.
-XML format is human-readable, however, because of the html-tags and `\texttt{<>} characters, the YAML format might be easier to read.
-If XML would lack human readability as it is stated in the article, it would not be such a popular format, and I don't think PrIMe and ReSpecTh would be using it.}
+\begin{response}
 
-\textbf{Our response:}
-We agree with the reviewer that XML is technically human readable by an expert (as opposed to, say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really mean human-\emph{understandable}.
+\end{response}
 
-\textit{7.~An initial version of the ChemKED format was introduced by Niemeyer, as is mentioned in the introduction.
-The mentioned article describes a Python-package called PyTeCK.
-How does this PyTeCK package relate to the PyKED package that is presented in the submitted article?}
+\begin{reviewer}
+    %
+        6. I don't completely agree with the second part of the statement on page 2 line 16-18:
+        ``Second, the XML format is intended to be a machine-readable mark-up language rather than a
+        data format, and its lack of human readability presents a barrier to creating and working
+        with database files.''
 
-\textbf{Our response:}
+        In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
+        however, because of the html-tags and `\texttt{<>} characters, the YAML format might be easier
+        to read. If XML would lack human readability as it is stated in the article, it would not be
+        such a popular format, and I don't think PrIMe and ReSpecTh would be using it.
+    %
+\end{reviewer}
 
-\textit{8.~Page 3, line 10: ``Our motivations are similar to those of '', `to' is missing.}
+\begin{response}
+    %
+    We agree with the reviewer that XML is technically human readable by an expert (as opposed to,
+    say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really
+    mean human-\emph{understandable}.
+    %
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    7.~An initial version of the ChemKED format was introduced by Niemeyer, as is mentioned in the
+    introduction. The mentioned article describes a Python-package called PyTeCK. How does this
+    PyTeCK package relate to the PyKED package that is presented in the submitted article?
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+%
+8.~Page 3, line 10: ``Our motivations are similar to those of '', `to' is missing.
+%
+\end{reviewer}
+
+\begin{response}
+    %
+    We have edited that sentence to correct the error.
+    %
+\end{response}
 
@@
 
-\textit{9.~A reference should not be a required mapping of an experimental dataset.
-If the intention is to make the community to adopt the ChemKED file structure, it might be easier to do so if experimentalists can use the structure internally as well, i.e.
-to read, store, and share non-published internal data.}
+\begin{reviewer}
+    %
+    9.~A reference should not be a required mapping of an experimental dataset. If the intention is
+    to make the community to adopt the ChemKED file structure, it might be easier to do so if
+    experimentalists can use the structure internally as well, i.e. to read, store, and share
+    non-published internal data.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textit{10.~Page 4, line 32: ``The second section of the file encodes the experimental data''.
-Perhaps it is easier to follow if this is briefly mentioned before the authors explain the details of the first section of the ChemKED format.}
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    10.~Page 4, line 32: ``The second section of the file encodes the experimental data''. Perhaps
+    it is easier to follow if this is briefly mentioned before the authors explain the details of
+    the first section of the ChemKED format.
+    %
+\end{reviewer}
 
-\textit{11.~Why do both \texttt{atomic-composition} and \texttt{elemental-composition} exist as synonyms? Wouldn't it be clearer if only one of the two can be used as a valid input?}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textit{12.~Is it possible to create for example two \texttt{common-properties} \texttt{composition} blocks A and B and refer to them in the \texttt{DataPoint} keys?}
+\begin{reviewer}
+    %
+    11.~Why do both \texttt{atomic-composition} and \texttt{elemental-composition} exist as
+    synonyms? Wouldn't it be clearer if only one of the two can be used as a valid input?
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textit{13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED files, on GitHub.
-Is this the most efficient way for making a publically available database?
-One issue with this is the fact that adding files is done via pull requests, which have to be verified and approved by the owners of the repository.
-Therefore a certain delay is inevitable.}
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    12.~Is it possible to create for example two \texttt{common-properties} \texttt{composition}
+    blocks A and B and refer to them in the \texttt{DataPoint} keys?
+    %
+\end{reviewer}
 
-\textit{14.~Page 6, line 18 states that a converter exists from ReSpecTh to ChemKED. But isn't this feature added when going from v0.1.6 to v0.2.0, which makes page 3 line 21: v0.1.6 not correct.}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textit{15.~Page 6, line 21, is the `,' necessary after fork?}
+\begin{reviewer}
+    %
+    13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED
+    files, on GitHub. Is this the most efficient way for making a publically available database? One
+    issue with this is the fact that adding files is done via pull requests, which have to be
+    verified and approved by the owners of the repository. Therefore a certain delay is inevitable.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    14.~Page 6, line 18 states that a converter exists from ReSpecTh to ChemKED. But isn't this
+    feature added when going from v0.1.6 to v0.2.0, which makes page 3 line 21: v0.1.6 not correct.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    15.~Page 6, line 21, is the `,' necessary after fork?
+    %
+\end{reviewer}
+
+\begin{response}
+    %
+    We removed the extraneous comma.
+    %
+\end{response}
 
@@
 
-\textit{16.~The Python package Cerberus is used to validate the format and content of the given ChemKED file.
-How does this package handles/reports the inconsistencies with the schemas?
-As far as my local test went with giving a corrupted YAML file to the ChemKED initiator, the package just spits out a Python error, which might be not the most user-friendly solution.}
+\begin{reviewer}
+    %
+    16.~The Python package Cerberus is used to validate the format and content of the given ChemKED
+    file. How does this package handles/reports the inconsistencies with the schemas? As far as my
+    local test went with giving a corrupted YAML file to the ChemKED initiator, the package just
+    spits out a Python error, which might be not the most user-friendly solution.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textit{17.~A recurring issue with actively developing codes is backwards compatibility. How are the authors planning on addressing this issue?}
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    17.~A recurring issue with actively developing codes is backwards compatibility. How are the
+    authors planning on addressing this issue?
+    %
+\end{reviewer}
 
-\textit{18.~Page 7, line 12, \texttt{mole-percent}, \texttt{mole-fraction}, and \texttt{mass-fraction} include a dash in their name, whereas this is not the case for the example on page 5 line 48 and in the description of the composition element on page 4, line 47.}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textit{19.~Same issue as \#18 holds for page 7, line 17.}
+\begin{reviewer}
+    %
+    18.~Page 7, line 12, \texttt{mole-percent}, \texttt{mole-fraction}, and \texttt{mass-fraction}
+    include a dash in their name, whereas this is not the case for the example on page 5 line 48 and
+    in the description of the composition element on page 4, line 47.
 
-\textbf{Our response:}
+    19.~Same issue as \#18 holds for page 7, line 17.
+    %
+\end{reviewer}
 
-\textit{20.~I think it would be more clear if \texttt{volume\_history}, \texttt{compression\_time}, \texttt{compressed\_temperature},
-and \texttt{compressed\_pressure} would be combined. The line: `if the ChemKED file encodes an
-RCM experiment' is unnecessarily repeated.}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textit{21.~Page 7, line 39: ``Future versions of PyKED may support this feature via, e.g. online lookup''.
-There are two main caveats with online lookups: a) It will make the code slower if the internet
-connection is not strong, b) an internet connection is necessary for the code to be executed.}
+\begin{reviewer}
+    %
+    20.~I think it would be more clear if \texttt{volume\_history}, \texttt{compression\_time},
+    \texttt{compressed\_temperature}, and \texttt{compressed\_pressure} would be combined.
 
-\textbf{Our response:}
+    The line: `if the ChemKED file encodes an RCM experiment' is unnecessarily repeated.
+    %
+\end{reviewer}
 
-\textit{22.~Page 7, line 40: Why is it interesting to create the YAML file again with the \texttt{write\_file()} function.
-Creating a summary excel/CSV file would be more useful.}
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    21.~Page 7, line 39: ``Future versions of PyKED may support this feature via, e.g. online
+    lookup''. There are two main caveats with online lookups: a) It will make the code slower if the
+    internet connection is not strong, b) an internet connection is necessary for the code to be
+    executed.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    22.~Page 7, line 40: Why is it interesting to create the YAML file again with the
+    \texttt{write\_file()} function. Creating a summary excel/CSV file would be more useful.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
 
@@
 
-\textit{23.~Page 10, line 29--30: ``A user might want to load the information'' Avoid the term `might'.
-It's one of the main goals of using ChemKED and its associated Python-package PyKED to load the information necessary for simulating the experiment.}
+\begin{reviewer}
+    %
+    23.~Page 10, line 29--30: ``A user might want to load the information'' Avoid the term `might'.
+    It's one of the main goals of using ChemKED and its associated Python-package PyKED to load the
+    information necessary for simulating the experiment.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textit{24.~The Python code snippets are not always initiated in the text.}
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    24.~The Python code snippets are not always initiated in the text.
+    %
+\end{reviewer}
 
-\textit{25.~Page 10, line 55: add more commentary text. It is not clear that those lines create the
-necessary instances for Cantera, rather than ChemKED.}
+\begin{response}
 
-\textbf{Our response:}
+\end{response}
 
-\textit{26.~Page 11, line 7: Where is this `air.xml' file coming from. This is not clear from the text.}
+\begin{reviewer}
+    %
+    25.~Page 10, line 55: add more commentary text. It is not clear that those lines create the
+    necessary instances for Cantera, rather than ChemKED.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+
+\end{response}
+
+\begin{reviewer}
+    %
+    26.~Page 11, line 7: Where is this `air.xml' file coming from. This is not clear from the text.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
 
@@
 
-\textit{27.~The table overlapping page 11 and 12 can be reduced by truncating the \texttt{DataPoints}, similarly as is done with the table starting at page 9, line 31 and ending at page 10, line 27.}
+\begin{reviewer}
+    %
+    27.~The table overlapping page 11 and 12 can be reduced by truncating the \texttt{DataPoints},
+    similarly as is done with the table starting at page 9, line 31 and ending at page 10, line 27.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
 
-\textit{28.
-Python code snippet on page 13: (a.) Add more commentary text.
-(b.) Changing the example to enable multithreading might make it more appealing to start using ChemKED as a data format in addition with Cantera as a simulation tool.}
+\end{response}
 
-\textbf{Our response:}
+\begin{reviewer}
+    %
+    28. Python code snippet on page 13: (a.) Add more commentary text. (b.) Changing the example to
+    enable multithreading might make it more appealing to start using ChemKED as a data format in
+    addition with Cantera as a simulation tool.
+    %
+\end{reviewer}
+
+\begin{response}
+
+\end{response}
 
@@
 
-\textit{29.~No conclusion is mentioned for the usage examples.}
+\begin{reviewer}
+    %
+    29.~No conclusion is mentioned for the usage examples.
+    %
+\end{reviewer}
 
-\textbf{Our response:}
+\begin{response}
+
+\end{response}
 

### 2017-09-27 rebuttal.tex
Fix compilation errors in the response
added rebuttal text to address #16
Additional ReSpecTh references to address #13

@@
 \newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup\vspace{0.5\baselineskip}}
-\newenvironment{response}{\vspace{0.5\baselineskip}\begin{response}
-
-\end{response}}{\vspace{0.5\baselineskip}}
+\newenvironment{response}{\vspace{0.5\baselineskip}\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
 \newcommand{\review}[1]{{\itshape#1}}
-\newcommand{\respond}[1]{\begin{response}
-
-\end{response} #1}
+\newcommand{\respond}[1]{\textbf{Our Response:} #1}
 
@@
 \begin{response}
-
+    We added the suggested text and reference.
 \end{response}
@@
     %
-    Missing data in the references: Year is missing is refs. [5], [6], [9], [10], [11], [12],
+    Missing data in the references: Year is missing is refs.~[5], [6], [9], [10], [11], [12],
     [21],[25],[27].
-    %
-\end{reviewer}
 
-\begin{reviewer}
-    %
     [3] Please add reference: \url{https://www.itv.rwth-aachen.de/index.php?id=13}  Also, the
@@
 \begin{response}
-
+    We corrected those references; thank you for pointing out the missing\slash old information.
 \end{response}

### 2017-09-27 pyked.tex
Additional ReSpecTh references to address #13

@@
 syngas~\autocite{Varga:2016gj}, methanol\slash formaldehyde~\autocite{Olm:2017a},
-and ethanol~\autocite{Olm:2016et}. In all four
-cases, they converted numerous experimental datasets from the literature into
-the ReSpecTh standard.
+and ethanol~\autocite{Olm:2016et}. In all five cases, they converted numerous
+experimental datasets from the literature into the ReSpecTh standard.
+Their database currently contains \num{93240} experimental data points in 1376
+ReSpecTh XML files~\autocite{respecth}.
 

### 2017-09-28 rebuttal.tex
added more text for #18
add Bryan's suggested text and rebuttal for #12

@@
 \begin{response}
-
+    We have updated the text to reflect this information; however, we argue that requiring 
+    registration limits use and differs from the openness ``baked-in'' to our approach.
 \end{response}
@@
     Chemkin-format chemical kinetic model.
+
+    Furthermore, we are willing to work (and currently are) with experimentalists to provide them easy-to-use
+    conversion scripts from internal data formats, and they will help us define what needs to go in the
+    specification for their experiment.
     %

### 2017-09-28 pyked.tex
add Bryan's suggested text and rebuttal for #12

@@
 XML-based format, and, as such, suffer from the same usability issues as PrIMe.
-Moreover, the standard remains closed, requiring potential users to register
-with the standard's authors to access the specification and the growing database.
+Moreover, the standard remains inaccessible to the public, requiring potential users
+to register with the standard's authors using a valid institutional email address
+to access the specification and the growing database.
 

### 2017-09-28 rebuttal.tex
draft response for #11
modified article and rebuttal for #10

@@
 \begin{response}
-
+    We have incorporated most of the recommended text in our revision. However, we did not add
+    ``human-readable'', consistent with the definition used throughout (i.e., less verbose and easily understandable).
 \end{response}
@@
 \begin{response}
-
+    By ``usability'' of a format, we mean whether files are easily usable and understandable by people,
+    rather than the technical limitations of a format.
 \end{response}

### 2017-09-28 rebuttal.tex
rebuttal for #6

@@
     p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind
-    considerably, although use of the former continually grows due to Cantera?s open availability.''
+    considerably, although use of the former continually grows due to Cantera's open availability.''
 
@@
 \begin{response}
-
+    We attempted to access the GitLab repository at \url{https://git.rwth-aachen.de/ITV/FlameMaster}, but
+    it is a closed repository. Thus, while it may be freely available to those with access, it is not
+    \emph{openly} available.
 \end{response}

### 2017-09-28 pyked.tex
modified article and rebuttal for #10
change listings stuff in paper and class, add instructions to README

--- 2017-09-27 pyked.tex
@@
 % !TEX option = -shell-escape
+%% comment out the above line before submitting
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
+More recently, Varga and coworkers~\autocite{Varga2015a,Varga2015b} developed the
 ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
@@
 and more recently develop improved models for hydrogen\slash
-syngas~\autocite{Varga:2016gj}, methanol\slash formaldehyde~\autocite{Olm:2017a},
-and ethanol~\autocite{Olm:2016et}. In all four
+syngas~\autocite{Varga:2016gj} and ethanol~\autocite{Olm:2016et}. In all four
 cases, they converted numerous experimental datasets from the literature into
@@
 software tool for validating and working with \ck{} files: \pk{}. Our
-motivations resemble those of the PrIMe and ReSpecTh teams: the \ck{} project
+motivations are similar those of the PrIMe and ReSpecTh teams: the \ck{} project
 will enable easy sharing and use of fundamental combustion data, for the
@@
 the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
-currently provides converters to and from the ReSpecTh format~\autocite{Varga2017}.
+currently provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
 While we plan to continue growing this database---and hope for submissions from
-the community---others can freely download files, or in fact copy (``fork'' in GitHub
+the community---others can freely download files, or in fact copy (``fork,'' in GitHub
 parlance) the entire database for their own purposes.

### 2017-09-29 rebuttal.tex
temporary response to #30, not done
Work on #5, justify a longer text rather than shorter

@@
 \begin{response}
-
+    %
+    Our intention is to describe the system to an audience that may not be familiar with the
+    technical details of how the system is implemented. Thus, we feel that the length of the
+    manuscript is a necessary outcome of providing the background for this information. However, the
+    reviewer is correct that it is not ideal to leave much of the information in the documentation
+    only. In combination with the suggestions of Reviewer 2, we have expanded the paper to provide
+    more detail where necessary.
+    %
 \end{response}
@@
 \begin{response}
-
+    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some delay,
+    however
 \end{response}

### 2017-09-29 rebuttal.tex
Work on #7, #23. Clarify that the problem with XML is verbosity

@@
 \begin{response}
+    %
+    We certainly recognize that XML is typically written in ASCII or UTF-8 (plain text) files, which
+    a human can hypothetically decipher. Our argument is that XML is significantly harder for a
+    human to read and write than YAML. This is primarily because XML has a lower signal-to-noise
+    ratio than YAML; it requires many more characters to express a data structure in XML than in
+    YAML. For instance, XML requires opening and closing tags, with the associated angle brackets,
+    for every field while YAML does not; in XML, information can be specified as properties of tags
+    or as elements (requiring the user to know where to look for information) while YAML has only 3 basic data structures (keys, values, and lists).
 
+    We note that although this verbosity in XML is a disadvantage when humans attempt to read or
+    write such files, it allows much greater flexibility in specifying how a document should be
+    marked up and provides XML with many advantages in other areas. However, since YAML is intended
+    as a simple data format, and has a much simpler overall format, we argue that it is therefore
+    much easier for humans to work with.
+
+    We have modified the text to clarify this point, including describing the lack of human
+    readability as a formatting issue (but one that is inherent to XML), rather than a technical
+    issue.
+    %
 \end{response}

### 2017-09-29 pyked.tex
Work on #7, #23. Clarify that the problem with XML is verbosity

--- 2017-09-28 pyked.tex
@@
 effects necessary to properly simulate certain experiments. Second, the XML
-format is intended to be a machine-readable markup language rather than a data format, and
-its lack of human readibility presents a barrier to creating and working with
-database files. Furthermore, the
-focus on using internal identifiers for bibliographic references and
+format is intended to be a machine-readable markup language rather than a data format.
+As a general-purpose document markup language, XML is able to represent very detailed
+information about a dataset, but implementing this detail requires XML-formatted
+files to be significantly more verbose than files formatted with a language focused
+only on representing data structures. The verbosity inherent in the XML format
+limits the human readability of database files constructed in XML and presents a barrier to
+creating and working with them. Furthermore, PrIMe uses
+internal identifiers for bibliographic references and
 species\slash reactions, rather than standard identifiers (e.g., DOIs for
-scholarly products, InCHI or SMILES for species), limits usability outside
+scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
 the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
@@
 a field for the definition of autoignition and the ability to specify
-facility-specific effects. However, ReSpecTh experimental files are another
+facility-specific effects. It can also describe speciation measurements in
+perfectly stirred reactors, flow reactors, and burner-stabilized flames, as well as
+burning velocity data. However, ReSpecTh experimental files are another
 XML-based format, and, as such, suffer from the same usability issues as PrIMe.

### 2017-09-29 rebuttal.tex
Reply to #8, add new PrIMe url to text, but don't add other suggestions

@@
 \begin{response}
-
+    %
+    We do not think that the technical details of the implementation of the original PrIMe system
+    are relevant to this work, so we have not included them. In addition, we cannot know
+    the plans of Prof. Frenklach, and we do not feel it would be appropriate to comment on them in
+    our manuscript. As such, we have not added these suggestions to the paper. We have, however,
+    added the new URL to the current manuscript, and thank the reviewer for pointing us to that.
+    %
 \end{response}

### 2017-09-29 pyked.tex
Reply to #8, add new PrIMe url to text, but don't add other suggestions

@@
 
-PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/}
+has a number of features that make
 it a strong standardization format, but it suffers from several flaws that

### 2017-09-29 pyked.tex
Work on #9, add modified version of suggested text to manuscript

@@
 system. For example, where PrIMe uses internal bibliographic references,
-ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
+ReSpecTh adds a field for typical bibliographic data, including the DOI of
+the cited article. The ReSpecTh standard also
 provides machine-readable formats for describing ignition experiments, including
@@
 perfectly stirred reactors, flow reactors, and burner-stabilized flames, as well as
-burning velocity data. However, ReSpecTh experimental files are another
+burning velocity data. To support this, ReSpecTh files allow the specification of species by
+standard identifiers, including InChI or SMILES. Finally, each ReSpecTh file
+has a unique DOI and URL assigned to it.
+
+However, ReSpecTh experimental files are another
 XML-based format, and, as such, suffer from the same usability issues as PrIMe.

### 2017-09-29 rebuttal.tex
Work on #9, add modified version of suggested text to manuscript
Slightly modify response for #10. Closes #10.

@@
 \begin{response}
-
+    %
+    We have incorporated these modifications into the suggested paragraph, although we have modified
+    the format slightly to better fit the flow of the text.
+    %
 \end{response}
@@
 \begin{response}
+    %
     We have incorporated most of the recommended text in our revision. However, we did not add
-    ``human-readable'', consistent with the definition used throughout (i.e., less verbose and easily understandable).
+    ``human-readable'', consistent with the definition used throughout (i.e., the XML file is more
+    verbose and less easily understood).
+    %
 \end{response}

### 2017-09-29 rebuttal.tex
Update response for #11. Change usability to readability to clarify

@@
 \begin{response}
-    By ``usability'' of a format, we mean whether files are easily usable and understandable by people,
-    rather than the technical limitations of a format.
+    %
+    By ``usability'' of a format, we mean whether files are easily usable and understandable by
+    people, rather than the technical limitations of a format. We have changed the word
+    ``usability'' to ``readability'' to emphasize this.
+    %
 \end{response}

### 2017-09-29 rebuttal.tex
Slightly modify response for #12

@@
 \begin{response}
+    %
     We have updated the text to reflect this information; however, we argue that requiring
-    registration limits use and differs from the openness ``baked-in'' to our approach.
+    registration to view the specification and data limits the possibility of use of the standard
+    and differs from the openness that we are advocating, where the specification and data files are
+    open for anyone to view, with or without registration.
+    %
 \end{response}

### 2017-09-30 rebuttal.tex
Work on #14.
I made a PR (pr-omethe-us/PyKED#77) to close the related issue
(pr-omethe-us/PyKED#71). This should address the reviewer's comment.

@@
 \begin{response}
-
+    %
+    We have implemented an option for baseline extrapolation of the slope. We have added the
+    description of this ignition type to the mentioned portion of the manuscript. To our knowledge,
+    the options now supported represent all of the common methods of specifying the ignition delay
+    in rapid compression machine and shock tube experiments.
+    %
 \end{response}

### 2017-09-30 pyked.tex
Update response for #11. Change usability to readability to clarify
Work on #14.
I made a PR (pr-omethe-us/PyKED#77) to close the related issue
(pr-omethe-us/PyKED#71). This should address the reviewer's comment.

@@
 However, ReSpecTh experimental files are another
-XML-based format, and, as such, suffer from the same usability issues as PrIMe.
+XML-based format, and, as such, suffer from the same readability issues as PrIMe.
 Moreover, the standard remains inaccessible to the public, requiring potential users
@@
         \yabox|max| or \yabox|min| (indicates the ignition point was at the
-        maximum or minimum of the \yabox|target|), or \yabox|1/2 max| (
-        indicates the half-maximum point of \yabox|target|)
+        maximum or minimum of the \yabox|target|), \yabox|1/2 max| (
+        indicates the half-maximum point of \yabox|target|), or \yabox|d/dt max extrapolated|
+        (indicates the maximum slope of the target extrapolated to the baseline)
         \item \yabox|target| (required, string): The target for the ignition

### 2017-09-30 rebuttal.tex
Work on #15.
Modify the sentence to indicate incompatibility runs both ways

@@
 \begin{response}
-
+    %
+    This is still the case, even in ReSpecTh format 2.0. For instance, ChemKED files support
+    specifying the initial composition as a mass fraction, which ReSpecTh does not support. We have
+    modified the sentence to indicate that there are some incompatibilities in both directions, not
+    all ChemKED files can be converted to ReSpecTh files, and not all ReSpecTh files can be
+    converted to ChemKED files due to missing data elements in both directions.
+    %
 \end{response}

### 2017-09-30 pyked.tex
Work on #15.
Modify the sentence to indicate incompatibility runs both ways

@@
 of the class, while \pybox|convert_to_ReSpecTh()| converts the \pybox|ChemKED|
-instance to a ReSpecTh XML file. In the latter case, some information may be lost, as \ck{}
-files both encode more details and specify more granularity of some information than
-ReSpecTh files.
+instance to a ReSpecTh XML file. In the latter case, some information may be lost or the conversion
+may not be possible, as \ck{} files support some data that ReSpecTh files do not support. We note
+that the reverse is also true, not all ReSpecTh files may be converted to \ck{} files because not
+all of the data that can be stored in a ReSpecTh format is supported by the \ck{} format.
 

### 2017-09-30 rebuttal.tex
Respond to #17 clarifying that other formats have the same disadvantages

@@
 \begin{response}
+    %
     We corrected those references; thank you for pointing out the missing\slash old information.
+    %
 \end{response}
@@
 \end{reviewer}
+
+\begin{response}
+    %
+    We agree with the reviewer, insofar as manual creation of the files may lead to errors in the
+    encoded data. However, this will be the case wherever manual processing is required, regardless
+    of the database format, and could even be the case for tables or figures published in journal
+    articles. Thus, we do not feel that manual file creation is an inherent disadvantage of our
+    format. Moreover, as we discuss below, we are working with experimental teams to integrate the
+    creation of ChemKED files into their workflows to try to mitigate some of these problems.
+
+    On the second point, we agree with the reviewer that CSV or Excel files may give an adequate
+    summary or overview of the data, but only if the files are opened with some appropriate
+    software, e.g., Microsoft Excel (particularly for Excel-formatted files). Thus, in this sense,
+    CSV or Excel files do not offer any advantage over ChemKED files (or any other database files)
+    that must also be processed with an appropriate software package (PyKED, in the case of ChemKED
+    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED,
+    it is trivial to produce an Excel or CSV file so that one may view such a summary if desired.
+    See also the response to a similar comment below.
+    %
+\end{response}
 

### 2017-09-30 rebuttal.tex
Slightly modify text for #18

@@
     ecosystem with desirable features---including inherent openness and usability---experimentalists
-    will begin adopting a community-standard format. In fact, we have already received interest from
-    various groups, and hope this will grow as we continue to improve and add features. We also hope
-    that at a certain point, with wide adoption, sharing experimental data in a standard format will
-    become the expected norm (e.g., when submitting an article for publication) like sharing a
-    Chemkin-format chemical kinetic model.
+    will begin adopting a community-standard format. In fact, we have already received expressions
+    of interest from various experimental groups, and hope this will grow as we continue to improve
+    and add features. We also hope that at a certain point, with wide adoption, sharing experimental
+    data in a standard format will become the expected norm (e.g., when submitting an article for
+    publication) like sharing a Chemkin-format chemical kinetic model.
 
-    Furthermore, we are willing to work (and currently are) with experimentalists to provide them easy-to-use
-    conversion scripts from internal data formats, and they will help us define what needs to go in the
-    specification for their experiment.
+    Furthermore, we are willing to work (and are currently working) directly with experimentalists
+    to provide them easy-to-use conversion scripts from internal data formats. In return, they are
+    helping us define what data and information needs to go in the specification for their
+    experiment.
     %

### 2017-09-30 rebuttal.tex
Slightly modify response for #19
Respond to #21
Point out the ongoing GUI work to create ChemKED files, and clarify that
PyKED is for validating and working with CK files

@@
     Yes, ChemKED files are files in YAML format, with a prescribed structure and set of contents.
+    The required structure and contents are defined by our schema, which is documented in this
+    manuscript and in the online documentation.
     %
@@
 \begin{response}
+    %
+    The reviewer is correct, and this was a typo in the abstract. We have clarified that PyKED is
+    intended as software to validate and work with ChemKED files.
 
+    That said, we do have ongoing work to create a graphical user interface to create ChemKED files,
+    as this was requested by several of the experimental teams that have expressed interest in the
+    ChemKED format. The work is available publicly at
+    \url{https://github.com/pr-omethe-us/ChemKED-gui}.
+    %
 \end{response}

### 2017-09-30 pyked.tex
Respond to #21
Point out the ongoing GUI work to create ChemKED files, and clarify that
PyKED is for validating and working with CK files

@@
 \begin{abstract} % not to exceed 200 words
-Fundamental experimental measurements of quantities such as ignition delay times,
-laminar flame speeds, and species profiles (among others) serve important roles in understanding
-fuel chemistry and validating chemical kinetic models. However, despite both the
-importance and abundance of such information in the literature, the community lacks
-a widely adopted standard format for this data. This impedes both sharing and wide
-use by the community. Here we introduce a new chemical kinetics
-experimental data format, ChemKED, and the related Python-based package for creating
-and validating ChemKED-formatted files called PyKED. We also review past and
-related efforts, and motivate the need for a new solution. ChemKED currently supports
-the representation of autoignition delay time measurements from shock tubes and rapid
-compression machines.
-ChemKED-formatted files contain all of the information needed to simulate experimental
+    %
+    Fundamental experimental measurements of quantities such as ignition delay times, laminar flame
+    speeds, and species profiles (among others) serve important roles in understanding fuel
+    chemistry and validating chemical kinetic models. However, despite both the importance and
+    abundance of such information in the literature, the community lacks a widely adopted standard
+    format for this data. This impedes both sharing and wide use by the community. Here we introduce
+    a new chemical kinetics experimental data format, ChemKED, and the related Python-based package
+    for validating and working with ChemKED-formatted files called PyKED. We also review past and
+    related efforts, and motivate the need for a new solution. ChemKED currently supports the
+    representation of autoignition delay time measurements from shock tubes and rapid compression
+    machines. ChemKED-formatted files contain all of the information needed to simulate experimental
     data points, including the uncertainty of the data. ChemKED is based on the YAML data
-serialization language, and is intended as a human- and machine-readable standard for
-easy creation and automated use.
-Development of ChemKED and PyKED occurs openly on GitHub under the
-BSD 3-clause license, and contributions from the community are welcome. Plans for
-future development include support for experimental data from laminar flame,
-jet stirred reactor, and speciation measurements.
+    serialization language, and is intended as a human- and machine-readable standard for easy
+    creation and automated use. Development of ChemKED and PyKED occurs openly on GitHub under the
+    BSD 3-clause license, and contributions from the community are welcome. Plans for future
+    development include support for experimental data from laminar flame, jet stirred reactor, and
+    speciation measurements.
+    %
 \end{abstract}

### 2017-09-30 pyked.tex
Respond to #22
Add JSR and flow reactor experiments to the conclusion, and update the
Roadmap

@@
 
-\ck{} and \pk{} are currently under active development, and contributions from
-the community are welcomed. All development occurs under the BSD 3-clause open source
-license and the code is housed on GitHub at
-\url{https://github.com/pr-omethe-us/PyKED}. Future directions for development
-are outlined in the public
-roadmap.\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}
-The highest priority issues are currently adding support for other types of
-fundamental experiments, including speciation in shock tubes, laminar flame speeds,
-and flame extinction measurements.
-Questions, comments, or suggestions are welcomed and can be posted as issues in the
-GitHub repository or emailed to the authors. In particular, we welcome proposals for new
-experimental types; see issue \#60 in the \pk{} repository for an example of how this may
-be done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
+\ck{} and \pk{} are currently under active development, and contributions from the community are
+welcomed. All development occurs under the BSD 3-clause open source license and the code is housed
+on GitHub at \url{https://github.com/pr-omethe-us/PyKED}. Future directions for development are
+outlined in the public roadmap.\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}
+The highest priority issues are currently adding support for other types of fundamental experiments,
+including speciation measurements in shock tubes, jet-stirred reactors, and flow reactors, and
+laminar flame speed and flame extinction measurements. Questions, comments, or suggestions are
+welcomed and can be posted as issues in the GitHub repository or emailed to the authors. In
+particular, we welcome proposals for new experimental types; see issue \#60 in the \pk{} repository
+for an example of how this may be
+done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
 

### 2017-09-30 rebuttal.tex
Respond to #22
Add JSR and flow reactor experiments to the conclusion, and update the
Roadmap
Slightly modify response for #23

@@
 \begin{response}
-
+    %
+    Thank you for the comment, we have corrected this oversight in the roadmap and the conclusions.
+    %
 \end{response}
@@
     ``Second, the XML format is intended to be a machine-readable mark-up language rather than a
-        data format, and its lack of human readability presents a barrier to creating and working
-        with database files.''
+    data format, and its lack of human readability presents a barrier to creating and working with
+    database files.''
 
     In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
-        however, because of the html-tags and `\texttt{<>} characters, the YAML format might be easier
-        to read. If XML would lack human readability as it is stated in the article, it would not be
-        such a popular format, and I don't think PrIMe and ReSpecTh would be using it.
+    however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to
+    read. If XML would lack human readability as it is stated in the article, it would not be such a
+    popular format, and I don't think PrIMe and ReSpecTh would be using it.
     %
@@
     say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really
-    mean human-\emph{understandable}.
+    mean human-\emph{understandable}. We have modified the text to clarify this point; please see
+    also the response to a similar comment from Reviewer \#1.
     %

### 2017-09-30 rebuttal.tex
Respond to #26

@@
 \begin{response}
+    %
+    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
+    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
+    users who wish to use ChemKED files in this way include an incomplete reference field since the
+    only required subfields of the reference field are the authors and the year.
 
+    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor.
+    This option turns off all of the validation, including the validation of the reference material
+    and enforcement of required fields. We recognize that this is a limitation at present, if users
+    wish to, for example, validate that their units are correct but do not wish to validate some
+    other aspect of the file. Work is ongoing to improve this situation, and discussion is contained
+    in a public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
+    %
 \end{response}

### 2017-09-30 rebuttal.tex
Respond to #30

@@
 \begin{response}
-    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some delay,
-    however
+    %
+    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some
+    delay, however we see the human intervention step as necessary and important. In much the same
+    way that review of manuscripts provides a certain surety of the minimal quality of the article,
+    we see review of data being added to the database as a minimal quality check; for instance, does
+    the reference exist, has the data already been added to the database, are the files styled
+    consistently, etc. Furthermore, such human review can help to catch some errors that may be
+    introduced by manual creation of the files, as pointed out by the reviewer previously. Finally,
+    we would note that all of the existing database have a similar requirement, where users submit
+    data files for inclusion that are approved by the database owners after some review. In this
+    sense, we actually see the Pull Request mechanism of GitHub as an advantage, since all reviews
+    can be done publicly and some of the review procedures (style checks, reference validity, etc)
+    may be done automatically by a script associated with the GitHub repository. In this way, the
+    entire community can contribute to improving the quality of the database.
+    %
 \end{response}

### 2017-09-30 rebuttal.tex
Draft response to #31

@@
 \begin{response}
-
+    %
+    In principle, the version of the ChemKED database format and the version of the PyKED Python
+    package are entirely independent. Thus, since adding the converters incurred no changes to the
+    database format, the description of the format in the manuscript is a description of v0.1.6,
+    which happens to be identical to v0.2.0. However, in practice, the versions of PyKED and the
+    ChemKED database have been coupled up to now, and we have updated the version mentioned in the
+    text to vX.X.X, which we released in conjunction with submitting this response after the helpful
+    comments of the reviewers.
+    %
 \end{response}

### 2017-09-30 rebuttal.tex
Respond to #31

@@
 \begin{response}
-
+    %
+    We recognize that this is a limitation of the current version of PyKED. The technical reason for
+    this in the Cerberus package has very recently been resolved (see
+    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error reporting
+    mechanism in PyKED to report well-formatted errors that have a detailed description of the
+    error. We have opened a public issue to track the progress on this:
+    \url{https://github.com/pr-omethe-us/PyKED/issues/78}. Thank you for suggesting this
+    improvement.
+    %
 \end{response}

### 2017-10-01 rebuttal.tex
Fix response to #26
Turns out the journal was also required, so make a note that we've
dropped that requirement. Also, fix the text where some optional fields
were specified as required.

@@
     the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
-    users who wish to use ChemKED files in this way include an incomplete reference field since the
-    only required subfields of the reference field are the authors and the year.
+    users who wish to use ChemKED files in this way include an incomplete reference field. We have
+    modified the ChemKED schema in response to this comment so that the only required subfields of
+    the reference field are the authors and the year.
 

### 2017-10-01 rebuttal.tex
Respond to #34

@@
 \begin{response}
+    %
+    We agree with the reviewer that backwards compatibility may become an issue. We have addressed
+    this concern in ChemKED by including the database version as one of the required keys in the
+    ChemKED file. Although the functionality doesn't exist yet, we intend to be able to check this
+    version and report any feature deprecation or removal to the user (see
+    \url{https://github.com/pr-omethe-us/PyKED/issues/22), and we plan to provide an automated
+    upgrade path for ChemKED files when backwards-incompatible changes are made. In addition, we use
+    the widely supported Semantic Versioning paradigm, where changes in the various parts of the
+    version number (vX.Y.Z) have specific meaning; changing the major version (X) indicates a
+    backwards incompatible change, changing the minor version (Y) indicates a backwards compatible,
+    but somewhat large, change, and changing the patch version (Z) indicates a bug fix with no
+    anticipated incompatibilities.
 
+    Finally, each released version of PyKED is permanently archived on Zenodo
+    (\url{https://doi.org/10.5281/zenodo.597935}) so that it is accessible should a user wish to use
+    an old version to solve a compatibility issue temporarily.
+    %
 \end{response}

### 2017-10-02 pyked.tex
Fix response to #26
Turns out the journal was also required, so make a note that we've
dropped that requirement. Also, fix the text where some optional fields
were specified as required.
Respond to #35
Correct the typo in the manuscript regarding hyphens in composition type
names

@@
         mappings
-        \item \yabox|journal| (required, string): The name of the publishing
+        \item \yabox|journal| (optional, string): The name of the publishing
         journal
@@
         \item \yabox|doi| (optional, string): The article DOI
-        \item \yabox|volume| (required, integer): The journal volume number
+        \item \yabox|volume| (optional, integer): The journal volume number
         \item \yabox|pages| (optional, string): The article pages
@@
     \item \pybox|composition|: A list of dictionaries of the species and their
-    respective amounts. The values are validated so that \yabox|mole-percent|,
-    \yabox|mole-fraction|, or \yabox|mass-fraction| cannot be mixed for a single
+    respective amounts. The values are validated so that \yabox|mole percent|,
+    \yabox|mole fraction|, or \yabox|mass fraction| cannot be mixed for a single
     data point, and so that the sum of the values is approximately 1.0, or
-    100.0 for \yabox|mole-percent|.
+    100.0 for \yabox|mole percent|.
     \item \pybox|composition_type|: A string indicating the type of composition
-    information for the data point---one of \pybox|'mole-percent'|,
-    \pybox|'mole-fraction'|, or \pybox|'mass-fraction'|
+    information for the data point---one of \pybox|'mole percent'|,
+    \pybox|'mole fraction'|, or \pybox|'mass fraction'|
     \item \pybox|volume_history|: If the \yabox|volume-history| of an RCM

### 2017-10-02 rebuttal.tex
Respond to #35
Correct the typo in the manuscript regarding hyphens in composition type
names
Respond to #36
Consolidate the RCM-specific attributes into one list.

@@
 \begin{response}
-
+    %
+    This was a typo in the manuscript. All of the keywords here include spaces in their name, both
+    in the YAML schema and in the PyKED interface. This has been corrected in the manuscript and we
+    have added a test to the PyKED package to ensure the values are correct.
+    %
 \end{response}
@@
 \begin{response}
-
+    %
+    We have consolidated the RCM-specific information into a separate list, and eliminated the
+    duplicated phrase.
+    %
 \end{response}

### 2017-10-02 pyked.tex
Respond to #36
Consolidate the RCM-specific attributes into one list.

@@
     \pybox|'mole fraction'|, or \pybox|'mass fraction'|
+    \item \pybox|ignition_type|: A dictionary specifying the method of the
+    measurement of ignition delay
+\end{itemize}
+
+If the \ck{} file encodes an RCM experiment, there are several specialized fields that can be
+included.
+
+\begin{itemize}
     \item \pybox|volume_history|: If the \yabox|volume-history| of an RCM
@@
     \autocite{vanderWalt:2011np}
-    \item \pybox|compression_time|: If the \ck{} file encodes an RCM experiment,
-    the length of the compression stroke may be included
-    \item \pybox|compressed_temperature|: If the \ck{} file encodes an RCM experiment,
-    the temperature at the end of compression may be included
-    \item \pybox|compressed_pressure|: If the \ck{} file encodes an RCM experiment,
-    the pressure at the end of compression may be included
-    \item \pybox|ignition_type|: A dictionary specifying the method of the
-    measurement of ignition delay
+    \item \pybox|compression_time|: The duration of the compression stroke
+    \item \pybox|compressed_temperature|: The estimated reactor temperature at the end of the
+    compression stroke
+    \item \pybox|compressed_pressure|: The measured reactor pressure at the end of the compression
+    stroke
 \end{itemize}
@@
 
-Only the first author is printed when \pybox|Reference| or
+Only the first author is included in the \pybox|DataFrame| when \pybox|Reference| or
 \pybox|Reference:Authors| is selected because the whole author list may be

### 2017-10-02 rebuttal.tex
WIP response to #37

@@
 \begin{response}
-
+    %
+    In general, both of these points are correct. However, we believe there are several mitigating
+    factors that make the online lookups worthwhile. First, we explicitly test that our code fails
+    gracefully in the absence of an internet connection. No internet connection is necessary to run
+    the code at this time, and aside from specialized features, we do not anticipate that an
+    internet connection will ever be required to run the basic software. Second, for the specific
+    feature mentioned here, this will be something that the user is explicitly requesting (e.g., by
+    attempting to convert mass to mole fractions). Thus, we expect that they are willing to wait for
+    the connection to complete to use the feature. Third, we expect that the amount of data
+    transmitted for such a request would be small, on the order of a few kB, so that even slow
+    internet connections would be able to handle such requests. Finally, in addition to
+    internet-based options, we are also investigating the use of standalone packages, such as RDKit,
+    that would enable local retrieval of such information. We have not modified the text in response
+    to the comment.
+    %
 \end{response}

### 2017-10-02 rebuttal.tex
WIP response to #38
In general, the requested functionality is available in Pandas, so I
would prefer to defer to that

@@
 \begin{response}
+    %
+    The use case for the \texttt{write\_file()} is that users might modify the ChemKED instance
+    after it has been loaded, and wish to write their changes out to a new file.
 
+    Users can create Excel or CSV files containing all or some of the datapoints in the ChemKED
+    instance by using the \verb|get_dataframe()| function, and the associated functionality in
+    Pandas (\verb|DataFrame.to_csv()| and \verb|DataFrame.to_excel()|). Since this functionality
+    already exists, we do not wish to reinvent the wheel by writing our own functions that would
+    have to be maintained and tested. Finally, we would note that data can also be plotted and
+    viewed directly from Python using existing functionality in the Pandas DataFrame class
+    (\verb|DataFrame.plot()|).
+    %
 \end{response}

### 2017-10-02 pyked.tex
WIP respond to #39
Modify the text to remove "might"

@@
 \textcite{Mittal2006a}:
-%
+%r
 \begin{yamlbox}
@@
 %
-The values for the \yabox|volume-history| are truncated here to save space. A user
-might want to load the information from this file to perform a simulation using
-Cantera~\autocite{Cantera:2.3.0} to calculate the ignition delay. All information
-required to perform this simulation is present in the \ck{} file, with the exception
-of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
+The values for the \yabox|volume-history| are truncated here to save space. One application of the
+data stored in this file is to perform a simulation using Cantera~\autocite{Cantera:2.3.0} to
+calculate the ignition delay, including the facility-dependent effects represented in the volume
+trace. All information required to perform this simulation is present in the \ck{} file, with the
+exception of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
 %

### 2017-10-02 pyked.tex
WIP respond to #42

@@
 %
-To apply the effect of the volume trace to the \pybox|IdealGasReactor|, a
-\pybox|Wall| is installed between the reactor and environment, and assigned
-a velocity. Although we do not show the details here, a reference implementation
-of a class that computes a wall velocity given the volume history of the reactor
-is available in CanSen~\autocite{cansen}, in the
-\pybox|cansen.profile.VolumeProfile| class. Then, a Cantera \pybox|ReactorNet|
-can be used to advance the state through autoignition, in this case to an end
+In this example, \pybox|reac| represents the contents of the RCM reaction chamber and \pybox|env|
+represents the contents of the environment. \verb|air.xml| is an input file included with Cantera
+that represents an average composition of air. To apply the effect of the volume trace to the
+\pybox|IdealGasReactor|, a \pybox|Wall| is installed between the reactor and environment and
+assigned a velocity. The \pybox|Wall| allows the environment to do work on the reactor (or vice
+versa) and change the reactor's thermodynamic state; we use a \pybox|Reservoir| for the environment
+because in Cantera, \pybox|Reservoir|s always have a constant thermodynamic state and composition.
+Using a \pybox|Reservoir| accelerates the solution compared to using two \pybox|IdealGasReactor|s,
+since the composition and state of the environment are typically not necessary for the solution of
+auotignition problems. Although we do not show the details here, a reference implementation of a
+class that computes a wall velocity given the volume history of the reactor is available in
+CanSen~\autocite{cansen}, in the \pybox|cansen.profile.VolumeProfile| class. Then, a Cantera
+\pybox|ReactorNet| can be used to advance the state through autoignition, in this case to an end
 time of \SI{50}{\milli\second}:

### 2017-10-02 rebuttal.tex
WIP respond to #39
Modify the text to remove "might"
WIP respond to #42
Update response to #37
Respond to #27

@@
 \begin{response}
-
+    %
+    We have updated the text to incorporate the suggested change.
+    %
 \end{response}
@@
     internet-based options, we are also investigating the use of standalone packages, such as RDKit,
-    that would enable local retrieval of such information. We have not modified the text in response
-    to the comment.
+    that would enable local retrieval of such information.
     %
@@
 \begin{response}
-
+    %
+    We have modified the text to describe this example as an application.
+    %
 \end{response}
@@
 \begin{response}
-
+    %
+    \verb|air.xml| is an input file included with Cantera, representing an average composition of
+    air. We have modified the text to include this information.
+    %
 \end{response}

### 2017-10-02 pyked.tex
Respond to #27

@@
 the \ck{} format to include all information necessary to simulate a
-given experiment. A \ck{} file is generally broken into two main sections.
-The first section details all the ``meta'' information about the experiment
-and the \ck{} file itself. This section uses the common \yabox|author| mapping
-for any fields that specify an author. The \yabox|author| mapping
-contains the following fields:
+given experiment. A \ck{} file is generally broken into two main sections: a section containing the
+``meta'' information about the experiment and the \ck{} file itself, and a section containing the
+actual experimental data to be encoded.
+
+\subsection{Apparatus Information and Metadata}\label{sec:apparatus-information-and-metadata}
+
+We will first describe the fields in the ``meta'' section.
+This section uses a mapping that is called the \yabox|author| mapping in the schema. As this mapping
+is used for any fields that specify an author throughout the schema, we describe it first.
+The \yabox|author| mapping contains the following fields:
 %
@@
 %
-More generally for the meta information, the required keys in this section
-are:
+In general, the meta information describes the experimental facility and type of experiment, the
+reference from which the data was taken, and the author of the \ck{} file itself. For each of the
+keys below, we have indicated whether the key is required or optional, and the type of value
+associated with the key. The keys in the meta section include:
 %
 \begin{itemize}
+    \item \yabox|apparatus| (required, mapping): Information about the specific experimental
+    apparatus used to conduct the experiment. The fields in this mapping are:
+    \begin{itemize}
+        \item \yabox|facility| (optional, string): A unique name or identifier for the apparatus,
+        if the institution has several that are similar
+        \item \yabox|institution| (optional, string): The institution where the experimental
+        apparatus was located when the experiments were performed
+        \item \yabox|kind| (required, string): The type of apparatus used to
+        perform the experiment. Currently, only \yabox|shock tube| or
+        \yabox|rapid compression machine| are supported.
+    \end{itemize}
+    \item \yabox|chemked-version| (required, string): The version of the \ck{}
+    schema to which this file conforms
+    \item \yabox|experiment-type| (required, string): Currently, only
+    \yabox|ignition delay| is supported
     \item \yabox|file-author| (required, \yabox|author|-type mapping): The
@@
     file
-    \item \yabox|chemked-version| (required, string): The version of the \ck{}
-    schema to which this file conforms
     \item \yabox|reference| (required, mapping): The reference information for
@@
         mappings
+        \item \yabox|detail| (optional, string): A description of from where the
+        data originated (e.g., figure or table number)
+        \item \yabox|doi| (optional, string): The article DOI
         \item \yabox|journal| (optional, string): The name of the publishing
         journal
+        \item \yabox|pages| (optional, string): The article pages
+        \item \yabox|volume| (optional, integer): The journal volume number
         \item \yabox|year| (required, integer): The year of publication
-        \item \yabox|doi| (optional, string): The article DOI
-        \item \yabox|volume| (optional, integer): The journal volume number
-        \item \yabox|pages| (optional, string): The article pages
-        \item \yabox|detail| (optional, string): A description of from where the
-        data originated (e.g., figure or table number)
     \end{itemize}
-    \item \yabox|experiment-type| (required, string): Currently, only
-    \yabox|ignition delay| is supported
-    \item \yabox|apparatus| (required, string): The type of apparatus used to
-    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
 \end{itemize}
 
-The second section of the file encodes the experimental data. The \ck{} file
-describes these using a sequence of mappings, with the following information required
-in each element of the sequence:
+Several of these keys bear further discussion. The \yabox|chemked-version| is present in the file so
+that PyKED can determine whether the file conforms to a supported version of the schema. Although
+quite similar in name, the \yabox|file-version| represents the number of versions that have been
+created of the current file; it is an integer that should be incremented whenever changes are made
+to the file. Finally, the \yabox|reference| mapping is required in every file, even those that
+represent data that has not been published. However, the \yabox|reference| mapping only requires the
+\yabox|authors| sequence and the \yabox|year|, so that \ck{} files can be used internally within
+research groups to represent unpublished data prior to publication.
+
+\subsection{Ignition Delay Experimental Data}\label{sec:ignition-delay-experimental-data}
+
+The second section of the file encodes the experimental data as a sequence of mappings. The
+top-level key for the sequence is called \yabox|datapoints|. For the current version of the \ck{}
+schema, only autoignition delay experiments are supported. The following information os required in
+each element of the sequence:
 %

### 2017-10-02 pyked.tex
Clarify human readability as ease of use
In the YAML description section

@@
 \ck{} files use the YAML data serialization format~\autocite{yaml:1.2}. This
-format offers the advantages of being human readable, written in plain text, and
+format offers the advantages of being relatively easy for humans to read and write,
+encoded in plain text (ASCII or UTF-8 format), and
 having parsers in most common programming languages, including Python, C++, Java,

### 2017-10-02 pyked.tex
Clarify the purpose of the YAML section
Respond to #5

@@
 
-This section provides a brief introduction to the current version of the
-\ck{} database format, v0.1.6, which represents a functional standard
-for documenting autoignition measurements in shock tubes and rapid compression machines.
-Here we do not include a complete listing of available fields; a complete reference for the
-schema can be found in the online
-documentation.\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
-In Section~\ref{sec:usage-example}, we provide examples of \ck{} files and
-representative use cases. We indicate \yabox|yaml| keywords or values using text with
-gray background, and \pybox|Python| code using text with a yellow background.
+This section provides an introduction to the current version of the \ck{} database format, v0.1.6,
+which represents a functional standard for documenting autoignition measurements in shock tubes and
+rapid compression machines. Here we describe the fields that a user creating a \ck{} file is likely
+to use. There are several fields that are used internally in the schema that are not described here,
+but can be found in the online documentation for the
+schema.\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}} In
+Section~\ref{sec:usage-example}, we provide examples of \ck{} files and representative use cases. We
+indicate \yabox|yaml| keywords or values using text with gray background, and \pybox|Python| code
+using text with a yellow background.
 

### 2017-10-02 pyked.tex
Clarify where to specify common-properties

@@
 
-Frequently, experimental series hold certain properties constant, such as initial
-mixture composition or pressure, and use the same method to detect ignition delay.
-\ck{} files support a \yabox|common-properties| convenience section where these
-common details may be defined once, and referenced in each datapoint.
-This block uses the ability of YAML files to define an anchor with the
-\yabox|&| symbol and refer to that section later with the \yabox|*| symbol:
+Frequently, experimental series hold certain properties constant, such as initial mixture
+composition or pressure, and use the same method to detect ignition delay. \ck{} files support a
+\yabox|common-properties| convenience section where these common details may be defined once, and
+referenced in each element of the \yabox|datapoints| sequence. This block uses the ability of YAML
+files to define an anchor with the \yabox|&| symbol and refer to that section later with the
+\yabox|*| symbol:
 %

### 2017-10-02 pyked.tex
Add all possible keys to YAML section
We were only missing a few anyways...

@@
     with units and optionally uncertainty
+    \item \yabox|first-stage-ignition-delay| (optional, sequence): If two stages of ignition are
+    present in the experiment, this is the length of the first stage of ignition, and the
+    \yabox|ignition-delay| is then the overall ignition delay
+    \item \yabox|equivalence-ratio| (optional, float): The value of the equivalence ratio
     \item \yabox|composition| (required, mapping): The composition of the
-    mixture in the experiment, described via a nested mapping with the fields
+    mixture in the experiment, described via a nested mapping with the fields:
     \begin{itemize}
         \item \yabox|kind| (required, string): one of \yabox|mole fraction|, \yabox|mole percent|, or \yabox|mass fraction|
-        \item \yabox|species|: sequence of mappings, with the fields
+        \item \yabox|species|: sequence of mappings, with the fields:
         \begin{itemize}
@@
 
-The top-level key for the sequence is called \yabox|datapoints|. Each of the
+Each of the
 quantities in an element must be specified with units. This should be done as a
@@
 as seen in the previous code.
+
+\subsubsection{Shock Tube Experiments}\label{sec:shock-tube-experiments}
+
+There is one additional key that can be used to describe the facility-specific effects during a
+shock tube experiment:
+%
+\begin{itemize}
+    \item \yabox|pressure-rise| (optional, sequence): The pressure rise in the driven section after
+    the passage of the reflected shock, in dimensions of inverse time. Must include units and,
+    optionally, uncertainty
+\end{itemize}
+
+\subsubsection{Rapid Compression Machine Experiments}\label{sec:rcm-experiments}
+
+Rapid compression machine experiments typically report several additional pieces of information in
+addition to the ignition delay. As for shock tube experiments, these describe some of the
+facility-specific effects during an experiment. The keys to describe this information are:
+%
+\begin{itemize}
+    \item \yabox|volume-history| (optional, mapping): Specify the volume history of the reaction
+    chamber as a function of time during a rapid compression machine experiment. The fields in this
+    mapping are:
+    \begin{itemize}
+        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history.
+        The fields in this mapping are:
+        \begin{itemize}
+            \item \yabox|units| (required, string): the units of the volume, with dimensions length
+            cubed
+            \item \yabox|column| (required, integer): the 0-based index of the column storing the
+            volume information in the \yabox|values| array
+        \end{itemize}
+        \item \yabox|time| (required, mapping): A mapping describing the time in the history.
+        The fields in this mapping are:
+        \begin{itemize}
+            \item \yabox|units| (required, string): the units of the time
+            \item \yabox|column| (required, integer): the 0-based index of the column storing the
+            time information in the \yabox|values| array
+        \end{itemize}
+        \item \yabox|values| (required, sequence): A sequence of sequences describing the values of
+        the volume at the time points.
+    \end{itemize}
+    \item \yabox|compressed-temperature| (optional, sequence): The estimated temperature at the end
+    of the compression stroke
+    \item \yabox|compression-time| (optional, sequence): The time taken during the compression
+    stroke
+    \item \yabox|compressed-pressure| (optional, sequence): The measured pressure at the end of the
+    compression stroke
+\end{itemize}
+
+\subsection{Public Database of \ck{} Files}\label{sec:public-database}
 

### 2017-10-02 pyked.tex
Clarify RCM properties and other properties in PyKED
Work on #36

@@
 %====================================================================
-\section{PyKED architecture}\label{sec:pyked-architecture}
+\section{PyKED Architecture}\label{sec:pyked-architecture}
 %
@@
     value.
-    \item \pybox|ignition_delay|, \pybox|temperature|, \pybox|pressure|,
-    \pybox|pressure_rise|: The value of these data are stored as instances of
-    the \pybox|Quantity| class from the Pint \autocite{Grecco2016} package,
-    possibly with associated uncertainty
     \item \pybox|composition|: A list of dictionaries of the species and their
@@
     measurement of ignition delay
-\end{itemize}
-
-If the \ck{} file encodes an RCM experiment, there are several specialized fields that can be
-included.
-
-\begin{itemize}
     \item \pybox|volume_history|: If the \yabox|volume-history| of an RCM
@@
     \autocite{vanderWalt:2011np}
-    \item \pybox|compression_time|: The duration of the compression stroke
-    \item \pybox|compressed_temperature|: The estimated reactor temperature at the end of the
-    compression stroke
-    \item \pybox|compressed_pressure|: The measured reactor pressure at the end of the compression
-    stroke
 \end{itemize}
+%
+Other instance attributes are stored as instances of the \pybox|Quantity| class from the
+Pint \autocite{Grecco2016} package, possibly with an associated uncertainty. These include the
+\pybox|ignition_delay|, \pybox|temperature|, \pybox|pressure|, \pybox|pressure_rise|,
+\pybox|compression_time|, \pybox|compressed_temperature|, and \pybox|compressed_pressure|
+attributes, each of which represents the similarly-named field in the \ck{} schema.
 

### 2017-10-02 pyked.tex
Fix weirdness of lists of attributes
Add Appveyor to CI section

@@
 
-In addition, specific fields from the \pybox|Reference| and \pybox|Apparatus|
-attributes can be included by specifying the name after a colon. These options
+In addition, specific fields from the \pybox|`Reference'| and \pybox|`Apparatus'|
+columns can be included by specifying the name after a colon. These options
 are:
@@
     \item \pybox|'Apparatus:Facility'|
-\end{itemize*}
-
-% This weirdness is necessary to prevent the list from running over the margin
-\noindent\begin{itemize*}
     \item \pybox|'Apparatus:Institution'|
@@
 at least once.
-Travis-CI\footnote{\url{https://travis-ci.org/pr-omethe-us/PyKED}} provides continuous
-integration service, including running the test suite, building and distributing binary
-installer packages, and building and publishing the online documentation. This automation
-runs the test suite after every change to the source code, keeps the documentation
-up-to-date with the latest changes, and streamlines the release of new versions.
+
+Travis-CI\footnote{\url{https://travis-ci.org/pr-omethe-us/PyKED}} and
+Appveyor\footnote{\url{https://ci.appveyor.com/project/Prometheus/pyked}} provide continuous
+integration (CI) service for Linux, macOS, and Windows platforms. The CI services run the test suite
+on every change to the source code, build and distribute binary installer packages for every
+release, and build and publish the online documentation automatically. This automation ensures that
+test failures are caught quickly, keeps the documentation up-to-date with the latest changes, and
+streamlines the release of new versions.
 

### 2017-10-02 pyked.tex
Start more thorough description of Python examples
Work on #40

@@
 how the volume of the reactor varies with time, for simulating the compression
-stroke and post-compression processes.
+stroke and post-compression processes:
 %
@@
 trace. All information required to perform this simulation is present in the \ck{} file, with the
-exception of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
+exception of a chemical kinetic model for \ce{H2}\slash \ce{CO} combustion.
+
+In Python, additional functionality can be imported into a script or session by the \pybox|import|
+keyword. Cantera and PyKED both must be imported into the session so that we can work with the code.
+In the case of Cantera, we will use many functions from the library, so we assign it an abbreviation
+(\pybox|ct|) for convenience. From PyKED, we will only be using the \pybox|ChemKED| class, so this
+is all that is imported:
 %
@@
 from pyked import ChemKED
-
-# Load the ChemKED file and retrieve the first element of the
-# datapoints list, which is an instance of the DataPoint class
+\end{pythonbox}
+%
+Next, we have to load the ChemKED file and retrieve the first element of the \pybox|datapoints|
+list. Although this file only encodes a single experiment, the \pybox|datapoints| attribute will
+always be a list (in this case, of length 1). As mentioned previously, the elements of the
+\pybox|datapoints| list are instances of the \pybox|DataPoint| class, which we store in the variable
+\pybox|dp|.
+%
+\begin{pythonbox}
 ck = ChemKED('testfile_rcm.yaml')
@@
 units as input. Therefore, we use the built-in capabilities of Pint to convert
-the units from those specified in the \ck{} file to SI units:
+the units from those specified in the \ck{} file to SI units, and we use the \pybox|magnitude|
+attribute of the \pybox|Quantity| class to take only the numerical part. We also retrieve the
+initial mixture mole fractions in a format Cantera will understand:
 %
@@
 X_initial = dp.get_cantera_mole_fraction()
+\end{pythonbox}
+%
+With these properties defined, we have to create the objects in Cantera that represent the physical
+state of the system to be studied.
 

### 2017-10-02 pyked.tex
More work on Python description
Work on #40

@@
 With these properties defined, we have to create the objects in Cantera that represent the physical
-state of the system to be studied.
+state of the system to be studied. In Cantera, the \pybox|Solution| class stores the thermodynamic,
+kinetic, and transport data from an input file in the CTI format. After the \pybox|Solution| object
+is created, we can set the initial temperature, pressure, and mole fractions using the \pybox|TPX|
+attribute of the \pybox|Solution| class:
 
+\begin{pythonbox}
 # Load the mechanism and set the initial state of the mixture
@@
 gas.TPX = T_initial, P_initial, X_initial
+\end{pythonbox}
+
+With the thermodynamic and kinetic data loaded and the initial conditions defined, we  need to
+install the \pybox|Solution| instance into an \pybox|IdealGasReactor| which implements the equations
+for mass, energy, and species conservation.
 
@@
 since the composition and state of the environment are typically not necessary for the solution of
-auotignition problems. Although we do not show the details here, a reference implementation of a
+autoignition problems. Although we do not show the details here, a reference implementation of a
 class that computes a wall velocity given the volume history of the reactor is available in

### 2017-10-02 pyked.tex
added missing \begin{pythonbox} for example

@@
 state of the system to be studied.
-
+%
+\begin{pythonbox}
 # Load the mechanism and set the initial state of the mixture

### 2017-10-04 pyked.tex
Finish really explaining the first Python example
Work on #40

--- 2017-10-02 pyked.tex
@@
 In Python, additional functionality can be imported into a script or session by the \pybox|import|
-keyword. Cantera and PyKED both must be imported into the session so that we can work with the code.
-In the case of Cantera, we will use many functions from the library, so we assign it an abbreviation
-(\pybox|ct|) for convenience. From PyKED, we will only be using the \pybox|ChemKED| class, so this
-is all that is imported:
+keyword. Cantera, NumPy, and PyKED must be imported into the session so that we can work with the
+code. In the case of Cantera and NumPy, we will use many functions from these libraries, so we
+assign them abbreviations (\pybox|ct| and \pybox|np|, respectively) for convenience. From PyKED, we
+will only be using the \pybox|ChemKED| class, so this is all that is imported:
 %
@@
 import cantera as ct
+import numpy as np
 from pyked import ChemKED
@@
 install the \pybox|Solution| instance into an \pybox|IdealGasReactor| which implements the equations
-for mass, energy, and species conservation.
+for mass, energy, and species conservation. In addition, we create a \pybox|Reservoir| to represent
+the environment external to the reaction chamber. The input file used for the environment,
+\verb|air.xml|, is included with Cantera and represents an average composition of air.
 
+\begin{pythonbox}
 # Create the reactor and the outside environment
@@
 %
-In this example, \pybox|reac| represents the contents of the RCM reaction chamber and \pybox|env|
-represents the contents of the environment. \verb|air.xml| is an input file included with Cantera
-that represents an average composition of air. To apply the effect of the volume trace to the
-\pybox|IdealGasReactor|, a \pybox|Wall| is installed between the reactor and environment and
-assigned a velocity. The \pybox|Wall| allows the environment to do work on the reactor (or vice
-versa) and change the reactor's thermodynamic state; we use a \pybox|Reservoir| for the environment
-because in Cantera, \pybox|Reservoir|s always have a constant thermodynamic state and composition.
-Using a \pybox|Reservoir| accelerates the solution compared to using two \pybox|IdealGasReactor|s,
-since the composition and state of the environment are typically not necessary for the solution of
-autoignition problems. Although we do not show the details here, a reference implementation of a
-class that computes a wall velocity given the volume history of the reactor is available in
-CanSen~\autocite{cansen}, in the \pybox|cansen.profile.VolumeProfile| class. Then, a Cantera
-\pybox|ReactorNet| can be used to advance the state through autoignition, in this case to an end
-time of \SI{50}{\milli\second}:
+To apply the effect of the volume trace to the \pybox|IdealGasReactor|, a \pybox|Wall| must be
+installed between the reactor and environment and assigned a velocity. The \pybox|Wall| allows the
+environment to do work on the reactor (or vice versa) and change the reactor's thermodynamic state;
+we use a \pybox|Reservoir| for the environment because in Cantera, \pybox|Reservoir|s always have a
+constant thermodynamic state and composition. Using a \pybox|Reservoir| accelerates the solution
+compared to using two \pybox|IdealGasReactor|s, since the composition and state of the environment
+are typically not necessary for the solution of autoignition problems. Although we do not show the
+details here, a reference implementation of a class that computes the wall velocity given the volume
+history of the reactor is available in CanSen~\autocite{cansen}, in the
+\pybox|cansen.profile.VolumeProfile| class.
 %
 \begin{pythonbox}
+# Retrieve the time and volume from the
+# history in the datapoint
 time = dp.volume_history.time
 volume = dp.volume_history.volume
+# Install the Wall between the
+# reactor and environment
 ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))
+\end{pythonbox}
 
+Then, the \pybox|IdealGasReactor| is installed in a \pybox|ReactorNet|. The \pybox|ReactorNet|
+implements the connection to the numerical solver (CVODES \autocite{hindmarsh_sundials:_2005} is
+used in Cantera) to solve the energy and species equations. For this example, it is best practice
+to set the maximum time step allowed in the solution to be the minimum time difference in the time array from the volume trace:
+
+\begin{pythonbox}
 netw = ct.ReactorNet([reac])
+netw.set_max_time_step(np.min(np.diff(time)))
+\end{pythonbox}
 
+To calculate the ignition delay, we will follow the definition specified in the ChemKED file for
+this experiment, where the experimentalists used the maximum of the time derivative of the pressure
+to define the ignition delay. To calculate this derivative, we need to store the state variables and the composition on each time step, so we initialize several Python lists to act as storage:
+
+\begin{pythonbox}
+# Initialize lists to store solution information
+time = []
+temperature = []
+pressure = []
+volume = []
+mass_fractions = []
+\end{pythonbox}
+
+Finally, the problem is integrated using the \pybox|step| method of the \pybox|ReactorNet|. The
+\pybox|step| method takes one timestep forward on each call, with step size determined by the CVODES
+solver (CVODES is an adaptive time-stepping algorithm). On each step, we add the relevant variables
+to their respective lists. The problem is integrated until a user-specified end time, in this case
+\SI{50}{\milli\second}, although in principle, the user could end the simulation on any condition
+they choose:
+
+\begin{pythonbox}
 # Integrate for 50 ms
 while netw.time < 0.05:
+    time.append(netw.time)
+    temperature.append(reac.T)
+    pressure.append(reac.thermo.P)
+    volume.append(reac.volume)
+    mass_fractions.append(reac.Y)
     netw.step()
 \end{pythonbox}
-%
-Although not shown in this example, the user would typically store information
-about the \pybox|IdealGasReactor| (e.g., temperature, pressure, mass fractions)
-at each time step of the integration for post-processing.
+
+At this point, the user could post-process the information in the \pybox|pressure| list to calculate
+the derivative by whatever algorithm they choose.
 

### 2017-10-04 rebuttal.tex
More work on example explanation
Also add multiprocessing. I haven't checked the code yet though! Work on
#44, #43, #42, #41, #40, #39

@@
 \begin{response}
-
+    %
+    We have corrected this oversight.
+    %
 \end{response}
@@
 \begin{response}
-
+    %
+    We have expanded the explanation for all of the Python examples to give more detail of the
+    process of creating the simulation.
+    %
 \end{response}
@@
 \begin{response}
-
+    %
+    We have truncated the datapoints list as the reviewer suggests.
+    %
 \end{response}
@@
 \begin{response}
-
+    %
+    We have modified the example to include more explanatory text and to use Python's built-in
+    multiprocessing library to run the simulations on many processors. We note that if Cantera has
+    been built with multi-threaded BLAS and LAPACK libraries (such as the Intel MKL), then the 0-D
+    solver will automatically utilize the multithreading capabilities of those libraries. Therefore,
+    using the multiprocessing library may not give much benefit in terms of overall speedup,
+    depending on the complexity of the chemical kinetic mechanism (i.e., how long the simulation
+    takes to run), how many processors are available, and the availability of the multi-threaded
+    BLAS and LAPACk libraries. However, we have not explained this detail of Cantera in the text to
+    avoid confusing the reader.
+    %
 \end{response}

### 2017-10-04 pyked.tex
More work on example explanation
Also add multiprocessing. I haven't checked the code yet though! Work on
#44, #43, #42, #41, #40, #39

@@
 
-At this point, the user could post-process the information in the \pybox|pressure| list to calculate
+At this point, the user would post-process the information in the \pybox|pressure| list to calculate
 the derivative by whatever algorithm they choose.
@@
 shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
-meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
-%
+meta information below for space; the format is largely similar to the example in
+Section~\ref{sec:rcm-modeling}.
+
 This \ck{} file specifies multiple data points with some common
 conditions, including a common mixture composition and common definition of
-ignition delay. Therefore, a \yabox|common-properties| section is specified.
+ignition delay. Therefore, a \yabox|common-properties| section is specified, followed by the
+datapoints list:
 %
@@
     equivalence-ratio: 0.5
-  - temperature:
-      - 1497 kelvin
-    ignition-delay:
-      - 212 us
-    pressure:
-      - 1.55 atm
-    composition: *comp
-    ignition-type: *ign
-    equivalence-ratio: 0.5
-  - temperature:
-      - 1562 kelvin
-    ignition-delay:
-      - 105 us
-    pressure:
-      - 1.50 atm
-    composition: *comp
-    ignition-type: *ign
-    equivalence-ratio: 0.5
+    ...
 \end{yamlbox}
-
-In this example, we would like to run constant-volume simulations at each
+%
+As before, we have truncated the datapoints list for space.
+
+In this example, we will run constant-volume simulations at each
 pressure and temperature condition in the \yabox|datapoints| list. Once again,
 the \ck{} file specifies all the information required for the simulations except
-for the chemical kinetic model, and Cantera can be used to simulate autoignition:
+for the chemical kinetic model, and Cantera can be used to simulate autoignition. The setup steps
+are the same as in the previous example, with one exception. In this example, we also import
+Python's built-in multiprocessing library so that we can run the simulations on multiple cores. We
+import the \pybox|Pool| class, which offers facilities to manage a jobs on a pool of processors:
 %
@@
 import cantera as ct
+from multiprocessing import Pool
 from pyked import ChemKED
-
+\end{pythonbox}
+
+Then, we define a function that will be mapped onto each job. This function takes the initial
+temperature, pressure, and mole fractions as input and returns the simulated ignition delay time,
+defined as the time the temperature increases by \SI{400}{\K} over the initial temperature, a
+simplified definition for this example. In general, the user could process the mole fraction of
+\ce{OH^*} (provided that the mechanism included accurate chemistry for \ce{OH^*}) to match the
+definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
+be supplied by the user:
+%
+\begin{pythonbox}
+def run_simulation(T, P, X):
+    gas = ct.Solution('tbuoh-mechanism.cti')
+    gas.TPX = T, P, X
+    reac = ct.IdealGasReactor(gas)
+    netw = ct.ReactorNet([reac])
+    while reac.T < T + 400:
+        netw.step()
+
+    return netw.time
+\end{pythonbox}
+
+Then, we load the ChemKED file and generate a list of initial conditions that will be mapped onto
+the \pybox|run_simulation()| function. We first define a convenience function to collect the input
+from a single datapoint and return a Python tuple with the conditions (tuples are lightweight groups
+of data in Python). Then, we use the built-in \pybox|map| function to apply the
+\pybox|collect_input| function to each of the elements in the \pybox|ck.datapoints| list: convert a
+\pybox|Quantity| to some input units and return the magnitude, then use Python's list comprehension
+syntax to generate a list of tuples (tuples are lightweight groups of data in Python):
+%
+\begin{pythonbox}
 ck = ChemKED('Stranic2012-tbuoh.yaml')
 
-gas = ct.Solution('tbuoh-mechanism.cti')
-
-# Since the composition is specified by the common-properties,
-# just take the composition of the first DataPoint
-X_initial = ck.datapoints[0].get_cantera_mole_fraction()
-
-# Loop through each of the DataPoints
-for dp in ck.datapoints:
+def collect_input(dp):
     T_initial = dp.temperature.to('K').magnitude
     P_initial = dp.pressure.to('Pa').magnitude
-    gas.TPX = T_initial, P_initial, X_initial
-    reac = ct.IdealGasReactor(gas)
-    netw = ct.ReactorNet([reac])
-    # Define ignition delay as T_initial + 400 K
-    while reac.T < T_initial + 400:
-        netw.step()
-
-    print('The ignition delay for T_initial={}, P_initial={} is: '
-          '{} seconds'.format(T_initial, P_initial, netw.time)
-          )
-\end{pythonbox}
-%
-The ignition delay in this example is defined as equal to the \SI{400}{\kelvin}
-above the initial temperature, a simplified definition used for this example.
-In general, the user could post-process the concentration information from the
-simulation to determine one-half the maximum \ce{OH^*} concentration to match
-the experimental definition of ignition delay.
+    X_initial = dp.get_cantera_mole_fraction()
+    return (T_initial, P_initial, X_initial)
+
+initial_conditions = map(collect_input, ck.datapoints)
+initial_conditions = [(to_SI(dp.temperature, 'K'), to_SI(dp.pressure, 'Pa'), dp.get_cantera_mole_fraction()) for dp in ck.datapoints]
+\end{pythonbox}
+
+Finally, we create the processor \pybox|Pool| and send the jobs out to run:
+%
+\begin{pythonbox}
+with Pool(12) as pool:
+    ignition_delays = pool.starmap(run_simulation, initial_conditions)
+
+for (T, P, X), tau in zip(initial_conditions, ignition_delays):
+    print(f'The ignition delay for T_initial={T}, P_initial={P} is: {tau} seconds')
+\end{pythonbox}
+%
+The simulated ignition delay results are returned in the \pybox|ignition_delays| list. The results
+are printed to the screen using a Python formatted string (f-string).
 %====================================================================

### 2017-10-04 rebuttal.tex
Fix missing close bracket in response

@@
     version and report any feature deprecation or removal to the user (see
-    \url{https://github.com/pr-omethe-us/PyKED/issues/22), and we plan to provide an automated
+    \url{https://github.com/pr-omethe-us/PyKED/issues/22}), and we plan to provide an automated
     upgrade path for ChemKED files when backwards-incompatible changes are made. In addition, we use

### 2017-10-07 pyked.tex
Fix common-properties YAML demo
add explanation of "open" for #6

@@
 
+%discuss open source definition here
+(When we describe a software package or project as ``closed,'' we are generally
+contrasting it with the open source definition~\cite{opensource}. In brief, to be
+``open,'' a project must provide free and unrestricted access to source code and\slash
+or data, allow free redistribution, permit derived works, and not discriminate against
+any people or areas of work, along with a few other criteria.)
+
 More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
@@
 datapoints:
-  composition: *comp
+  - composition: *comp
     ignition-type: *ign
-  temperature: 1000 kelvin
+    temperature:
+      - 1000 kelvin
   ...

### 2017-10-07 rebuttal.tex
more rebuttal for #6
minor changes to close #23

--- 2017-10-04 rebuttal.tex
@@
     \emph{openly} available.
+
+    To clarify the meaning of ``open'' when used in our article, we have added a brief definition
+    and link to the definition of open source, which comes from the Open Source Initiative.
 \end{response}
@@
     for every field while YAML does not; in XML, information can be specified as properties of tags
-    or as elements (requiring the user to know where to look for information) while YAML has only 3 basic data structures (keys, values, and lists).
+    or as elements (requiring the user to know where to look for information) while YAML has only
+    three basic data structures (keys, values, and lists).
 
@@
 
-    We have modified the text to clarify this point, including describing the lack of human
+    We have modified the text on \textbf{page 2} to clarify this point, including describing the lack of human
     readability as a formatting issue (but one that is inherent to XML), rather than a technical

### 2017-10-07 pyked.tex
minor changes to close #23

@@
 format is intended to be a machine-readable markup language rather than a data format.
-As a general-purpose document markup language, XML is able to represent very detailed
+As a general-purpose document markup language, XML can represent detailed
 information about a dataset, but implementing this detail requires XML-formatted
-files to be significantly more verbose than files formatted with a language focused
+files to be much more verbose than files formatted with a language focused
 only on representing data structures. The verbosity inherent in the XML format

### 2017-10-07 rebuttal.tex
finalized response to #8

@@
 \setlength{\parindent}{0in}
-\setlength{\parskip}{0.4cm}
+\setlength{\parskip}{0.25cm}
 
-\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup\vspace{0.5\baselineskip}}
-\newenvironment{response}{\vspace{0.5\baselineskip}\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
+\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup}
+\newenvironment{response}{\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
 \newcommand{\review}[1]{{\itshape#1}}
@@
 
-    To clarify the meaning of ``open'' when used in our article, we have added a brief definition
-    and link to the definition of open source, which comes from the Open Source Initiative.
+    To clarify the meaning of ``open'' when used in our article, we have added to \textbf{page 2} a
+    brief definition and link to the definition of open source, which comes from the
+    Open Source Initiative.
 \end{response}
@@
     %
-    We do not think that the technical details of the implementation of the original PrIMe system
+    We have added the new PrIMe URL to the manuscript on \textbf{page 2}, and thank the reviewer for pointing us to that.
+
+    However, we do not think that the technical details of the implementation of the original PrIMe system
     are relevant to this work, so we have not included them. In addition, we cannot know
-    the plans of Prof. Frenklach, and we do not feel it would be appropriate to comment on them in
-    our manuscript. As such, we have not added these suggestions to the paper. We have, however,
-    added the new URL to the current manuscript, and thank the reviewer for pointing us to that.
+    the plans of Prof.~Frenklach, and we do not feel it would be appropriate to comment on them in
+    our manuscript. As such, we have not added these suggestions to the paper.
     %
@@
 
-    In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
+    In my opinioo n the YAML format resembles the XML format strongly. XML format is human-readable,
     however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to

### 2017-10-07 rebuttal.tex
finalized response to #9

@@
     %
-    We have incorporated these modifications into the suggested paragraph, although we have modified
-    the format slightly to better fit the flow of the text.
+    We have incorporated these modifications into the suggested paragraph on \textbf{page 2},
+    although we have modified the format slightly to better fit the flow of the text.
     %

### 2017-10-07 rebuttal.tex
add page reference in response to #10

@@
     %
-    We have incorporated most of the recommended text in our revision. However, we did not add
-    ``human-readable'', consistent with the definition used throughout (i.e., the XML file is more
-    verbose and less easily understood).
+    We have incorporated most of the recommended text in our revision on \textbf{page 2}.
+    However, we did not add ``human-readable'', consistent with the definition used throughout
+    and explained on \textbf{page 2} (i.e., the XML file is more verbose and less easily understood).
     %

### 2017-10-07 rebuttal.tex
adds page reference for #11
adds page references for #12

@@
     people, rather than the technical limitations of a format. We have changed the word
-    ``usability'' to ``readability'' to emphasize this.
+    ``usability'' to ``readability'' on \textbf{page 2} to emphasize this.
     %
@@
     %
-    We have updated the text to reflect this information; however, we argue that requiring
-    registration to view the specification and data limits the possibility of use of the standard
-    and differs from the openness that we are advocating, where the specification and data files are
-    open for anyone to view, with or without registration.
+    We have updated the text on \textbf{page 2} to reflect this information; however, we argue
+    that requiring registration to \emph{view} the specification and data limits use
+    of the standard and differs from the openness that we are advocating (now defined on \textbf{page 2}),
+    where the specification and data files are open for anyone to view, with or without
+    registration.
     %

### 2017-10-07 rebuttal.tex
adds page ref for #13
adds page ref for #14

@@
 \begin{response}
-    We added the suggested text and reference.
+    We added the suggested text and reference to \textbf{page 3}.
 \end{response}
@@
     %
-    We have implemented an option for baseline extrapolation of the slope. We have added the
-    description of this ignition type to the mentioned portion of the manuscript. To our knowledge,
+    We have implemented an option for baseline extrapolation of the slope, and added the
+    description of this ignition type to \textbf{page 6} of the manuscript. To our knowledge,
     the options now supported represent all of the common methods of specifying the ignition delay

### 2017-10-07 rebuttal.tex
minor edits for #15

@@
     specifying the initial composition as a mass fraction, which ReSpecTh does not support. We have
-    modified the sentence to indicate that there are some incompatibilities in both directions, not
-    all ChemKED files can be converted to ReSpecTh files, and not all ReSpecTh files can be
-    converted to ChemKED files due to missing data elements in both directions.
+    modified the sentence on \textbf{page 10} to indicate that there are some incompatibilities
+    in both directions, not all ChemKED files can be converted to ReSpecTh files, and
+    not all ReSpecTh files can be converted to ChemKED files due to missing data elements in both directions.
     %

### 2017-10-07 pyked.tex
minor edits for #15

@@
 
-PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/}
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/})
 has a number of features that make
@@
 may not be possible, as \ck{} files support some data that ReSpecTh files do not support. We note
-that the reverse is also true, not all ReSpecTh files may be converted to \ck{} files because not
-all of the data that can be stored in a ReSpecTh format is supported by the \ck{} format.
+that the reverse is also true: not all ReSpecTh files may be converted to \ck{} files because not
+all of the data that can be stored in a ReSpecTh format is (currently) supported by the \ck{} format.
 

### 2017-10-07 rebuttal.tex
final edits for #21
adds page ref for #22

@@
     ChemKED format. The work is available publicly at
-    \url{https://github.com/pr-omethe-us/ChemKED-gui}.
+    \url{https://github.com/pr-omethe-us/ChemKED-gui}. In addition, we have a built-in converter
+    for generating ChemKED files from existing ReSpecTh files.
     %
@@
     %
-    Thank you for the comment, we have corrected this oversight in the roadmap and the conclusions.
+    Thank you for the comment, we have corrected this oversight in the roadmap and the
+    conclusions (\textbf{page 17}).
     %

### 2017-10-07 rebuttal.tex
adds page ref for #23

@@
 
-    In my opinioo n the YAML format resembles the XML format strongly. XML format is human-readable,
+    In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
     however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to
@@
     say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really
-    mean human-\emph{understandable}. We have modified the text to clarify this point; please see
-    also the response to a similar comment from Reviewer \#1.
+    mean human-\emph{understandable}. We have modified the text on \textbf{page 2} to clarify this
+    point; please see also the response to a similar comment from Reviewer \#1.
     %

### 2017-10-07 rebuttal.tex
response to #24
adds page number for #25

@@
 \begin{response}
-
+    %
+    PyTeCK is a software package for automatically validating chemical kinetic models based
+    on experimental data encoded in the ChemKED format; thus, PyTeCK relies on the PyKED software
+    to interpret and validate ChemKED files.
+    %
 \end{response}
@@
     %
-    We have edited that sentence to correct the error.
+    We have edited that sentence on \textbf{page 3} to correct the error.
     %

### 2017-10-07 rebuttal.tex
adds page ref for #27

@@
     In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor.
-    This option turns off all of the validation, including the validation of the reference material
+    This option turns off all validation, including the validation of the reference material
     and enforcement of required fields. We recognize that this is a limitation at present, if users
@@
     %
-    We have updated the text to incorporate the suggested change.
+    We have updated the text on \textbf{page 4} to incorporate the suggested change.
     %

### 2017-10-07 pyked.tex
response to close #28

@@
             \item \yabox|amount| (required, sequence): The mole fraction or percent, or the mass fraction, optionally with uncertainty
-            \item \yabox|InChI|, \yabox|SMILES|, \yabox|atomic-composition|, or
-            \yabox|elemental-composition| (required, string or sequence): The InChI
-            or SMILES string representing the molecule, or its atomic or elemental
-            composition (\yabox|atomic-composition| and\\
-            \yabox|elemental-composition| are synonyms) as a sequence
+            \item \yabox|InChI|, \yabox|SMILES|, or \yabox|atomic-composition|
+            (required, string or sequence): The InChI
+            or SMILES string representing the molecule, or its atomic
+            composition as a sequence
         \end{itemize}

### 2017-10-07 rebuttal.tex
response to close #28
added a bit more to #30

@@
 \begin{response}
-
+    We agree with the reviewer on this point, and have removed \texttt{elemental-composition}
+    as an option for clarity. The text on \textbf{page 6} has been updated to reflect this.
 \end{response}
@@
 \begin{response}
+    %
 
+    %
 \end{response}
@@
     consistently, etc. Furthermore, such human review can help to catch some errors that may be
-    introduced by manual creation of the files, as pointed out by the reviewer previously. Finally,
-    we would note that all of the existing database have a similar requirement, where users submit
+    introduced by manual creation of the files, as pointed out by the reviewer previously.
+
+    All the existing database have a similar requirement, where users submit
     data files for inclusion that are approved by the database owners after some review. In this
     sense, we actually see the Pull Request mechanism of GitHub as an advantage, since all reviews
-    can be done publicly and some of the review procedures (style checks, reference validity, etc)
+    can be done publicly and some of the review procedures (style checks, reference validity, etc.)
     may be done automatically by a script associated with the GitHub repository. In this way, the
     entire community can contribute to improving the quality of the database.
+
+    Lastly, while we hope other researchers will contribute to this database (such that it becomes
+    more of a community effort than our effort), there is no requirement for others to use our
+    database to use ChemKED files or PyKED. Others are free to use the tools internally, or
+    download\slash fork our database and build their own.
     %

### 2017-10-07 rebuttal.tex
adds some page refs

@@
     ChemKED database have been coupled up to now, and we have updated the version mentioned in the
-    text to vX.X.X, which we released in conjunction with submitting this response after the helpful
-    comments of the reviewers.
+    text on \textbf{page 3} to vX.X.X, which we released in conjunction with submitting
+    this response after the helpful comments of the reviewers.
     %
@@
     %
-    We removed the extraneous comma.
+    We removed the extraneous comma on \textbf{page 8}.
     %

### 2017-10-07 pyked.tex
made all section names sentence case

@@
 %====================================================================
-\section{Overview of ChemKED format}\label{sec:overview-of-format}
+\section{Overview of ChemKED format}
+\label{sec:overview-of-format}
 
@@
 
-\subsection{Apparatus Information and Metadata}\label{sec:apparatus-information-and-metadata}
+\subsection{Apparatus information and metadata}
+\label{sec:apparatus-information-and-metadata}
 
@@
 
-\subsection{Ignition Delay Experimental Data}\label{sec:ignition-delay-experimental-data}
+\subsection{Ignition delay experimental data}
+\label{sec:ignition-delay-experimental-data}
 
@@
 
-\subsubsection{Shock Tube Experiments}\label{sec:shock-tube-experiments}
+\subsubsection{Shock tube experiments}
+\label{sec:shock-tube-experiments}
 
@@
 
-\subsubsection{Rapid Compression Machine Experiments}\label{sec:rcm-experiments}
+\subsubsection{Rapid compression machine experiments}
+\label{sec:rcm-experiments}
 
@@
 
-\subsection{Public Database of \ck{} Files}\label{sec:public-database}
+\subsection{Public database of \ck{} files}
+\label{sec:public-database}
 
@@
 %====================================================================
-\section{PyKED Architecture}\label{sec:pyked-architecture}
+\section{PyKED architecture}
+\label{sec:pyked-architecture}
 %
@@
 %====================================================================
-\section{Usage examples}\label{sec:usage-example}
+\section{Usage examples}
+\label{sec:usage-example}
 
@@
 
-\subsection{RCM modeling with varying reactor volume}\label{sec:rcm-modeling}
+\subsection{RCM modeling with varying reactor volume}
+\label{sec:rcm-modeling}
 
@@
 
-\subsection{Shock tube modeling with constant volume}\label{sec:shock-tube}
+\subsection{Shock tube modeling with constant volume}
+\label{sec:shock-tube}
 

### 2017-10-08 rebuttal.tex
adds page ref to #35

@@
     this in the Cerberus package has very recently been resolved (see
-    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error reporting
+    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error-reporting
     mechanism in PyKED to report well-formatted errors that have a detailed description of the
@@
     This was a typo in the manuscript. All of the keywords here include spaces in their name, both
-    in the YAML schema and in the PyKED interface. This has been corrected in the manuscript and we
-    have added a test to the PyKED package to ensure the values are correct.
+    in the YAML schema and in the PyKED interface. This has been corrected in the manuscript on
+    \textbf{page 9} and we have added a test to the PyKED package to ensure the values are correct.
     %

### 2017-10-08 rebuttal.tex
adds page ref to #36
minor edit to close #37

@@
     %
-    We have consolidated the RCM-specific information into a separate list, and eliminated the
+    We have consolidated the RCM-specific information into separate list for ChemKED on
+    \textbf{page 7--8} and PyKED on \textbf{page 9}, which eliminated the
     duplicated phrase.
@@
     the code at this time, and aside from specialized features, we do not anticipate that an
-    internet connection will ever be required to run the basic software. Second, for the specific
-    feature mentioned here, this will be something that the user is explicitly requesting (e.g., by
+    internet connection will ever be required to run the basic software.
+
+    Second, the specific feature mentioned here will be something that the user explicitly requests (e.g., by
     attempting to convert mass to mole fractions). Thus, we expect that they are willing to wait for

### 2017-10-08 rebuttal.tex
adds page ref for #39
adds page refs for #40
adds page ref for #42

@@
     %
-    We have modified the text to describe this example as an application.
+    We have modified the text on \textbf{page 12} to describe this example as an application.
     %
@@
     %
-    We have corrected this oversight.
+    We have corrected this oversight, with more explanation and introduction of the Python
+    examples on \textbf{pages 12--17}.
     %
@@
     \verb|air.xml| is an input file included with Cantera, representing an average composition of
-    air. We have modified the text to include this information.
+    air. We have modified the text on \textbf{page 13} to include this information.
     %

### 2017-10-08 rebuttal.tex
adds page ref and number of processes for #43

@@
     %
-    We have truncated the datapoints list as the reviewer suggests.
+    We have truncated the datapoints list on \textbf{pages 12 and 16} as the reviewer suggests.
     %
@@
     %
-    We have modified the example to include more explanatory text and to use Python's built-in
+    We have modified the example on \textbf{pages 16--17} to include more
+    explanatory text and to use Python's built-in
     multiprocessing library to run the simulations on many processors. We note that if Cantera has
@@
     using the multiprocessing library may not give much benefit in terms of overall speedup,
-    depending on the complexity of the chemical kinetic mechanism (i.e., how long the simulation
+    depending on the complexity of the chemical kinetic model (i.e., how long the simulation
     takes to run), how many processors are available, and the availability of the multi-threaded

### 2017-10-08 pyked.tex
mechanism -> kinetic model
adds page ref and number of processes for #43

@@
 simplified definition for this example. In general, the user could process the mole fraction of
-\ce{OH^*} (provided that the mechanism included accurate chemistry for \ce{OH^*}) to match the
+\ce{OH^*} (provided that the kinetic model includes accurate chemistry for \ce{OH^*}) to match the
 definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
@@
 
-Finally, we create the processor \pybox|Pool| and send the jobs out to run:
+Finally, we create the processor \pybox|Pool| (with 12 processes) and send the jobs out to run:
 %

### 2017-10-08 rebuttal.tex
Soften "trivial" in the response

@@
     that must also be processed with an appropriate software package (PyKED, in the case of ChemKED
-    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED,
-    it is trivial to produce an Excel or CSV file so that one may view such a summary if desired.
-    See also the response to a similar comment below.
+    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED, it
+    is already possible to produce an Excel or CSV file (using the Pandas DataFrame integration) so
+    that one may view such a summary if desired. See also the response to a similar comment below.
     %

### 2017-10-08 rebuttal.tex
Modify response to #26

@@
 
-    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor.
-    This option turns off all validation, including the validation of the reference material
-    and enforcement of required fields. We recognize that this is a limitation at present, if users
-    wish to, for example, validate that their units are correct but do not wish to validate some
-    other aspect of the file. Work is ongoing to improve this situation, and discussion is contained
-    in a public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
+    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor,
+    which turns off all validation of the ChemKED file when it is loaded with PyKED. Work is ongoing
+    to allow more granular selection of fields to be validated, and discussion is contained in a
+    public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
     %

### 2017-10-08 rebuttal.tex
Tentative response to #29
Basically say "Not right now, stay tuned"
A true community effort for #30
Also mention that we're trying to play nice with others

@@
     %
-
+    At present, this is not possible; only a single \verb|composition| block is allowed in the
+    \verb|common-properties| section. However, we are considering modifying the behavior of the
+    \verb|common-properties| section, and discussion is ongoing in
+    \url{https://github.com/pr-omethe-us/PyKED/issues/67}.
     %
@@
 
-    All the existing database have a similar requirement, where users submit
-    data files for inclusion that are approved by the database owners after some review. In this
-    sense, we actually see the Pull Request mechanism of GitHub as an advantage, since all reviews
-    can be done publicly and some of the review procedures (style checks, reference validity, etc.)
-    may be done automatically by a script associated with the GitHub repository. In this way, the
-    entire community can contribute to improving the quality of the database.
+    Moreover, all the existing databases have a similar requirement (whether explicitly stated or
+    not), where users submit data files for inclusion that are approved by the database owners after
+    some review. In this sense, we actually see the Pull Request mechanism of GitHub as an
+    advantage, since all reviews can be done publicly and some of the review procedures (style
+    checks, reference validity, etc.) may be done automatically by a script associated with the
+    GitHub repository. In this way, the entire community can contribute to improving the quality of
+    the database.
 
-    Lastly, while we hope other researchers will contribute to this database (such that it becomes
-    more of a community effort than our effort), there is no requirement for others to use our
-    database to use ChemKED files or PyKED. Others are free to use the tools internally, or
-    download\slash fork our database and build their own.
+    Lastly, while we hope other researchers will contribute to this database (such that it becomes a
+    true community effort), there is no requirement for others to use our database to use ChemKED
+    files or PyKED. Others are free to use the tools internally, or download\slash fork our database
+    and build their own. Moreover, as mentioned in the manuscript, we have included the ability to
+    generate ReSpecTh-compliant files from ChemKED files, so that the community can move data to
+    alternative formats easily and so there is a reduced chance of a negative ``lock-in'' effect.
     %

### 2017-10-08 pyked.tex
Update open definition to include open knowledge

@@
 definition of ignition or a standard way to express detailed facility-specific
-effects necessary to properly simulate certain experiments. Second, the XML
-format is intended to be a machine-readable markup language rather than a data format.
+effects necessary to properly simulate certain experiments. Second, PrIMe uses
+internal identifiers for bibliographic references and
+species\slash reactions, rather than standard identifiers (e.g., DOIs for
+scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
+the PrIMe ecosystem.
+
+Third, XML is intended to be a machine-readable markup language rather than a data format.
 As a general-purpose document markup language, XML can represent detailed
@@
 limits the human readability of database files constructed in XML and presents a barrier to
-creating and working with them. Furthermore, PrIMe uses
-internal identifiers for bibliographic references and
-species\slash reactions, rather than standard identifiers (e.g., DOIs for
-scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
-the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
-and associated Data Warehouse limit contributions and community development of
-the PrIMe standard and compatible data files and tools.
-
-%discuss open source definition here
-(When we describe a software package or project as ``closed,'' we are generally
-contrasting it with the open source definition~\cite{opensource}. In brief, to be
-``open,'' a project must provide free and unrestricted access to source code and\slash
-or data, allow free redistribution, permit derived works, and not discriminate against
-any people or areas of work, along with a few other criteria.)
+creating and working with them.
+
+Finally, the closed and opaque natures of the PrIMe standard and associated Data Warehouse limit
+contributions and community development of the PrIMe standard and compatible data files and tools.
+In this sense, we mean ``closed'' as a contrast to open source, open science, and open data, as
+defined by the Open Knowledge Institute~\cite{openknowledge} or the Open Source
+Initiative~\cite{opensource}. Although most commonly applied to software, the distinction between
+open and closed is becoming increasingly important in science and research. In brief, the Open
+Knowledge Institute defines open knowledge as ``free to access, use, modify, and share... — subject,
+at most, to measures that preserve provenance and openness''~\cite{openknowledge}, while the Open
+Source Institute focuses on software and declares that a project must provide free and unrestricted
+access to the source code, allow free redistribution, permit derived works, and not discriminate
+against any people or areas of work, along with a few other criteria~\cite{opensource}.
 

### 2017-10-08 rebuttal.tex
Move PrIMe URL from footnote to text

@@
     %
-    We have added the new PrIMe URL to the manuscript on \textbf{page 2}, and thank the reviewer for pointing us to that.
-
-    However, we do not think that the technical details of the implementation of the original PrIMe system
-    are relevant to this work, so we have not included them. In addition, we cannot know
-    the plans of Prof.~Frenklach, and we do not feel it would be appropriate to comment on them in
-    our manuscript. As such, we have not added these suggestions to the paper.
+    We do not think that the technical details of the implementation of the original PrIMe system
+    (including a soon-to-be outdated URL for PrIMe 2.0) are relevant to this work, so we have not
+    included them. In addition, we cannot know the plans of Prof.~Frenklach, and we do not feel it
+    would be appropriate to comment on them in our manuscript. As such, we have not added these
+    suggestions to the paper. However, we have moved the URL for PrIMe 3.0 from a footnote into the
+    course of the main text on \textbf{page 2}.
     %

### 2017-10-08 pyked.tex
Move PrIMe URL from footnote to text

@@
 format~\autocite{Frenklach:2007bm,You:2011hy} and an associated online database,
-the PrIMe Data Warehouse.\footnote{\url{http://primekinetics.org}}
+the PrIMe Data Warehouse (available at \url{http://primekinetics.org}).
 PrIMe files encode fundamental
@@
 
-PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/})
+PrIMe~\autocite{Frenklach:2007bm,You:2011hy}
 has a number of features that make

### 2017-10-08 pyked.tex
Add more standards URLs to the text instead of footnotes

@@
 More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
-ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
+ReSpecTh standard (available at
+\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}),
+which builds on the PrIMe format. ReSpecTh adds important
 features that make files better standalone representations of experimental
@@
 the Clean Combustion Research Center at King Abdullah University of Science and
-Technology (KAUST): CloudFlame.\footnote{\url{https://cloudflame.kaust.edu.sa}}
+Technology (KAUST): CloudFlame (available at \url{https://cloudflame.kaust.edu.sa}).
 Started in 2013~\autocite{Goteng:2013cf,Goteng:2014,ReynoChiasson:2015}, CloudFlame

### 2017-10-09 pyked.tex
Update examples so they work

@@
 The \ck{} file that will be used in this example can be found in the
-\texttt{tests} directory of the PyKED
+\texttt{tests} directory of the \pk{}
 repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
@@
 history of the reactor is available in CanSen~\autocite{cansen}, in the
-\pybox|cansen.profile.VolumeProfile| class.
-%
-\begin{pythonbox}
-# Retrieve the time and volume from the
-# history in the datapoint
-time = dp.volume_history.time
-volume = dp.volume_history.volume
-# Install the Wall between the
-# reactor and environment
-ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))
+\pybox|cansen.profiles.VolumeProfile| class.
+%
+\begin{pythonbox}
+# Retrieve the time and volume from the history in the datapoint
+time = dp.volume_history.time.magnitude
+volume = dp.volume_history.volume.magnitude
+keywords = {'vproTime': time, 'vproVol': volume}
+# Install the Wall between the reactor and environment
+ct.Wall(reac, env, velocity=VolumeProfile(keywords))
 \end{pythonbox}
@@
 
-To calculate the ignition delay, we will follow the definition specified in the ChemKED file for
+To calculate the ignition delay, we will follow the definition specified in the \ck{} file for
 this experiment, where the experimentalists used the maximum of the time derivative of the pressure
@@
 \pybox|step| method takes one timestep forward on each call, with step size determined by the CVODES
-solver (CVODES is an adaptive time-stepping algorithm). On each step, we add the relevant variables
-to their respective lists. The problem is integrated until a user-specified end time, in this case
-\SI{50}{\milli\second}, although in principle, the user could end the simulation on any condition
-they choose:
+solver (CVODES uses an adaptive time-stepping algorithm). On each step, we add the relevant
+variables to their respective lists. The problem is integrated until a user-specified end time, in
+this case \SI{50}{\milli\second}, although in principle, the user could end the simulation on any
+condition they choose:
 
@@
 ignition delay. Therefore, a \yabox|common-properties| section is specified, followed by the
-datapoints list:
+\yabox|datapoints| list:
 %
@@
 of data in Python). Then, we use the built-in \pybox|map| function to apply the
-\pybox|collect_input| function to each of the elements in the \pybox|ck.datapoints| list: convert a
-\pybox|Quantity| to some input units and return the magnitude, then use Python's list comprehension
-syntax to generate a list of tuples (tuples are lightweight groups of data in Python):
+\pybox|collect_input| function to each of the elements in the \pybox|ck.datapoints| list:
 %
@@
 
-initial_conditions = map(collect_input, ck.datapoints)
-initial_conditions = [(to_SI(dp.temperature, 'K'), to_SI(dp.pressure, 'Pa'), dp.get_cantera_mole_fraction()) for dp in ck.datapoints]
-\end{pythonbox}
-
-Finally, we create the processor \pybox|Pool| (with 12 processes) and send the jobs out to run:
-%
-\begin{pythonbox}
-with Pool(12) as pool:
+initial_conditions = list(map(collect_input, ck.datapoints))
+\end{pythonbox}
+
+Finally, we create the processor \pybox|Pool| (with 4 processes) and send the jobs out to run:
+%
+\begin{pythonbox}
+with Pool(processes=4) as pool:
     ignition_delays = pool.starmap(run_simulation, initial_conditions)

### 2017-10-09 pyked.tex
Small update to ST example variable names

@@
 # Retrieve the time and volume from the history in the datapoint
-time = dp.volume_history.time.magnitude
-volume = dp.volume_history.volume.magnitude
-keywords = {'vproTime': time, 'vproVol': volume}
+exp_time = dp.volume_history.time.magnitude
+exp_volume = dp.volume_history.volume.magnitude
+keywords = {'vproTime': exp_time, 'vproVol': exp_volume}
 # Install the Wall between the reactor and environment
@@
 netw = ct.ReactorNet([reac])
-netw.set_max_time_step(np.min(np.diff(time)))
+netw.set_max_time_step(np.min(np.diff(exp_time)))
 \end{pythonbox}
@@
 for (T, P, X), tau in zip(initial_conditions, ignition_delays):
-    print(f'The ignition delay for T_initial={T}, P_initial={P} is: {tau} seconds')
+    print(f'The ignition delay for T_initial={T} K, P_initial={P} Pa is: {tau} seconds')
 \end{pythonbox}

### 2017-10-09 pyked.tex
Add link to example notebooks in GitHub

@@
 
-The following usage examples provide a guide to the use of \pk{}. They are by no
-means an exhaustive treatment, and are meant to demonstrate the basic
-capabilities of the software.
+The following usage examples provide a guide to the use of \pk{}. They are by no means an exhaustive
+treatment, and are meant to demonstrate the basic capabilities of the software. Both examples shown
+below are also available as Jupyter Notebook files from the GitHub repository for
+PyKED\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/docs/rcm-example.ipynb} and
+\url{https://github.com/pr-omethe-us/PyKED/blob/master/docs/shock-tube-example.ipynb}} and in the
+Supplementary Material associated with this article.
 

### 2017-10-09 pyked.tex
Add paragraph to conclusion about examples

@@
 
+Finally, we presented several examples of the use of \pk{} to interpret and interact with data
+stored in \ck{} files. Both examples are also available online as Jupyter Notebook files in the
+GitHub repository and in the Supplementary Material. The first example involved the simulation of a
+rapid compression machine experiment, including the facility-specific effects that are critical to
+accurately predict ignition delay experiments. The second example demonstrated the use of built-in
+Python libraries to automatically run simulations of several shock tube experiments in parallel,
+with initial conditions loaded directly from the \ck{} file.
+
 \ck{} and \pk{} are currently under active development, and contributions from the community are

### 2017-10-09 rebuttal.tex
Add paragraph to conclusion about examples
Clarify backwards compatiblity response

@@
     version number (vX.Y.Z) have specific meaning; changing the major version (X) indicates a
-    backwards incompatible change, changing the minor version (Y) indicates a backwards compatible,
-    but somewhat large, change, and changing the patch version (Z) indicates a bug fix with no
+    backwards incompatible change, changing the minor version (Y) indicates somewhat large change,
+    possibly requiring small backwards-compatibility fixes (but should generally be entirely
+    backwards-compatible), and changing the patch version (Z) indicates a bug fix with no
     anticipated incompatibilities.
@@
 \begin{response}
-
+    %
+    We have added a paragraph to the conclusion on \textbf{page 17}.
+    %
 \end{response}
 
-\section*{References}
-\bibliography{rebuttal}
-\bibliographystyle{elsarticle-num}
-
 \end{document}

### 2017-10-09 rebuttal.tex
Small changes to response for clarification

@@
     brief definition and link to the definition of open source, which comes from the
-    Open Source Initiative.
+    Open Source Initiative and the Open Knowledge Institute.
 \end{response}
@@
     We agree that submitting files to the ChemKED database on GitHub will inevitably incur some
-    delay, however we see the human intervention step as necessary and important. In much the same
+    delay; however, we see the human intervention step as necessary and important. In much the same
     way that review of manuscripts provides a certain surety of the minimal quality of the article,
@@
     viewed directly from Python using existing functionality in the Pandas DataFrame class
-    (\verb|DataFrame.plot()|).
+    (\verb|DataFrame.plot()|), which leverages the Matplotlib package under-the-hood.
     %

### 2017-10-09 rebuttal.tex
Remove mention of highlights that we didn't add from the response

@@
 
-We thank the reviewers for their helpful and detailed comments, and have addressed their remarks here.
-The manuscript has been significantly revised, and we marked significant changes corresponding to suggestions from Reviewers \#1 and \#2 in \hlg{yellow} and \hly{green}, respectively.
+We thank the reviewers for their helpful and detailed comments, and have addressed their remarks
+here. The manuscript has been significantly revised, and we have included a marked copy of the
+manuscript with additions in blue and deletions in red.
 

### 2017-10-09 rebuttal.tex
Final edit to close #26

@@
     We agree with the reviewer that some of the choices currently enforced in the PyKED package and
-    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
+    the ChemKED schema do not lend themselves to easy informal\slash working use. We presently recommend that
     users who wish to use ChemKED files in this way include an incomplete reference field. We have
-    modified the ChemKED schema in response to this comment so that the only required subfields of
-    the reference field are the authors and the year.
+    modified the ChemKED schema (and \textbf{page 5} of the manuscript) in response to this comment
+    so that the only required subfields of the reference field are the authors and the year.
 

### 2017-10-09 pyked.tex
minor wording edits in conclusions

--- 2017-10-09 pyked.tex
@@
 
-\ck{} and \pk{} are currently under active development, and contributions from the community are
-welcomed. All development occurs under the BSD 3-clause open source license and the code is housed
+Finally, we presented several examples using \pk{} to interpret and interact with data
+stored in \ck{} files. Both examples are also available online as Jupyter Notebook files in the
+GitHub repository and in the Supplementary Material. The first example simulates a
+rapid compression machine experiment, including the facility-specific effects critical to
+accurately predict ignition delay experiments. The second example demonstrates the use of built-in
+Python libraries to automatically simulate several shock tube experiments in parallel,
+with initial conditions loaded directly from the \ck{} file.
+
+We are actively developing \ck{} and \pk{}, and we welcome contributions from the community.
+All development occurs under the BSD 3-clause open source license and the code is housed
 on GitHub at \url{https://github.com/pr-omethe-us/PyKED}. Future directions for development are

### 2017-10-09 rebuttal.tex
minor wording fix in rebuttal

@@
 
-    Moreover, all the existing databases have a similar requirement (whether explicitly stated or
+    Moreover, all existing databases have a similar requirement (whether explicitly stated or
     not), where users submit data files for inclusion that are approved by the database owners after

### 2017-10-09 pyked.tex
minor edits to paragraph about open

@@
 open and closed is becoming increasingly important in science and research. In brief, the Open
-Knowledge Institute defines open knowledge as ``free to access, use, modify, and share... — subject,
-at most, to measures that preserve provenance and openness''~\cite{openknowledge}, while the Open
-Source Institute focuses on software and declares that a project must provide free and unrestricted
-access to the source code, allow free redistribution, permit derived works, and not discriminate
+Knowledge Institute defines open knowledge as ``free to access, use, modify, and
+share\textellipsis — subject, at most, to measures that preserve provenance and
+openness''~\cite{openknowledge}, while the Open Source Institute focuses on software and declares
+that a project must provide free and unrestricted access to the source code, allow free
+redistribution, permit derived works, and not discriminate
 against any people or areas of work, along with a few other criteria~\cite{opensource}.

### 2017-10-09 pyked.tex
extended shock tube example, with figure

@@
 Section~\ref{sec:rcm-modeling}.
-
 This \ck{} file specifies multiple data points with some common
@@
 ignition delay. Therefore, a \yabox|common-properties| section is specified, followed by the
-\yabox|datapoints| list:
+\yabox|datapoints| list (as before, we have truncated the \yabox|datapoints| list for space):
 %
@@
 \end{yamlbox}
-%
-As before, we have truncated the datapoints list for space.
 
@@
 pressure and temperature condition in the \yabox|datapoints| list. Once again,
-the \ck{} file specifies all the information required for the simulations except
+the \ck{} file specifies all information required for the simulations except
 for the chemical kinetic model, and Cantera can be used to simulate autoignition. The setup steps
-are the same as in the previous example, with one exception. In this example, we also import
-Python's built-in multiprocessing library so that we can run the simulations on multiple cores. We
-import the \pybox|Pool| class, which offers facilities to manage a jobs on a pool of processors:
+match those from the previous example, with one exception: in this example, we also import
+Python's built-in multiprocessing library so that we can run the simulations on multiple cores.
+We import the \pybox|Pool| class, which offers tools to manage a set of jobs on a
+pool of processors:
 %
@@
 \ce{OH^*} (provided that the kinetic model includes accurate chemistry for \ce{OH^*}) to match the
-definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
-be supplied by the user:
+definition of ignition delay in the experiments. We use the chemical kinetic model
+for butanol isomers from Sarathy et al.~\cite{Sarathy:2012fj}, represented as
+\verb|LLNL_sarathy_butanol.cti|:
 %
@@
 def run_simulation(T, P, X):
-    gas = ct.Solution('tbuoh-mechanism.cti')
+    gas = ct.Solution('LLNL_sarathy_butanol.cti')
     gas.TPX = T, P, X
@@
 
-Finally, we create the processor \pybox|Pool| (with 4 processes) and send the jobs out to run:
+Finally, we create the processor \pybox|Pool| (with four processes) and send the jobs out to run:
 %
@@
 The simulated ignition delay results are returned in the \pybox|ignition_delays| list. The results
-are printed to the screen using a Python formatted string (f-string).
+are printed to the screen using a Python formatted string (f-string). We can also visually
+compare the results using Matplotlib~\cite{matplotlib}, as
+Figure~\ref{fig:ignition-delays} shows:
+%
+\begin{pythonbox}
+import matplotlib.pyplot as plt
+
+inv_temps = [1000/i[0] for i in initial_conditions]
+exp_ignition_delays = [dp.ignition_delay.to('ms').magnitude for dp in ck.datapoints]
+sim_ignition_delays = np.array(ignition_delays)*1.0E3
+
+plt.figure()
+plt.scatter(inv_temps, exp_ignition_delays, label='Experimental ignition delays')
+plt.scatter(inv_temps, sim_ignition_delays, label='Simulated ignition delays', marker='s')
+plt.legend(loc='best')
+plt.yscale('log')
+plt.ylabel('Ignition delay [ms]')
+plt.xlabel('1000/T [1/K]')
+\end{pythonbox}
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=0.7\textwidth]{ignition-delays.pdf}
+    \caption{Comparison of experimental and simulated ignition delays of
+    \textit{tert}-butanol from shock tube example.}
+    \label{fig:ignition-delays}
+\end{figure}
+
 %====================================================================

### 2017-10-09 pyked.tex
extended RCM example, with figs

@@
 
+\usepackage[english]{babel}
+\usepackage{csquotes}
+\usepackage{textcomp}
+\usepackage{latexsym,amsmath,amssymb}
+
 \usepackage{fancyvrb,newverbs,xcolor}
+\usepackage{graphicx}
+\usepackage{caption,subcaption}
 
@@
 \end{pythonbox}
-%
+
 With these properties defined, we have to create the objects in Cantera that represent the physical
 state of the system to be studied. In Cantera, the \pybox|Solution| class stores the thermodynamic,
-kinetic, and transport data from an input file in the CTI format. After the \pybox|Solution| object
-is created, we can set the initial temperature, pressure, and mole fractions using the \pybox|TPX|
-attribute of the \pybox|Solution| class:
-
+kinetic, and transport data from the GRI Mech 3.0~\cite{GRI3.0} model included with Cantera.
+After the \pybox|Solution| object is created, we can set the initial temperature, pressure,
+and mole fractions using the \pybox|TPX| attribute of the \pybox|Solution| class:
+%
 \begin{pythonbox}
 # Load the mechanism and set the initial state of the mixture
-gas = ct.Solution('h2-co-mechanism.cti')
+gas = ct.Solution('gri30.cti')
 gas.TPX = T_initial, P_initial, X_initial
@@
 \verb|air.xml|, is included with Cantera and represents an average composition of air.
-
+%
 \begin{pythonbox}
@@
 \begin{pythonbox}
+from cansen.profiles import VolumeProfile
 # Retrieve the time and volume from the history in the datapoint
@@
 used in Cantera) to solve the energy and species equations. For this example, it is best practice
-to set the maximum time step allowed in the solution to be the minimum time difference in the time array from the volume trace:
-
+to set the maximum time step allowed in the solution to be the minimum time difference in
+the time array from the volume trace:
+%
 \begin{pythonbox}
@@
 this experiment, where the experimentalists used the maximum of the time derivative of the pressure
-to define the ignition delay. To calculate this derivative, we need to store the state variables and the composition on each time step, so we initialize several Python lists to act as storage:
-
+to define the ignition delay. To calculate this derivative, we need to store the state variables
+and the composition on each time step, so we initialize several Python lists to act as storage:
+%
 \begin{pythonbox}
@@
 condition they choose:
-
+%
 \begin{pythonbox}
@@
 At this point, the user would post-process the information in the \pybox|pressure| list to calculate
-the derivative by whatever algorithm they choose.
+the derivative by whatever algorithm they choose. Here, we plot pressure versus time from the
+simulation, and compare the simulated and experimental volume traces,
+using the Matplotlib library~\cite{matplotlib} and shown in Figure~\ref{fig:rcm}:
+%
+\begin{pythonbox}
+import matplotlib.pyplot as plt
+
+plt.figure()
+plt.plot(time, pressure)
+plt.ylabel('Pressure [Pa]')
+plt.xlabel('Time [s]')
+
+plt.figure()
+plt.plot(exp_time, exp_volume/exp_volume[0], label='Experimental volume', linestyle='--')
+plt.plot(time, volume, label='Simulated volume')
+plt.legend(loc='best')
+plt.ylabel('Volume [m^3]')
+plt.xlabel('Time [s]')
+\end{pythonbox}
+
+\begin{figure}[htbp]
+    \centering
+    \begin{subfigure}{0.7\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{rcm-pressure.pdf}
+        \caption{Simulated pressure trace}
+    \end{subfigure}
+    \\
+    \begin{subfigure}{0.7\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{rcm-volume.pdf}
+        \caption{Experimental and simulated volume histories}
+    \end{subfigure}
+    \caption{Results from RCM example}
+    \label{fig:rcm}
+\end{figure}
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[htbp]
     \centering

### 2017-10-09 rebuttal.tex
response for #20 on figures
version X.X.X -> 0.3.0

@@
 \begin{response}
-
+    In response to one of your later comments, we have extended our examples to make them
+    more usable as-is, and also added figures to show results from these examples on
+    \textbf{pages 16 and 19}.
 \end{response}
@@
     ChemKED database have been coupled up to now, and we have updated the version mentioned in the
-    text on \textbf{page 3} to vX.X.X, which we released in conjunction with submitting
+    text on \textbf{page 3} to v0.3.0, which we released in conjunction with submitting
     this response after the helpful comments of the reviewers.

### 2017-10-09 rebuttal.tex
lowercase "response"
updated page numbers in rebuttal

@@
 \newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup}
-\newenvironment{response}{\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
+\newenvironment{response}{\textbf{Our response:}}{\vspace{0.5\baselineskip}}
 \newcommand{\review}[1]{{\itshape#1}}
-\newcommand{\respond}[1]{\textbf{Our Response:} #1}
+\newcommand{\respond}[1]{\textbf{Our response:} #1}
 
@@
     %
-    We would like to thank the reviewer for their comments. Detailed responses to each comment are
+    We thank the reviewer for their comments. Detailed responses to each comment are
     below.
+
+    Regarding the manuscript, please see our response to your next point; in addition, we added more
+    detail to our examples in response to Reviewer \#2's comments, which increased the length of
+    the manuscript slightly.
     %
@@
     helping us define what data and information needs to go in the specification for their
-    experiment.
+    experiments.
     %
@@
     Thank you for the comment, we have corrected this oversight in the roadmap and the
-    conclusions (\textbf{page 17}).
+    conclusions (\textbf{page 20}).
     %
@@
     We have corrected this oversight, with more explanation and introduction of the Python
-    examples on \textbf{pages 12--17}.
+    examples on \textbf{pages 12--19}.
     %
@@
     %
-    We have truncated the datapoints list on \textbf{pages 12 and 16} as the reviewer suggests.
+    We have truncated the datapoints list on \textbf{pages 12 and 17} as the reviewer suggests.
     %
@@
     %
-    We have modified the example on \textbf{pages 16--17} to include more
+    We have modified the example on \textbf{pages 17--19} to include more
     explanatory text and to use Python's built-in
@@
     takes to run), how many processors are available, and the availability of the multi-threaded
-    BLAS and LAPACk libraries. However, we have not explained this detail of Cantera in the text to
+    BLAS and LAPACK libraries. However, we have not explained this detail of Cantera in the text to
     avoid confusing the reader.
@@
     %
-    We have added a paragraph to the conclusion on \textbf{page 17}.
+    We have added a paragraph to the conclusion on \textbf{page 20}.
     %

### 2017-10-10 pyked.tex
updated PyKED version in paper

@@
 
-This section provides an introduction to the current version of the \ck{} database format, v0.1.6,
+This section provides an introduction to the current version of the \ck{} database format, v0.3.0,
 which represents a functional standard for documenting autoignition measurements in shock tubes and

### 2017-11-14 pyked.tex
changes to paper and class for submission
IJCK submission system doesn't like minted, so for submission we have
to use the "frozencache" option.
made various easy changes for second round of reviewer comments

@@
-% !TEX option = -shell-escape
+%% !TEX option = -shell-escape
 %% comment out the above line before submitting
@@
 \setlist{noitemsep}
+
+% For the version with changes, use these four commands
+% *** Be very precise and careful about including whitespace and punctuation in your edits ***
+\usepackage[normalem]{ulem}
+\newcommand{\addone}[1]{{\sloppy\textcolor{teal}{\uline{#1}}}}  % add
+\newcommand{\deleteone}[1]{\sloppy\textcolor{teal}{\sout{#1}}}    % delete
+\newcommand{\addtwo}[1]{{\sloppy\textcolor{red}{\uline{#1}}}}  % add
+\newcommand{\deletetwo}[1]{\sloppy\textcolor{red}{\sout{#1}}}    % delete
 
@@
 hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models,
-and more recently develop improved models for hydrogen\slash
-syngas~\autocite{Varga:2016gj}, methanol\slash formaldehyde~\autocite{Olm:2017a},
-and ethanol~\autocite{Olm:2016et}. In all five cases, they converted numerous
+and more recently develop improved models for \addtwo{hydrogen~\autocite{Varga:2015hy}},
+hydrogen\slash syngas~\autocite{Varga:2016gj},
+methanol\slash formaldehyde~\autocite{Olm:2017a},
+and ethanol~\autocite{Olm:2016et}. In all \deletetwo{five}~\addtwo{six} cases, they converted numerous
 experimental datasets from the literature into the ReSpecTh standard.
-Their database currently contains \num{93240} experimental data points in 1376
+Their database currently contains \deletetwo{\num{93240}}~\addtwo{\num{93326}}
+experimental data points in \deletetwo{1376}~\addtwo{1396}
 ReSpecTh XML files~\autocite{respecth}.
@@
 top-level key for the sequence is called \yabox|datapoints|. For the current version of the \ck{}
-schema, only autoignition delay experiments are supported. The following information os required in
-each element of the sequence:
+schema, only autoignition delay experiments are supported. The following information
+\deleteone{os}~\addone{is} required in each element of the sequence:
 %
@@
 may not be possible, as \ck{} files support some data that ReSpecTh files do not support. We note
-that the reverse is also true: not all ReSpecTh files may be converted to \ck{} files because not
+that the reverse is also true: not all ReSpecTh files may be
+\addtwo{equivalently} converted to \ck{} files because not
 all of the data that can be stored in a ReSpecTh format is (currently) supported by the \ck{} format.

### 2017-11-14 pyked.tex
revision and response about ReSpecTh email to close #46

@@
-%% !TEX option = -shell-escape
+% !TEX option = -shell-escape
 %% comment out the above line before submitting
@@
 XML-based format, and, as such, suffer from the same readability issues as PrIMe.
-Moreover, the standard remains inaccessible to the public, requiring potential users
+Moreover, \deletetwo{the standard remains inaccessible to the public, requiring potential users
 to register with the standard's authors using a valid institutional email address
-to access the specification and the growing database.
+to access the specification and the growing database}
+\addtwo{potential users must register using a valid institutional email address
+to access the growing database}.
 

### 2017-11-15 pyked.tex
clarify description of RCM attributes to address #47

@@
             cubed
-            \item \yabox|column| (required, integer): the 0-based index of the column storing the
-            volume information in the \yabox|values| array
+            \item \yabox|column| (required, integer): the \deletetwo{0-based} \addtwo{zero-based} index of the column storing the
+            volume information in the \yabox|values| array \addtwo{(e.g., 0 or 1)}
         \end{itemize}
@@
             \item \yabox|units| (required, string): the units of the time
-            \item \yabox|column| (required, integer): the 0-based index of the column storing the
-            time information in the \yabox|values| array
-        \end{itemize}
-        \item \yabox|values| (required, sequence): A sequence of sequences describing the values of
-        the volume at the time points.
+            \item \yabox|column| (required, integer): the \deletetwo{0-based} \addtwo{zero-based} index of the column storing the
+            time information in the \yabox|values| array \addtwo{(e.g., 0 or 1)}
+        \end{itemize}
+        \item \yabox|values| (required, sequence): A \deletetwo{sequence of sequences describing the values of
+        the volume at the time points.}\addtwo{sequence of time-volume pairs describing the values
+        of the volume at different times}
     \end{itemize}

### 2017-11-15 pyked.tex
new footnote and response to address #48

@@
     \begin{itemize}
-        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history.
+        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history\footnote{\addtwo{Future versions
+        of \ck{} will support specifying additional time histories for rapid compression machine and shock tubes (e.g.,
+        pressure, volume, light emission, OH emission). These will follow the specification format given for volume.}}
         The fields in this mapping are:

### 2017-11-15 pyked.tex
add "experiments" to footnote about RCM & ST time-histories plan

--- 2017-11-15 pyked.tex
@@
     \begin{itemize}
-        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history.
+        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history\footnote{\addtwo{Future versions
+        of \ck{} will support specifying additional time histories for rapid compression machine and shock tube
+        experiments (e.g., pressure, volume, light emission, OH emission). These will follow the specification format given for volume.}}
         The fields in this mapping are:

### 2017-11-16 pyked.tex
removed formatting for final revision; this is the final version for IJCK

--- 2017-11-15 pyked.tex
@@
 % *** Be very precise and careful about including whitespace and punctuation in your edits ***
-\usepackage[normalem]{ulem}
-\newcommand{\addone}[1]{{\sloppy\textcolor{teal}{\uline{#1}}}}  % add
-\newcommand{\deleteone}[1]{\sloppy\textcolor{teal}{\sout{#1}}}    % delete
-\newcommand{\addtwo}[1]{{\sloppy\textcolor{red}{\uline{#1}}}}  % add
-\newcommand{\deletetwo}[1]{\sloppy\textcolor{red}{\sout{#1}}}    % delete
+% \usepackage[normalem]{ulem}
+% \newcommand{\addone}[1]{{\sloppy\textcolor{teal}{\uline{#1}}}}  % add
+% \newcommand{\deleteone}[1]{\sloppy\textcolor{teal}{\sout{#1}}}    % delete
+% \newcommand{\addtwo}[1]{{\sloppy\textcolor{red}{\uline{#1}}}}  % add
+% \newcommand{\deletetwo}[1]{\sloppy\textcolor{red}{\sout{#1}}}    % delete
 
@@
 XML-based format, and, as such, suffer from the same readability issues as PrIMe.
-Moreover, \deletetwo{the standard remains inaccessible to the public, requiring potential users
-to register with the standard's authors using a valid institutional email address
-to access the specification and the growing database}
-\addtwo{potential users must register using a valid institutional email address
-to access the growing database}.
+Moreover, potential users must register using a valid institutional email address
+to access the growing database.
 
@@
 hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models,
-and more recently develop improved models for \addtwo{hydrogen~\autocite{Varga:2015hy}},
+and more recently develop improved models for hydrogen~\autocite{Varga:2015hy},
 hydrogen\slash syngas~\autocite{Varga:2016gj},
 methanol\slash formaldehyde~\autocite{Olm:2017a},
-and ethanol~\autocite{Olm:2016et}. In all \deletetwo{five}~\addtwo{six} cases, they converted numerous
+and ethanol~\autocite{Olm:2016et}. In all six cases, they converted numerous
 experimental datasets from the literature into the ReSpecTh standard.
-Their database currently contains \deletetwo{\num{93240}}~\addtwo{\num{93326}}
-experimental data points in \deletetwo{1376}~\addtwo{1396}
-ReSpecTh XML files~\autocite{respecth}.
+Their database currently contains \num{93326}
+experimental data points in 1396 ReSpecTh XML files~\autocite{respecth}.
 
@@
 schema, only autoignition delay experiments are supported. The following information
-\deleteone{os}~\addone{is} required in each element of the sequence:
+is required in each element of the sequence:
 %
@@
     \begin{itemize}
-        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history.
+        \item \yabox|volume| (required, mapping): A mapping describing the volume in the history\footnote{Future versions
+        of \ck{} will support specifying additional time histories for rapid compression machine and shock tube
+        experiments (e.g., pressure, volume, light emission, OH emission). These will follow the specification format given for volume.}
         The fields in this mapping are:
@@
             cubed
-            \item \yabox|column| (required, integer): the \deletetwo{0-based} \addtwo{zero-based} index of the column storing the
-            volume information in the \yabox|values| array \addtwo{(e.g., 0 or 1)}
+            \item \yabox|column| (required, integer): the zero-based index of the column storing the
+            volume information in the \yabox|values| array (e.g., 0 or 1)
         \end{itemize}
@@
             \item \yabox|units| (required, string): the units of the time
-            \item \yabox|column| (required, integer): the \deletetwo{0-based} \addtwo{zero-based} index of the column storing the
-            time information in the \yabox|values| array \addtwo{(e.g., 0 or 1)}
-        \end{itemize}
-        \item \yabox|values| (required, sequence): A \deletetwo{sequence of sequences describing the values of
-        the volume at the time points.}\addtwo{sequence of time-volume pairs describing the values
-        of the volume at different times}
+            \item \yabox|column| (required, integer): the zero-based index of the column storing the
+            time information in the \yabox|values| array (e.g., 0 or 1)
+        \end{itemize}
+        \item \yabox|values| (required, sequence): A sequence of time-volume pairs describing the values
+        of the volume at different times
     \end{itemize}
@@
 that the reverse is also true: not all ReSpecTh files may be
-\addtwo{equivalently} converted to \ck{} files because not
+equivalently converted to \ck{} files because not
 all of the data that can be stored in a ReSpecTh format is (currently) supported by the \ck{} format.
